### 2016-08-08 docs/latex/ver1.tex
added generative model doc

\documentclass[12pt]{article}
\pdfoutput=1

\addtolength{\oddsidemargin}{-.875in}
\addtolength{\evensidemargin}{-.875in}
\addtolength{\textwidth}{1.75in}

\addtolength{\topmargin}{-.875in}
\addtolength{\textheight}{1.75in}

\openup 1em

%macro for commenting
\usepackage{color}
\newcommand{\leo}[1]{{\color{blue}{\it leo: #1}}}
\newcommand{\xbeta}{\boldsymbol X_i^T \boldsymbol\beta}

\usepackage[round]{natbib}

\usepackage{rotating}
\usepackage{graphicx}
\usepackage{subcaption}

\usepackage{float}

\usepackage{amsthm,amsmath} 
\usepackage{amssymb}
\usepackage{subcaption}
\captionsetup{compatibility=false}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}

\thispagestyle{empty}
\baselineskip=28pt

\title
    {{jk-means}}

\date{}

\begin{document}

\maketitle

\section{Generative Model}

For $i=1\ldots n$, 
\begin{equation}
\begin{aligned}
(w_{1},\ldots,w_{K}) & \sim Dir(\alpha,\ldots,\alpha)\\
Pr(u_{i1},\ldots,u_{iK}) & = \frac{1}{\binom{K}{J}} \\
m_i& \sim Multi(w_1 u_{i1},\ldots,w_K u_{iK})\\
y_i& \sim G(\theta_{m_i})
\end{aligned}
\end{equation}
where $(u_{i1},\ldots,u_{iK})$ is a random sample vector with $J$ $1$'s and $(K-J)$ 0's.

\bibliographystyle{plainnat}
\bibliography{reference}

\end{document}

### 2016-08-14 Rmd/1DimTest.Rmd
upload Rmd file

```
###Prerequisite
###Install the package if you haven't
setwd("~/git/")
require('devtools')
build('jkmeans')
install.packages("jkmeans_1.0.tar.gz", repos = NULL, type = "source")
```

```{r include=FALSE}
require('ggplot2')
require('reshape')
```

#Running jk-means, step-by-step

###1. Generate Data
Let's generate data from 3 clusters, each with n=100

N(1,1)
N(2,1)
N(6,1)

```{r}
mu<- rep(c(1,2,6),each=100)
table(mu)
y<- rnorm(100*3, mu, 1)
Z<- matrix(0,300,3)
clusterLabel<- rep(c(1:3),each=100)
for(i in 1:300){
  Z[i,clusterLabel[i]]<-1
}

hist(y,breaks=50)
```

###2. Run jk-means
Run with jk-means with j=2, k=3:

options: 
max iteration=100
convergence error tolerance = 1E-8
```{r}
K=3
j=2
jk <- jkmeans::jkmeansEM(as.matrix(y), K, j ,  1000, tol = 1E-8)

```

###3. Results
Here are the estimates for $\mu_k$ (sorted in ascending order)
```{r}
new_order<- order(jk$mu)
mu_est<- jk$mu[new_order]
```

Here are the cluster assignment probability:

$$\zeta_{ik} = \frac{N(y_i|\mu_k,\sigma^2)}{\sum_k N(y_i|\mu_k,\sigma^2)}$$

```{r echo=FALSE}
#set up a coloring scheme using colorRampPalette
zeta_est<- jk$zeta[,new_order]
m <-melt(zeta_est)
names(m)<- c("i","k","prob")
p <- ggplot(data=m, aes(x=i, y=k, fill=prob)) + geom_tile()
p<- p+ scale_fill_continuous(low = "#FFFFFF", high="red", name = "")
p
```

###4. Validation

Compute RMSE
```{r}
RMSE <- function(a, b) {
  sqrt(mean((a - b) ^ 2))
}

RMSE(c(1,2,6),mu_est)

```

Use $\zeta_{ik}$ as  estimator for cluster assignment, compute the expected misclassification loss for each $\sum_k \zeta_{ik} (1- Z_{ik})$:

```{r}

#misclassified ones:
m <-rowSums( zeta_est*(1-Z))

#misclassification rate
sum(m)/300
```

Use $\hat{Z}_{i}= \max_k\zeta_{ik}$ as MAP estimator for cluster assignment, compute the misclassification error ($\sum_k \delta_{\hat{Z}_{i},k}(1-{Z_{ik}})$):

```{r}

misclsError<-function(zeta_est, n, Z){
  Z_est <- t(apply(zeta_est, MARGIN = 1, function(x){x==max(x)}))*1
  m <-rowSums(Z_est*(1-Z))
  sum(m)/n
}

#misclassification rate
misclsError(zeta_est,300, Z)
```

#Run jk-means with different n (repeat 50 times for each)

Define the functions:
```{r}

jkmeansExperiment<-function(n,j=3, sigma=1){

  mu<- rep(c(1,2,6),each=n)

  Z<- matrix(0,n*3,3)
  clusterLabel<- rep(c(1:3),each=n)
  for(i in 1:(n*3)){
    Z[i,clusterLabel[i]]<-1
  }

  y<- as.matrix(rnorm(n*3, mu, sigma))

  jk <- jkmeans::jkmeansEM(y, K, j=j ,  1000, tol = 1E-8)
  new_order<- order(jk$mu)
  mu_est<- jk$mu[new_order]
  zeta_est<- jk$zeta[,new_order]
c(RMSE(c(1,2,6),mu_est),misclsError(zeta_est,3*n,Z))
}

runExp<- function(j=3,n=100, m=50){

  rmse<- numeric(m)
  mce<- numeric(m)

  for(i in 1:m){
      result<- jkmeansExperiment(n,j=j)
      rmse[i]<- result[1]
      mce[i]<- result[2]
  }
  return(list("RMSE"=rmse,"MCE"=mce))
}

```

Run n=seq(10,500,by=10), get average RMSE and MCE over 50 times for each n:

run with $j=3$
```{r}

n_seq<- seq(10,500,by=10)

avg_rmse_3<- numeric(length(n_seq))
avg_mce_3<- numeric(length(n_seq))

for(i in 1:length(n_seq)){
  n<- n_seq[i]
  Exp<- runExp(j=3,n=n,m=50)
  avg_rmse_3[i]<- mean(Exp$RMSE)
  avg_mce_3[i]<- mean(Exp$MCE)
  print(i)
}

```

run with $j=2$
```{r}

avg_rmse_2<- numeric(length(n_seq))
avg_mce_2<- numeric(length(n_seq))

for(i in 1:length(n_seq)){
  n<- n_seq[i]
  Exp<- runExp(j=2,n=n,m=50)
  avg_rmse_2[i]<- mean(Exp$RMSE)
  avg_mce_2[i]<- mean(Exp$MCE)
  print(i)
}

```

run with $j=1$
```{r}

avg_rmse_1<- numeric(length(n_seq))
avg_mce_1<- numeric(length(n_seq))

for(i in 1:length(n_seq)){
  n<- n_seq[i]
  Exp<- runExp(j=1,n=n,m=50)
  avg_rmse_1[i]<- mean(Exp$RMSE)
  avg_mce_1[i]<- mean(Exp$MCE)
  print(i)
}

```

Plot of the RMSE vs n

```{r}

plot1<- data.frame( "n" = rep(n_seq, 3),
            "RMSE_avg"=c(avg_rmse_1,avg_rmse_2,avg_rmse_3),
            "j"= as.factor(rep(c(1:3),each= length(n_seq)))
            )

p<- ggplot(data=plot1, aes())
p<- p+  geom_line(aes(x= n, y=RMSE_avg, colour=j)) 
p<- p+ theme_bw() + labs(title = "mu = 1,2,6")
p
```

Plot of the misclassification error vs n

```{r}

plot2<- data.frame( "n" = rep(n_seq, 3),
            "MCE_avg"=c(avg_mce_1,avg_mce_2,avg_mce_3),
            "j"= as.factor(rep(c(1:3),each= length(n_seq)))
            )

p<- ggplot(data=plot2, aes())
p<- p+  geom_line(aes(x= n, y=MCE_avg, colour=j)) 
p<- p+ theme_bw() + labs(title = "mu = 1,2,6")
p
```

### 2016-08-14 Rmd/render.r

@@ -1,0 +1,4 @@
+require("rmarkdown")
+rmarkdown::render("1DimTest.Rmd")
+
+

### 2016-08-14 src/mixture.hpp

@@ -248,7 +248,7 @@ class Mixture {
 
 
 
-      cout << newLoglik << endl;
+      //cout << newLoglik << endl;
     }
   }
 
@@ -267,7 +267,7 @@ class Mixture {
       else
         mu0 = mu;
 
-      cout << compTotalLoglik() << endl;
+     // cout << compTotalLoglik() << endl;
     }
   }
 

### 2016-08-29 Rmd/bigKtest.Rmd
First draft of a K=10 test :construction: [neurodata/weekly-experiments#196]

---
title: "$jk$-means tests"
author: "Jesse Leigh Patsolic"
date: '`r Sys.Date()`'
output:
  html_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 5
    fig_retina: 2.7
    highlight: pygments
    keep_md: yes
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
---

```{r cc-setup, eval=FALSE, include=FALSE}
require(rmarkdown)
rmarkdown::render("bigKtest.Rmd")
system('open bigKtest.html')
```

```{r knitrOpts, include=FALSE, results='asis', message=FALSE, warning=FALSE}
require(rmarkdown)
require(knitr)
knitr::opts_chunk$set(cache=TRUE, autodep=TRUE)
dep_auto() # figure out dependencies automatically
opts_chunk$set(cache=TRUE, echo=TRUE, 
               opts_chunk$set(fig.path = './'),
               warning=FALSE, message=FALSE, 
               comment="#",
               dpi=227,dev=c('png','pdf'))
```

```{r include=FALSE}
require('ggplot2')
require('reshape')
require(mvtnorm)
require(plotly)
require(jkmeans)
require(mclust)
require(gridExtra)
source("~/neurodata/synaptome-stats/Code/kmpp.r")
```

# Generate Data

Let $X \sim \sum_i^k \pi_i \mathcal{N}_i(\mu_i, \sigma^2_i I)$. 
With $k = 10$ and $n = 1000$, $N = nk$.

```{r cc-genParams}
n <- 1e3
k <- 10
N <- n*k

set.seed(49)
(mux <- sort(sample(-20:20, 10)))
(muy <- sample(-20:20, 10))

(mu0 <- cbind(mux, muy))
(sigma1 <- sample(c(1,1,1,2,3,4,5), 10, replace=TRUE))
sigma0 <- lapply(sigma1, function(x) matrix(x*c(1,0,0,1), 2, 2))

Z <- rep(1:k, each = n)
```

Here are the means
```{r p-means}
plot(mu0, col = 1:10, pch = 19, cex = 3)
```

We now generate the data with 
```{r cc-genXs}
X0 <- data.frame(t(sapply(Z, function(x) rmvnorm(1, mu0[x,], sigma0[[x]]))), Z)
names(X0) <- c('x1', 'x2', 'Z')

head(X0[sample(dim(X0)[1]),])
dat <- as.matrix(X0[, 1:2])
```

And here are the data colored with the truth.
```{r p-X0s}
ggCol <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a')
ggplot(data = X0, aes(x=x1, y=x2, color = as.factor(Z))) +
  scale_color_manual(values = ggCol) + 
  geom_point(alpha = 0.5)
```

# Run algorithms

## Run $jk$-means
Run with $jk$-means with $j=2, k=10$:

options: 
max iteration=100
convergence error tolerance = 1E-8
```{r}
K <- k
j <- 5

set.seed(5431)
set.seed(2^10)
system.time(
  jk <- jkmeans::jkmeansEM(dat, K, j , 100, tol = 1E-8)
)
jkv <- apply(jk$zeta, 1,  function(x) which( x == max(x) ))
```

## Run $k=10$-means initialized at the truth

```{r cc-km}
system.time({
  kv <- kmeans(dat, centers = mu0)
}
)
```

## Run MCLUST

```{r cc-mclust}
system.time(
  mcv <- Mclust(dat, 
              G = K, 
              modelNames="EII")
)
mcv
```
# Results

```{r, echo=FALSE}
ari <- c(mclust::adjustedRandIndex(Z, kv$cluster),
mclust::adjustedRandIndex(Z, jkv),
mclust::adjustedRandIndex(Z, mcv$classification))

Names <- c("Truth v km", "Truth v jk", "Truth v mclust")

data.frame(Names, ari)
```

```{r p-results}
Y <- data.frame(X0, jk=jkv, kv = kv$cluster,
                mc=mcv$classification, kvpp=kvpp$cluster)

p1 <- ggplot(data = Y, aes(x=x1, y=x2)) +
  scale_color_manual(values = ggCol) 
```

```{r p-all, echo=FALSE, fig.height=12,fig.width=16}
layoutm <- matrix(c(NA,2,NA,3,1,4,NA,NA,NA),3,3)

grid.arrange(
 p1 + geom_point(alpha = 0.25, aes(color = as.factor(Z))) +
   ggtitle("Truth"),
 p1 + geom_point(alpha = 0.25, aes(color = as.factor(jkv))) +
   ggtitle("(2,10)=jk-means"),
 p1 + geom_point(alpha = 0.25, aes(color = as.factor(mc))) +
   ggtitle("10=k, mclust"),
 p1 + geom_point(alpha = 0.25, aes(color = as.factor(kv))) +
   ggtitle("10=k-means"),
# p1 + geom_point(alpha = 0.25, aes(color = as.factor(kvpp))) +
#   ggtitle("10=k-meanspp"), 
 layout_matrix = layoutm
)
```

### 2016-08-29 Rmd/bigKtest.Rmd

@@
 set.seed(49)
-(mux <- sort(sample(-20:20, 10)))
-(muy <- sample(-20:20, 10))
+mux <- sort(sample(-20:20, 10))
+muy <- sample(-20:20, 10)
 
@@
 K <- k
-j <- 5
+j <- 2
 

### 2016-08-29 Rmd/bigKtest.Rmd
started 1d stuff @jovo [neurodata/weekly-experiments#196]

@@
 
-```{r cc-setup, eval=FALSE, include=FALSE}
+```{r cc-run, eval=FALSE, include=FALSE}
 require(rmarkdown)
@@
 
-# Generate Data
+# Generate Data 1d
+
+Let $X \sim \sum_i^k \pi_i \mathcal{N}_i(\mu_i, \sigma^2)$. 
+With $k = 10$ and $n = 1000$, $N = nk$, $\pi_i = 1/n$.
+
+```{r cc-1genParams}
+n <- 2e3
+k <- 10
+N <- n*k
+
+set.seed(53)
+mux <- c(seq(1,5,by=1),8, 10,15, 20, 25)
+
+mu0 <- mux
+(sigma1 <- rep(1,10))
+
+Z <- rep(1:k, each = n)
+```
+
+Here are the means
+```{r p-1means}
+plot(mu0, col = 1:10, pch = 19, cex = 3)
+```
+
+We now generate the data with 
+```{r cc-1genXs}
+X0 <- data.frame(x1 = sapply(Z, function(x) rnorm(1, mu0[x], sigma1[x])), Z = Z)
+
+head(X0[sample(dim(X0)[1]),])
+dat <- X0[, 1]
+```
+
+And here are the data colored with the truth.
+```{r p-1X0s}
+ggCol <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a')
+ggplot(data = X0, aes(x=x1, y = Z, ,color = as.factor(Z))) +
+  scale_color_manual(values = ggCol) + 
+  geom_point(alpha = 0.5)
+```
+
+# Run algorithms
+
+## Run $jk$-means
+Run with $jk$-means with $j=2, k=10$:
+
+options: 
+max iteration=100
+convergence error tolerance = 1E-8
+```{r, eval=FALSE}
+K <- k
+j <- 2
+
+set.seed(2^10)
+system.time(
+  jk <- jkmeans::jkmeansEM(as.matrix(dat), K, j , 100, tol = 1E-8)
+)
+jkv <- apply(jk$zeta, 1,  function(x) which( x == max(x) ))
+```
+
+## Run $k=10$-means initialized at the truth
+
+```{r cc-1km}
+system.time({
+  kv <- kmeans(dat, centers = mu0)
+}
+)
+```
+
+## Run MCLUST
+
+```{r cc-1mclust}
+system.time(
+  mcv <- Mclust(dat, 
+              G = K, 
+              modelNames="E")
+)
+mcv
+```
+# Results
+
+```{r, echo=FALSE}
+ari <- c(mclust::adjustedRandIndex(Z, kv$cluster),
+         mclust::adjustedRandIndex(Z, jkv),
+         mclust::adjustedRandIndex(Z, mcv$classification))
+
+Names <- c("Truth v km", "Truth v jk", "Truth v mclust")
+
+data.frame(Names, ari)
+```
+
+```{r p-1results}
+Y <- data.frame(X0, kv = kv$cluster,
+                mc=mcv$classification,
+                jk=jkv)
+
+p1d <- ggplot(data = Y, aes(x=x1, y=Z)) +
+  scale_color_manual(values = ggCol) 
+```
+
+```{r p-1all, echo=FALSE, fig.height=12,fig.width=16}
+layoutm <- matrix(c(NA,2,NA,3,1,4,NA,NA,NA),3,3)
+
+grid.arrange(
+ p1d + geom_point(alpha = 0.25, aes(color = as.factor(Z))) +
+   ggtitle("Truth"),
+ p1d + geom_point(alpha = 0.25, aes(color = as.factor(mc))) +
+   ggtitle("10=k, mclust"),
+ p1d + geom_point(alpha = 0.25, aes(color = as.factor(kv))) +
+   ggtitle("10=k-means"),
+ p1d + geom_point(alpha = 0.25, aes(color = as.factor(jk))) +
+   ggtitle("(2,10)=jk-means"),
+ layout_matrix = layoutm
+)
+```
+
+# Generate Data 2d
 
@@
 Y <- data.frame(X0, jk=jkv, kv = kv$cluster,
-                mc=mcv$classification, kvpp=kvpp$cluster)
+                mc=mcv$classification)
 

### 2016-09-01 docs/latex/ver1.tex
add new write-up for jkmeans

@@
 
+\usepackage[utf8]{inputenc}
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+
 \newtheorem{theorem}{Theorem}
@@
 \title
-    {{jk-means}}
+    {{Sparse Mixture Model (with jk-means a special case)}}
 
@@
 
-\section{Generative Model}
+\section{Sparse Mixture Model}
 
-For $i=1\ldots n$, 
+\subsection{General Framework}
+
+On a separable metrics space $\{\mathcal{X},\mathcal{F}\}$, we assume each data $y_i$ has the following probabilty density as the likelihood:
+
 \begin{equation}
 \begin{aligned}
-(w_{1},\ldots,w_{K}) & \sim Dir(\alpha,\ldots,\alpha)\\
-Pr(u_{i1},\ldots,u_{iK}) & = \frac{1}{\binom{K}{J}} \\
-m_i& \sim Multi(w_1 u_{i1},\ldots,w_K u_{iK})\\
-y_i& \sim G(\theta_{m_i})
+& [ y_i ] = \sum_{k=1}^{\kappa} w_{k,i} \delta_{\theta_{k}}(y_i) \\
+& w_{k,i} = \frac{ u_{k,i} v_{k} } { \sum_{k=1}^{\kappa}  u_{k,i} v_{k} }\\
+& u_{k,i} \in \{0,1\}
+\end{aligned}
+\label{marginal_lik}
+\end{equation}
+where $\delta_{\theta_{k}}(y_i)$ is a probabilty density corresponds to a component distribution with parameter $\theta_{k}$, and $w_{k,i}$ is the component weight that varies by index $i$ and $\sum_{k=1}^{\kappa} w_{k,i} =1$.
+The weight varying is due to the randomness in $u_{k,i}$.
+Due to the $0$'s $u_{k,i}$ introduces, this induces sparsity in the membership probabilty $w_{k,i}$.
+Therefore, we refer this as the sparse mixture model.
+
+\subsection{JK-Means}
+
+With the above general framework, we focus on a special case $\{ u_{1,i},u_{2,i},\ldots u_{\kappa,i}\} \sim \text{ J out of K draws}$ and $\nu_k = 1/\kappa$.
+We refer this as JK-Means as when $J=1$, this reduces to standard K-means method.
+
+We now use the latent variable $\{ z_{1,i},z_{2,i},\ldots z_{\kappa,i} \} \sim MultiNomial (\{ w_{1,i},w_{2,i},\ldots w_{\kappa,i}\} )$ that takes value of $\{0,\ldots,1,\ldots 0\}$ that assign randomly one $1$ to one of $\kappa$ vertices in the simplex.
+The likelihood can be rewritten as:
+
+\begin{equation}
+\begin{aligned}
+& [ y_i ] =   \sum_{k=1}^{\kappa} \int z_{k,i} \delta_{\theta_{k}}(y_i) P(d z_{k,i}) = \int \sum_{k=1}^{\kappa} z_{k,i} \delta_{\theta_{k}}(y_i) P(d z_{k,i})
 \end{aligned}
 \end{equation}
-where $(u_{i1},\ldots,u_{iK})$ is a random sample vector with $J$ $1$'s and $(K-J)$ 0's.
+where $P(d z_{k,i})$ is the measure of the multinomial distribution aforementioned, the last equation is due to Fubini theorem.
+Note due to the $z_{k,i}$ takes only one $1$ out of $K$, this allows a simple log-density augmented with $z_{k,i}$:
+
+\begin{equation}
+\begin{aligned}
+& \log[ \{y_i\}_i, \{z_{k,i}\}_{k,i} ] =  \sum_{i}\sum_{k=1}^{\kappa} z_{k,i} \log \{\delta_{\theta_{k}}(y_i) v_k \}
+\end{aligned}
+\label{conditional_lik}
+\end{equation}
+
+ This allows us to utilize Expectation-Maximization algorithm for parameter estimation:
+
+\begin{algorithm}[H]
+\caption{EM algorithm for JK-Means}\label{alg:euclid}
+\begin{algorithmic}[1]
+% \Procedure{Euclid}{$a,b$}\Comment{The g.c.d. of a and b}
+% \State $r\gets a\bmod b$
+\While{$||\theta^{(s)}-\theta^{(s+1)}||>\epsilon$}%\Comment{We have the answer if r is 0}
+\State Maximize over $\{ u_{i,k}\}_k$ by $ \underset{\{u_{i,k}\}_k} {\text{argmax }}\{u_{i,k}  v_{k}\delta_{\theta_{k}}(y_i)\}$ for all $i$; \Comment{\small pick top J out K based on (\ref{marginal_lik})}
+\State Compute $ \mathbb{E}(z_{i,k})= \frac{w_{k,i} \delta_{\theta_{k}}(y_i)}{\sum_k w_{k,i} \delta_{\theta_{k}}(y_i)}$ with $w_{k,i} = u_{i,k} v_{k}$; \Comment{E step  based on (\ref{conditional_lik})}
+\State Set $\theta^{(s+1)}_{k}= \text{argmax} \sum_{k=1}^{\kappa} \mathbb{E}z_{k,i} \log \delta_{\theta_{k}}(y_i) $ for all $k$ \Comment{M step: obtain MLE based on (\ref{conditional_lik})}
+\EndWhile\label{euclidendwhile}
+% \State \textbf{return} $b$\Comment{The gcd is b}
+% \EndProcedure
+\end{algorithmic}
+\end{algorithm}
 

### 2016-09-01 Rmd/RenzeTest.Rmd
tested old setting

```
###Prerequisite
###Install the package if you haven't
setwd("~/git/")
require('devtools')
build('jkmeans')
install.packages("jkmeans_1.0.tar.gz", repos = NULL, type = "source")
```

```{r include=FALSE}
require('ggplot2')
require('reshape')
```

###1. Generate Data
Let's generate data from 2 clusters, each of size n

N( c(0.7,0.3), diag(1.7661/n))
N( c(0.3,0.7), diag(1.7661/n))

```{r}
n<- 30
K<- 2
p<- 2
batchN<- 1000

yBatch<- array(0,dim = c(n*K,p,batchN))

mu0<- c(c(0.7,0.3), c(0.3,0.7))
mu<- matrix( 0,n*2,p)
mu[1:n,]<- rep(c(0.7,0.3),each=n)
mu[(n+1):(2*n),]<- rep(c(0.3,0.7),each=n)

sigma<-  sqrt(1.7661/n)

for(b in 1:batchN){
  yBatch[,,b]<- matrix( rnorm(n*p*2,mu,sigma), n*K, p)
}

mu_ini<- matrix(mu0,K,p)

computeMError<- function(jk, n){
  error<- numeric(batchN)
  for(b in 1:batchN){
    newOrder<- order(jk$mu[,1,b],decreasing = T)
    trueM<- rep( c(0,1)[newOrder], each=n)
    error[b]<- sum (jk$M[,b] != trueM) /2/n
  }
  error
}

```

###2. Test jk-means with different j

Run 12-Means, get average error

```{r}

jk12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini)

error12 <- computeMError(jk12,n)

mean(error12)
```

Run 22-Means, get average error

```{r}

jk22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini)

error22 <- computeMError(jk22,n)

mean(error22)
```

Plot overlay error rate histograms:

```{r}

dataHist<- data.frame("J"=rep(c(1,2),each=batchN),
                  "ErrorRate"= c(error12,error22))

ggplot(dataHist, aes(ErrorRate, fill = as.factor(J))) + geom_histogram(alpha = 0.2,bins = 50,  position="identity")
```

### 2016-09-01 Rmd/RenzeTest.Rmd
add ability to run with non-fixed w; revisted GMM vs kmeans test

@@
 require('reshape')
+require('jkmeans')
+
 ```
@@
 p<- 2
-batchN<- 1000
+batchN<- 10000
 
@@
 ```
+
+###2. Run j-sparse-GMM with different j
+
+Run 12-GMM, get average error
+
+```{r}
+
+jk12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini,fixW = F)
+
+error12 <- computeMError(jk12,n)
+
+mean(error12)
+```
+
+Run 22-GMM, get average error
+
+```{r}
+
+jk22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini,fixW = F)
+
+error22 <- computeMError(jk22,n)
+
+mean(error22)
+```
+
+Plot overlay error rate histograms:
+
+```{r}
+
+dataHist<- data.frame("J"=rep(c(1,2),each=batchN),
+                  "ErrorRate"= c(error12,error22))
+
+ggplot(dataHist, aes(ErrorRate, fill = as.factor(J))) + geom_histogram(alpha = 0.2,bins = 50,  position="identity")
+```
+
+#Conclusion:
+
+The better performance of the k-means over GMM in terms of clustering, reported by Renze, is more likely due to fixing w=1/k in k-means.
+
+On the other hand, we do see 1-sparse-GMM outperforms 2-sparse-GMM in clustering. This suggests focusing applying the sparsity idea in general mixture model would be a good direction.

### 2016-09-01 R/RcppExports.R

@@ -1,11 +1,11 @@
 # This file was generated by Rcpp::compileAttributes
 # Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
 
-jkmeansEM <- function(y, k, j, steps = 1000L, tol = 1E-8, useKmeansIni = TRUE, meansIni = 0L) {
-    .Call('jkmeans_jkmeansEM', PACKAGE = 'jkmeans', y, k, j, steps, tol, useKmeansIni, meansIni)
+jkmeansEM <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, useKmeansIni = TRUE, meansIni = 0L) {
+    .Call('jkmeans_jkmeansEM', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, useKmeansIni, meansIni)
 }
 
-jkmeansEMBatch <- function(y, k, j, steps = 1000L, tol = 1E-8, useKmeansIni = TRUE, meansIni = 0L) {
-    .Call('jkmeans_jkmeansEMBatch', PACKAGE = 'jkmeans', y, k, j, steps, tol, useKmeansIni, meansIni)
+jkmeansEMBatch <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, useKmeansIni = TRUE, meansIni = 0L) {
+    .Call('jkmeans_jkmeansEMBatch', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, useKmeansIni, meansIni)
 }
 

### 2016-09-01 src/RcppExports.cpp

@@ -7,8 +7,8 @@
 using namespace Rcpp;
 
 // jkmeansEM
-Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps, double tol, bool useKmeansIni, const arma::mat& meansIni);
-RcppExport SEXP jkmeans_jkmeansEM(SEXP ySEXP, SEXP kSEXP, SEXP jSEXP, SEXP stepsSEXP, SEXP tolSEXP, SEXP useKmeansIniSEXP, SEXP meansIniSEXP) {
+Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps, double tol, bool fixW, bool useKmeansIni, const arma::mat& meansIni);
+RcppExport SEXP jkmeans_jkmeansEM(SEXP ySEXP, SEXP kSEXP, SEXP jSEXP, SEXP stepsSEXP, SEXP tolSEXP, SEXP fixWSEXP, SEXP useKmeansIniSEXP, SEXP meansIniSEXP) {
 BEGIN_RCPP
     Rcpp::RObject __result;
     Rcpp::RNGScope __rngScope;
@@ -17,15 +17,16 @@ BEGIN_RCPP
     Rcpp::traits::input_parameter< int >::type j(jSEXP);
     Rcpp::traits::input_parameter< int >::type steps(stepsSEXP);
     Rcpp::traits::input_parameter< double >::type tol(tolSEXP);
+    Rcpp::traits::input_parameter< bool >::type fixW(fixWSEXP);
     Rcpp::traits::input_parameter< bool >::type useKmeansIni(useKmeansIniSEXP);
     Rcpp::traits::input_parameter< const arma::mat& >::type meansIni(meansIniSEXP);
-    __result = Rcpp::wrap(jkmeansEM(y, k, j, steps, tol, useKmeansIni, meansIni));
+    __result = Rcpp::wrap(jkmeansEM(y, k, j, steps, tol, fixW, useKmeansIni, meansIni));
     return __result;
 END_RCPP
 }
 // jkmeansEMBatch
-Rcpp::List jkmeansEMBatch(const arma::cube& y, int k, int j, int steps, double tol, bool useKmeansIni, const arma::mat& meansIni);
[…]

### 2016-09-01 src/jkmeans.cpp

@@ -13,15 +13,15 @@ using namespace arma;
 
 // [[Rcpp::export]]
 Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps = 1000,
-                     double tol = 1E-8, bool useKmeansIni = true,
-                     const arma::mat& meansIni = 0) {
+                     double tol = 1E-8, bool fixW = true,
+                     bool useKmeansIni = true, const arma::mat& meansIni = 0) {
   Mixture mix(y, k, j);
 
   if (j > k) {
     throw std::range_error("j needs be no bigger than k");
   }
 
-  mix.initialize(meansIni, useKmeansIni);
+  mix.initialize(meansIni, useKmeansIni, fixW);
 
   mix.runEM(steps, tol);
 
@@ -33,7 +33,8 @@ Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps = 1000,
 
 // [[Rcpp::export]]
 Rcpp::List jkmeansEMBatch(const arma::cube& y, int k, int j, int steps = 1000,
-                          double tol = 1E-8, bool useKmeansIni = true,
+                          double tol = 1E-8, bool fixW = true,
+                          bool useKmeansIni = true,
                           const arma::mat& meansIni = 0) {
   if (j > k) {
     throw std::range_error("j needs be no bigger than k");
@@ -53,7 +54,7 @@ Rcpp::List jkmeansEMBatch(const arma::cube& y, int k, int j, int steps = 1000,
   for (int i = 0; i < batchN; ++i) {
     mat localY = y.slice(i);
     Mixture mix(localY, k, j);
-    mix.initialize(meansIni, useKmeansIni);
+    mix.initialize(meansIni, useKmeansIni, fixW);
     mix.runEM(steps, tol);
 
     mu.slice(i) = mix.mu;

### 2016-09-01 src/mixture.hpp

@@ -16,6 +16,8 @@ class Mixture {
   vec gradient;
   mat hessian;
 
+  bool fixW;
+
   Mixture(mat _y, int _K, int _J) {
     y = _y;
     K = _K;
@@ -40,7 +42,7 @@ class Mixture {
     hessian = zeros<mat>(p * K, p * K);
   }
 
-  void initialize(mat meansInput, bool useKmeansIni) {
+  void initialize(mat meansInput, bool useKmeansIni, bool _fixW) {
     if (useKmeansIni) {
       mat means;
       bool status = kmeans(means, trans(y), K, random_subset, 10, false);
@@ -48,6 +50,8 @@ class Mixture {
     } else {
       mu = meansInput;
     }
+
+    fixW = _fixW;
   }
 
   double loglik(rowvec x, rowvec mu, double sigma2) {
@@ -130,7 +134,7 @@ class Mixture {
       // (total_weights * p + 0.1);
 
       // compute w
-      // w(k) = accu(weights) / (double)n;
+      if (!fixW) w(k) = accu(weights) / (double)n;
     }
 
     sigma2 = (diff2 / sum_total_weights);

### 2016-09-01 docs/latex/ver1.tex
sync

@@
 
-\subsection{JK-Means}
-
-With the above general framework, we focus on a special case $\{ u_{1,i},u_{2,i},\ldots u_{\kappa,i}\} \sim \text{ J out of K draws}$ and $\nu_k = 1/\kappa$.
-We refer this as JK-Means as when $J=1$, this reduces to standard K-means method.
+\subsection{Estimation}
 
@@
 \begin{algorithm}[H]
-\caption{EM algorithm for JK-Means}\label{alg:euclid}
+\caption{EM algorithm}\label{alg:euclid}
 \begin{algorithmic}[1]
@@
 \State Maximize over $\{ u_{i,k}\}_k$ by $ \underset{\{u_{i,k}\}_k} {\text{argmax }}\{u_{i,k}  v_{k}\delta_{\theta_{k}}(y_i)\}$ for all $i$; \Comment{\small pick top J out K based on (\ref{marginal_lik})}
-\State Compute $ \mathbb{E}(z_{i,k})= \frac{w_{k,i} \delta_{\theta_{k}}(y_i)}{\sum_k w_{k,i} \delta_{\theta_{k}}(y_i)}$ with $w_{k,i} = u_{i,k} v_{k}$; \Comment{E step  based on (\ref{conditional_lik})}
+\State Compute $ \mathbb{E}(z_{k,i})= \frac{w_{k,i} \delta_{\theta_{k}}(y_i)}{\sum_k w_{k,i} \delta_{\theta_{k}}(y_i)}$ with $w_{k,i} = u_{i,k} v_{k}$; \Comment{E step  based on (\ref{conditional_lik})}
 \State Set $\theta^{(s+1)}_{k}= \text{argmax} \sum_{k=1}^{\kappa} \mathbb{E}z_{k,i} \log \delta_{\theta_{k}}(y_i) $ for all $k$ \Comment{M step: obtain MLE based on (\ref{conditional_lik})}
-\EndWhile\label{euclidendwhile}
+\State Set $v^{(s+1)}_{k}=  \sum_{i}\mathbb{E}z_{k,i} /  \sum_{i}\sum_{k}\mathbb{E}z_{k,i}$ \Comment{M step: obtain MLE based on (\ref{conditional_lik})}\label{flex_w_step}
+\EndWhile
 % \State \textbf{return} $b$\Comment{The gcd is b}
@@
 
+\subsubsection{JK-Means}
+
+With the above general framework, there is a special case $\{ u_{1,i},u_{2,i},\ldots u_{\kappa,i}\} \sim \text{ J out of K draws}$ and $\nu_k = 1/\kappa$.
+We refer this as JK-Means as when $J=1$, this reduces to standard K-means method.
+The estimation procedure is the same as above except step \ref{flex_w_step} is skipped.
+
 \bibliographystyle{plainnat}

### 2016-09-06 Rmd/unequalClustersTestSmallN.Rmd
sync

--- 2016-09-01 Rmd/RenzeTest.Rmd
@@
 ###1. Generate Data
-Let's generate data from 2 clusters, each of size n
+Let's generate data from 2 clusters, with size n/2, n/2*3
 
@@
 mu<- matrix( 0,n*2,p)
-mu[1:n,]<- rep(c(0.7,0.3),each=n)
-mu[(n+1):(2*n),]<- rep(c(0.3,0.7),each=n)
+mu[1:(n/2),]<- rep(c(0.7,0.3),each=(n/2))
+mu[(n/2+1):(2*n),]<- rep(c(0.3,0.7),each=(n/2*3))
 
-sigma<-  sqrt(1.7661/n)
+sigma<-  sqrt(1.7661/30)
 
@@
     newOrder<- order(jk$mu[,1,b],decreasing = T)
-    trueM<- rep( c(0,1)[newOrder], each=n)
+    tM0<- c(0,1)[newOrder]
+    trueM<- c(rep(tM0[1],(n/2)),rep(tM0[2],(n/2*3)))
     error[b]<- sum (jk$M[,b] != trueM) /2/n
@@
   error
+}
+
+computeRMSE<- function(jk, mu0){
+  rmse<- numeric(batchN)
+  for(b in 1:batchN){
+    newOrder<- order(jk$mu[,1,b],decreasing = T)
+
+    muEst<-jk$mu[,,b][newOrder,]
+
+    rmse[b]<- sqrt( mean((muEst - mu0)^2))
+  }
+  rmse
 }
@@
 error12 <- computeMError(jk12,n)
+rmse12 <- computeRMSE(jk12, mu0)
 
 mean(error12)
+mean(rmse12)
 ```
@@
 error22 <- computeMError(jk22,n)
+rmse22 <- computeRMSE(jk22, mu0)
 
 mean(error22)
+mean(rmse22)
 ```
@@
 error22 <- computeMError(jk22,n)
+rmse22 <- computeRMSE(jk22, mu0)
 
 mean(error22)
+mean(rmse22)
+
 ```

### 2016-09-06 Rmd/convergenceRates.Rmd
sync

```
###Prerequisite
###Install the package if you haven't
setwd("~/git/")
require('devtools')
build('jkmeans')
install.packages("jkmeans_1.0.tar.gz", repos = NULL, type = "source")
```

```{r include=FALSE}
require('ggplot2')
require('reshape')
require('jkmeans')
```

###1. Generate Data
Let's generate data from 2 clusters, with size n/2, n/2*3

N( c(0.7), diag(1.7661/30))
N( c(0.3), diag(1.7661/30))

```{r}
n<- 100
K<- 2
p<- 1
batchN<- 1000

yBatch<- array(0,dim = c(n*K,p,batchN))

mu0<- c(c(0.7), c(0.3))
mu<- matrix( 0,n*2,p)
mu[1:(n/2),]<- rep(c(0.7),each=(n/2))
mu[(n/2+1):(2*n),]<- rep(c(0.3),each=(n/2*3))

sigma<-  sqrt(1.7661/30)

for(b in 1:batchN){
  yBatch[,,b]<- matrix( rnorm(n*p*2,mu,sigma), n*K, p)
}

mu_ini<- matrix(mu0,K,p)

```

Here is a view of the two clusters
```{r}

dataHist<- data.frame("Mu"=as.factor(mu),    "y"= yBatch[,,1])

ggplot(dataHist, aes(y, fill = as.factor(Mu))) + geom_histogram(alpha = 0.2,bins = 30,  position="identity")

```

functions to compute misclassification error and rmse of mean estimate:
NB: GMM can get stuck in saddle point at (w1=0, w2=1), so I removed the test result when it happened.
```{r}
computeMError<- function(jk, n,batchN){
  error<- numeric(batchN)
  for(b in 1:batchN){
    newOrder<- order(jk$mu[,1,b],decreasing = T)
    tM0<- c(0,1)[newOrder]
    trueM<- c(rep(tM0[1],(n/2)),rep(tM0[2],(n/2*3)))
    error[b]<- sum (jk$M[,b] != trueM) /2/n
    if( min(jk$w[,b])<0.1){
        error[b]<- NA
    }
  }
  error
}

computeRMSE<- function(jk, mu0, batchN){
  rmse<- numeric(batchN)
  for(b in 1:batchN){
    newOrder<- order(jk$mu[,1,b],decreasing = T)

    muEst<-as.matrix(jk$mu[,,b])[newOrder,]

    rmse[b]<- sqrt( mean((muEst - mu0)^2))
    if( min(jk$w[,b])<0.1)
      rmse[b]<- NA
  }
  rmse
}

```

###2. Test jk-means/ j-sparse-GMM with different n

```{r}

experiment<- function(n){

  K<- 2
  p<- 1
  batchN<- 1000

  yBatch<- array(0,dim = c(n*K,p,batchN))

  mu0<- c(c(0.7), c(0.3))
  mu<- matrix( 0,n*2,p)
  mu[1:(n/2),]<- rep(c(0.7),each=(n/2))
  mu[(n/2+1):(2*n),]<- rep(c(0.3),each=(n/2*3))

  sigma<-  sqrt(1.7661/30)

  for(b in 1:batchN){
    yBatch[,,b]<- matrix( rnorm(n*p*2,mu,sigma), n*K, p)
  }

  mu_ini<- matrix(mu0,K,p)

  jk12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)
  jk22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)
  jGMM12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F)
  jGMM22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F)

  error <- cbind( computeMError(jk12,n, batchN), computeMError(jk22,n,batchN), computeMError(jGMM12,n,batchN),computeMError(jGMM22,n,batchN))
  rmse <- cbind( computeRMSE(jk12,mu0,batchN), computeRMSE(jk22,mu0,batchN), computeRMSE(jGMM12,mu0,batchN),computeRMSE(jGMM22,mu0,batchN))

  list("MCE"=error, "RMSE"=rmse)

}

```

```{r}
#code to run on an increasing series of n

if(FALSE)
{
  nSeries<- c(seq(20,100,by = 20),150,200,300,400,500)

MCEmean<- numeric()
MCEq25<- numeric()
MCEq975<- numeric()
RMSEmean<- numeric()
RMSEq25<- numeric()
RMSEq975<- numeric()

for(n in nSeries){
  test<-experiment(n)

  MCEmean <- rbind(MCEmean,colMeans(test$MCE,na.rm = T))
  MCEq25 <- rbind(MCEq25, apply(test$MCE, MARGIN = 2, function(x){quantile(x,probs = 0.025,na.rm = T)}))
  MCEq975 <- rbind(MCEq975, apply(test$MCE, MARGIN = 2, function(x){quantile(x,probs = 0.975,na.rm = T)}))

  RMSEmean<- rbind(RMSEmean, colMeans(test$RMSE,na.rm = T))
  RMSEq25<- rbind(RMSEq25, apply(test$RMSE, MARGIN = 2, function(x){quantile(x,probs = 0.025,na.rm = T)}))
  RMSEq975<-  rbind(RMSEq975, apply(test$RMSE, MARGIN = 2, function(x){quantile(x,probs = 0.975,na.rm = T)}))

  print(n)
}

result<- list("MCEmean"=MCEmean,
     "MCEq25"=MCEq25,
     "MCEq975"=MCEq975,
     "RMSEmean"=RMSEmean,
     "RMSEq25"=RMSEq25,
     "RMSEq975"=RMSEq975)

save(result, file="resultRates.Rda")
}

```

```{r}

load("resultRates.Rda")

```

### 2016-09-06 Rmd/convergenceRates.R

@@ -1,0 +1,98 @@
+require('jkmeans')	
+
+computeMError<- function(jk, n,batchN){	
+  error<- numeric(batchN)	
+  for(b in 1:batchN){	
+    newOrder<- order(jk$mu[,1,b],decreasing = T)	
+    tM0<- c(0,1)[newOrder]	
+    trueM<- c(rep(tM0[1],(n/2)),rep(tM0[2],(n/2*3)))	
+    error[b]<- sum (jk$M[,b] != trueM) /2/n	
+    if( min(jk$w[,b])<0.1){	
+      error[b]<- NA	
+    }	
+  }	
+  error	
+}	
+
+computeRMSE<- function(jk, mu0, batchN){	
+  rmse<- numeric(batchN)	
+  for(b in 1:batchN){	
+    newOrder<- order(jk$mu[,1,b],decreasing = T)	
+    
+    muEst<-as.matrix(jk$mu[,,b])[newOrder,]	
+    
+    rmse[b]<- sqrt( mean((muEst - mu0)^2))	
+    if( min(jk$w[,b])<0.1)	
+      rmse[b]<- NA	
+  }	
+  rmse	
+}
+
+experiment<- function(n){	
+  
+  K<- 2	
+  p<- 1	
+  batchN<- 1000	
+  
+  yBatch<- array(0,dim = c(n*K,p,batchN))	
+  
+  mu0<- c(c(0.7), c(0.3))	
+  mu<- matrix( 0,n*2,p)	
+  mu[1:(n/2),]<- rep(c(0.7),each=(n/2))	
+  mu[(n/2+1):(2*n),]<- rep(c(0.3),each=(n/2*3))	
+  
+  sigma<-  sqrt(1.7661/30)	
+  
+  for(b in 1:batchN){	
+    yBatch[,,b]<- matrix( rnorm(n*p*2,mu,sigma), n*K, p)	
+  }	
+  
+  mu_ini<- matrix(mu0,K,p)	
+  
+  
+  jk12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)	
+  jk22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)	
[…]

### 2016-09-06 Rmd/convergenceRates.Rmd
add empirical estimates for the convergence rate

@@
 require('jkmeans')
+
+setwd("~/git/jkmeans/Rmd/")
 ```
@@
 
-###2. Test jk-means/ j-sparse-GMM with different n
+###2. Test jk-means/ j-sparse-GMM with different n's
 
@@
 #code to run on an increasing series of n
+#ran on cluster
+
+nSeries<- c(seq(20,500,by = 20))
 
@@
 {
-  nSeries<- c(seq(20,100,by = 20),150,200,300,400,500)
 
@@
 
+#Misclassification error
+plot without & with pointwise 95% confidence band
+
 ```{r}
@@
 
+plot1<- data.frame( "n" = rep(nSeries, 4),
+            "MC14Error"= c(result$MCEmean),
+            "MC14ErrorL"= c(result$MCEq25),
+            "MC14ErrorU"= c(result$MCEq975),
+            "Model"= rep(c("K-Means","W-constraint-GMM (22-Means)","1-sparse-GMM","GMM"),each= length(nSeries))
+            )
+
+p<- ggplot(data=plot1, aes())
+p<- p+  geom_line(aes(x= n, y=MC14Error, colour=Model)) 
+pE<- p+  geom_line(aes(x= n, y=MC14ErrorL, colour=Model),linetype=2) 
+pE<- pE+  geom_line(aes(x= n, y=MC14ErrorU, colour=Model),linetype=2)
+
+# p+ geom_errorbar( aes(x=n,ymax = MC14ErrorU, ymin=MC14ErrorL, colour=Model),width=0.2)
+p+ theme_bw()
+pE + theme_bw()
 ```
+
+#RMSE for the mean estimate
+plot without & with pointwise 95% confidence band
+
+```{r}
+
+plot1<- data.frame( "n" = rep(nSeries, 4),
+            "RMSE"= c(result$RMSEmean),
+            "RMSEL"= c(result$RMSEq25),
+            "RMSEU"= c(result$RMSEq975),
+            "Model"= rep(c("K-Means","W-constraint-GMM (22-Means)","1-sparse-GMM","GMM"),each= length(nSeries))
+            )
+
+p<- ggplot(data=plot1, aes())
+p<- p+  geom_line(aes(x= n, y=RMSE, colour=Model))
+
+pE<- p+  geom_line(aes(x= n, y=RMSEL, colour=Model),linetype=2) 
+pE<- pE+  geom_line(aes(x= n, y=RMSEU, colour=Model),linetype=2)
+
+p+ theme_bw()
+pE + theme_bw()
+```

### 2016-09-26 Rmd/convergenceRatesBalancedLabels.Rmd
add balanced label test

--- 2016-09-06 Rmd/convergenceRates.Rmd
@@
 ###1. Generate Data
-Let's generate data from 2 clusters, with size n/2, n/2*3
+Let's generate data from 2 clusters, with size n, n
 
@@
 mu<- matrix( 0,n*2,p)
-mu[1:(n/2),]<- rep(c(0.7),each=(n/2))
-mu[(n/2+1):(2*n),]<- rep(c(0.3),each=(n/2*3))
+mu[1:(n),]<- rep(c(0.7),each=(n))
+mu[(n+1):(2*n),]<- rep(c(0.3),each=(n))
 
@@
     tM0<- c(0,1)[newOrder]
-    trueM<- c(rep(tM0[1],(n/2)),rep(tM0[2],(n/2*3)))
+    trueM<- c(rep(tM0[1],(n)),rep(tM0[2],(n)))
     error[b]<- sum (jk$M[,b] != trueM) /2/n
@@
   mu<- matrix( 0,n*2,p)
-  mu[1:(n/2),]<- rep(c(0.7),each=(n/2))
-  mu[(n/2+1):(2*n),]<- rep(c(0.3),each=(n/2*3))
+  mu[1:(n),]<- rep(c(0.7),each=(n))
+  mu[(n+1):(2*n),]<- rep(c(0.3),each=(n))
 
@@
 
-  save(result, file="resultRates.Rda")
+  save(result, file="resultRatesBalanced.Rda")
 }
@@
 
-load("resultRates.Rda")
+load("resultRatesBalanced.Rda")
 

### 2016-09-26 Rmd/convergenceRatesBalancedLabels.R

@@ -1,0 +1,97 @@
+require('jkmeans')	
+
+computeMError<- function(jk, n,batchN){	
+  error<- numeric(batchN)	
+  for(b in 1:batchN){	
+    newOrder<- order(jk$mu[,1,b],decreasing = T)	
+    tM0<- c(0,1)[newOrder]	
+    trueM<- c(rep(tM0[1],(n)),rep(tM0[2],(n)))	
+    error[b]<- sum (jk$M[,b] != trueM) /2/n	
+    if( min(jk$w[,b])<0.1){	
+      error[b]<- NA	
+    }	
+  }	
+  error	
+}	
+
+computeRMSE<- function(jk, mu0, batchN){	
+  rmse<- numeric(batchN)	
+  for(b in 1:batchN){	
+    newOrder<- order(jk$mu[,1,b],decreasing = T)	
+    
+    muEst<-as.matrix(jk$mu[,,b])[newOrder,]	
+    
+    rmse[b]<- sqrt( mean((muEst - mu0)^2))	
+    if( min(jk$w[,b])<0.1)	
+      rmse[b]<- NA	
+  }	
+  rmse	
+}
+
+experiment<- function(n){	
+  
+  K<- 2	
+  p<- 1	
+  batchN<- 1000	
+  
+  yBatch<- array(0,dim = c(n*K,p,batchN))	
+  
+  mu0<- c(c(0.7), c(0.3))	
+  mu<- matrix( 0,n*2,p)	
+  mu[1:(n),]<- rep(c(0.7),each=(n))	
+  mu[(n+1):(2*n),]<- rep(c(0.3),each=(n))	
+  
+  sigma<-  sqrt(1.7661/30)	
+  
+  for(b in 1:batchN){	
+    yBatch[,,b]<- matrix( rnorm(n*p*2,mu,sigma), n*K, p)	
+  }	
+  
+  mu_ini<- matrix(mu0,K,p)	
+  
+  
+  jk12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)	
+  jk22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)	
[…]

### 2016-10-04 Rmd/convergenceRatesBalancedLabels.Rmd
update

@@
+---
+title: "jk test with balanced label"
+author: "Author"
+date: '`r Sys.Date()`'
+output:
+  md_document: 
+    variant: markdown_github
+---
+
 ```
@@
   for(b in 1:batchN){
-    newOrder<- order(jk$mu[,1,b],decreasing = T)
-    tM0<- c(0,1)[newOrder]
-    trueM<- c(rep(tM0[1],(n)),rep(tM0[2],(n)))
-    error[b]<- sum (jk$M[,b] != trueM) /2/n
+    tM0<- rank(jk$mu[,1,b])-1
+    trueM<- rep(tM0,each=n)
+    error[b]<- sum ( jk$M[,b] != trueM) /K/n
     if( min(jk$w[,b])<0.1){
@@
   for(b in 1:batchN){
-    newOrder<- order(jk$mu[,1,b],decreasing = T)
+    newOrder<- order(jk$mu[,1,b],decreasing = F)
 

### 2016-10-04 Rmd/convergenceRatesBigK.Rmd
update

--- 2016-10-04 Rmd/convergenceRatesBalancedLabels.Rmd
@@
 ---
-title: "jk test with balanced label"
+title: "jk test with big K, balanced label"
 author: "Author"
@@
 ###1. Generate Data
-Let's generate data from 2 clusters, with size n, n
-
-N( c(0.7), diag(1.7661/30))
-N( c(0.3), diag(1.7661/30))
+Let's generate data:
 
@@
 n<- 100
-K<- 2
+K<- 10
 p<- 1
@@
 
-mu0<- c(c(0.7), c(0.3))
-mu<- matrix( 0,n*2,p)
-mu[1:(n),]<- rep(c(0.7),each=(n))
-mu[(n+1):(2*n),]<- rep(c(0.3),each=(n))
-
-sigma<-  sqrt(1.7661/30)
+mu0<- c(1:K)
+mu<- matrix( 0,n*K,p)
+mu <- rep(mu0, each = n)
+sigma<-  0.5
 
 for(b in 1:batchN){
-  yBatch[,,b]<- matrix( rnorm(n*p*2,mu,sigma), n*K, p)
+  yBatch[,,b]<- matrix( rnorm(n*p*K,mu,sigma), n*K, p)
 }
@@
 
-  K<- 2
   p<- 1
@@
   yBatch<- array(0,dim = c(n*K,p,batchN))
+  K<- 10
 
-  mu0<- c(c(0.7), c(0.3))
-  mu<- matrix( 0,n*2,p)
-  mu[1:(n),]<- rep(c(0.7),each=(n))
-  mu[(n+1):(2*n),]<- rep(c(0.3),each=(n))
-
-  sigma<-  sqrt(1.7661/30)
+  mu0<- c(1:K)
+  mu<- matrix( 0,n*K,p)
+  mu <- rep(mu0, each = n)
+  sigma<-  0.5
 
   for(b in 1:batchN){
-    yBatch[,,b]<- matrix( rnorm(n*p*2,mu,sigma), n*K, p)
+    yBatch[,,b]<- matrix( rnorm(n*p*K,mu,sigma), n*K, p)
   }
@@
 
-  jk12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)
-  jk22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)
-  jGMM12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F)
-  jGMM22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F)
+  jk12 <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)
+  jk22 <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = K,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)
+  jGMM12 <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F)
+  jGMM22 <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = K,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F)
 

### 2016-10-04 Rmd/convergenceRatesBigK.r

@@ -1,0 +1,131 @@
+
+require('ggplot2')
+require('reshape')
+require('jkmeans')
+
+setwd("~/git/jkmeans/Rmd/")
+
+n<- 100
+K<- 10
+p<- 1
+batchN<- 1000
+
+yBatch<- array(0,dim = c(n*K,p,batchN))
+
+mu0<- c(1:K)
+mu<- matrix( 0,n*K,p)
+mu <- rep(mu0, each = n)
+sigma<-  0.5
+
+for(b in 1:batchN){
+  yBatch[,,b]<- matrix( rnorm(n*p*K,mu,sigma), n*K, p)
+}
+
+mu_ini<- matrix(mu0,K,p)
+
+
+
+dataHist<- data.frame("Mu"=as.factor(mu),    "y"= yBatch[,,1])
+
+ggplot(dataHist, aes(y, fill = as.factor(Mu))) + geom_histogram(alpha = 0.2,bins = 30,  position="identity")
+
+
+computeMError<- function(jk, n,batchN){
+  error<- numeric(batchN)
+  for(b in 1:batchN){
+    tM0<- rank(jk$mu[,1,b])-1
+    trueM<- rep(tM0,each=n)
+    error[b]<- sum ( jk$M[,b] != trueM) /K/n
+    if( min(jk$w[,b])<0.1){
+        error[b]<- NA
+    }
+  }
+  error
+}
+
+computeRMSE<- function(jk, mu0, batchN){
+  rmse<- numeric(batchN)
+  for(b in 1:batchN){
+    newOrder<- order(jk$mu[,1,b],decreasing = F)
+    
+    muEst<-as.matrix(jk$mu[,,b])[newOrder,]
+   
+    rmse[b]<- sqrt( mean((muEst - mu0)^2))
+    if( min(jk$w[,b])<0.1)
+      rmse[b]<- NA
+  }
+  rmse
+}
+
+experiment<- function(n){
+  
+  p<- 1
+  batchN<- 1000
+  
+  yBatch<- array(0,dim = c(n*K,p,batchN))
+  K<- 10
+  
+  mu0<- c(1:K)
+  mu<- matrix( 0,n*K,p)
+  mu <- rep(mu0, each = n)
+  sigma<-  0.5
+
+
+  for(b in 1:batchN){
+    yBatch[,,b]<- matrix( rnorm(n*p*K,mu,sigma), n*K, p)
+  }
+  
+  mu_ini<- matrix(mu0,K,p)
+  
+  
[…]

### 2016-10-04 Rmd/convergenceRatesImbalanced.Rmd
update labeling issue

--- 2016-10-04 Rmd/convergenceRatesBalancedLabels.Rmd
@@
 ---
-title: "jk test with balanced label"
+title: "jk test with unbalanced label"
 author: "Author"
@@
 ###1. Generate Data
-Let's generate data from 2 clusters, with size n, n
+Let's generate data from 2 clusters, with size n/2, n/2*3
 
+N( c(0.3), diag(1.7661/30))
 N( c(0.7), diag(1.7661/30))
-N( c(0.3), diag(1.7661/30))
 
@@
 
-mu0<- c(c(0.7), c(0.3))
+mu0<- c(c(0.3), c(0.7))
 mu<- matrix( 0,n*2,p)
-mu[1:(n),]<- rep(c(0.7),each=(n))
-mu[(n+1):(2*n),]<- rep(c(0.3),each=(n))
+mu[1:(n/2),]<- rep(c(0.3),each=(n/2))
+mu[(n/2+1):(2*n),]<- rep(c(0.7),each=(n/2*3))
 
@@
 ```{r}
+
 computeMError<- function(jk, n,batchN){
@@
     tM0<- rank(jk$mu[,1,b])-1
-    trueM<- rep(tM0,each=n)
+    trueM<- c(rep(tM0[1],(n/2)),rep(tM0[2],(n/2*3)))
     error[b]<- sum ( jk$M[,b] != trueM) /K/n
@@
 
-  mu0<- c(c(0.7), c(0.3))
+  mu0<- c(c(0.3), c(0.7))
   mu<- matrix( 0,n*2,p)
-  mu[1:(n),]<- rep(c(0.7),each=(n))
-  mu[(n+1):(2*n),]<- rep(c(0.3),each=(n))
+  mu[1:(n/2),]<- rep(c(0.3),each=(n/2))
+  mu[(n/2+1):(2*n),]<- rep(c(0.7),each=(n/2*3))
 
@@
 
-  save(result, file="resultRatesBalanced.Rda")
+  save(result, file="resultRatesImbalanced.Rda")
 }
@@
 
-load("resultRatesBalanced.Rda")
+load("resultRatesImbalanced.Rda")
 

### 2016-10-04 Rmd/convergenceRates.R

@@ -1,97 +1,0 @@
-require('jkmeans')	
-
-computeMError<- function(jk, n,batchN){	
-  error<- numeric(batchN)	
-  for(b in 1:batchN){	
-    newOrder<- order(jk$mu[,1,b],decreasing = T)	
-    tM0<- c(0,1)[newOrder]	
-    trueM<- c(rep(tM0[1],(n/2)),rep(tM0[2],(n/2*3)))	
-    error[b]<- sum (jk$M[,b] != trueM) /2/n	
-    if( min(jk$w[,b])<0.1){	
-      error[b]<- NA	
-    }	
-  }	
-  error	
-}	
-
-computeRMSE<- function(jk, mu0, batchN){	
-  rmse<- numeric(batchN)	
-  for(b in 1:batchN){	
-    newOrder<- order(jk$mu[,1,b],decreasing = T)	
-    
-    muEst<-as.matrix(jk$mu[,,b])[newOrder,]	
-    
-    rmse[b]<- sqrt( mean((muEst - mu0)^2))	
-    if( min(jk$w[,b])<0.1)	
-      rmse[b]<- NA	
-  }	
-  rmse	
-}
-
-experiment<- function(n){	
-  
-  K<- 2	
-  p<- 1	
-  batchN<- 1000	
-  
-  yBatch<- array(0,dim = c(n*K,p,batchN))	
-  
-  mu0<- c(c(0.7), c(0.3))	
-  mu<- matrix( 0,n*2,p)	
-  mu[1:(n/2),]<- rep(c(0.7),each=(n/2))	
-  mu[(n/2+1):(2*n),]<- rep(c(0.3),each=(n/2*3))	
-  
-  sigma<-  sqrt(1.7661/30)	
-  
-  for(b in 1:batchN){	
-    yBatch[,,b]<- matrix( rnorm(n*p*2,mu,sigma), n*K, p)	
-  }	
-  
-  mu_ini<- matrix(mu0,K,p)	
-  
-  
-  jk12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)	
-  jk22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)	
[…]

### 2016-10-04 Rmd/convergenceRatesBigK.r

@@ -127,5 +127,5 @@ nSeries<- c(seq(20,500,by = 20))
                 "RMSEq25"=RMSEq25,
                 "RMSEq975"=RMSEq975)
   
-  save(result, file="resultRatesBalanced.Rda")
+  save(result, file="resultRatesBigK.Rda")
 }

### 2016-10-04 Rmd/convergenceRatesImbalanced.R

@@ -14,8 +14,8 @@ yBatch<- array(0,dim = c(n*K,p,batchN))
 
 mu0<- c(c(0.3), c(0.7))
 mu<- matrix( 0,n*2,p)
-mu[1:(n),]<- rep(c(0.3),each=(n))
-mu[(n+1):(2*n),]<- rep(c(0.7),each=(n))
+mu[1:(n/2),]<- rep(c(0.3),each=(n/2))
+mu[(n/2+1):(2*n),]<- rep(c(0.7),each=(n/2*3))
 
 sigma<-  sqrt(1.7661/30)
 
@@ -31,13 +31,11 @@ dataHist<- data.frame("Mu"=as.factor(mu),    "y"= yBatch[,,1])
 ggplot(dataHist, aes(y, fill = as.factor(Mu))) + geom_histogram(alpha = 0.2,bins = 30,  position="identity")
 
 
-
-
 computeMError<- function(jk, n,batchN){
 error<- numeric(batchN)
 for(b in 1:batchN){
 tM0<- rank(jk$mu[,1,b])-1
-trueM<- rep(tM0,each=n)
+trueM<- c(rep(tM0[1],(n/2)),rep(tM0[2],(n/2*3)))
 error[b]<- sum ( jk$M[,b] != trueM) /K/n
 if( min(jk$w[,b])<0.1){
 error[b]<- NA
@@ -61,7 +59,6 @@ rmse
 }
 
 
-
 experiment<- function(n){
   
   K<- 2
@@ -72,8 +69,9 @@ experiment<- function(n){
   
   mu0<- c(c(0.3), c(0.7))
   mu<- matrix( 0,n*2,p)
-  mu[1:(n),]<- rep(c(0.3),each=(n))
-  mu[(n+1):(2*n),]<- rep(c(0.7),each=(n))
+  mu[1:(n/2),]<- rep(c(0.3),each=(n/2))
+  mu[(n/2+1):(2*n),]<- rep(c(0.7),each=(n/2*3))
+  
   
   sigma<-  sqrt(1.7661/30)
   
@@ -97,11 +95,9 @@ experiment<- function(n){
   
 }
 
-
-
-
 nSeries<- c(seq(20,500,by = 20))
 
+# if(FALSE)
 {
   
   MCEmean<- numeric()
@@ -134,5 +130,5 @@ nSeries<- c(seq(20,500,by = 20))
                 "RMSEq25"=RMSEq25,
                 "RMSEq975"=RMSEq975)
   
[…]

### 2016-10-06 Rmd/convergenceRatesBigK2.Rmd
added tests with K=10 J=1,2,3 with balanced labels

--- 2016-10-04 Rmd/convergenceRatesBigK.Rmd
@@
     error[b]<- sum ( jk$M[,b] != trueM) /K/n
-    if( min(jk$w[,b])<0.1){
+    if( min(jk$w[,b])<0.01){
         error[b]<- NA
@@
     rmse[b]<- sqrt( mean((muEst - mu0)^2))
-    if( min(jk$w[,b])<0.1)
+    if( min(jk$w[,b])<0.01)
       rmse[b]<- NA
@@
 
-  save(result, file="resultRatesBalanced.Rda")
+  save(result, file="resultRatesBigK2.Rda")
 }
@@
 
-load("resultRatesBalanced.Rda")
+load("resultRatesBigK2.Rda")
 

### 2016-10-09 Rmd/convergenceRatesImbalanced.Rmd
added flex J option

@@
 
-  error <- cbind( computeMError(jk12,n, batchN), computeMError(jk22,n,batchN), computeMError(jGMM12,n,batchN),computeMError(jGMM22,n,batchN))
-  rmse <- cbind( computeRMSE(jk12,mu0,batchN), computeRMSE(jk22,mu0,batchN), computeRMSE(jGMM12,mu0,batchN),computeRMSE(jGMM22,mu0,batchN))
+  jGMMFlex <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F,flexJ = T,zetaTrunc = 0.05)
+
+  error <- cbind( computeMError(jk12,n, batchN), computeMError(jk22,n,batchN), computeMError(jGMM12,n,batchN),computeMError(jGMM22,n,batchN), computeMError(jGMMFlex,n,batchN)  )
+  rmse <- cbind( computeRMSE(jk12,mu0,batchN), computeRMSE(jk22,mu0,batchN), computeRMSE(jGMM12,mu0,batchN),computeRMSE(jGMM22,mu0,batchN), computeRMSE(jGMMFlex,mu0,batchN))
 
@@
 
-plot1<- data.frame( "n" = rep(nSeries, 4),
+plot1<- data.frame( "n" = rep(nSeries, 5),
             "MC14Error"= c(result$MCEmean),
@@
             "MC14ErrorU"= c(result$MCEq975),
-            "Model"= rep(c("K-Means","W-constraint-GMM (22-Means)","1-sparse-GMM","GMM"),each= length(nSeries))
+            "Model"= rep(c("K-Means","W-constraint-GMM (22-Means)","1-sparse-GMM","GMM","flexJGMM"),each= length(nSeries))
             )
@@
 
-plot1<- data.frame( "n" = rep(nSeries, 4),
+plot1<- data.frame( "n" = rep(nSeries, 5),
             "RMSE"= c(result$RMSEmean),
@@
             "RMSEU"= c(result$RMSEq975),
-            "Model"= rep(c("K-Means","W-constraint-GMM (22-Means)","1-sparse-GMM","GMM"),each= length(nSeries))
+            "Model"= rep(c("K-Means","W-constraint-GMM (22-Means)","1-sparse-GMM","GMM","flexJGMM"),each= length(nSeries))
             )

### 2016-10-09 R/RcppExports.R

@@ -1,11 +1,11 @@
 # This file was generated by Rcpp::compileAttributes
 # Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
 
-jkmeansEM <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, useKmeansIni = TRUE, meansIni = 0L) {
-    .Call('jkmeans_jkmeansEM', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, useKmeansIni, meansIni)
+jkmeansEM <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, flexJ = FALSE, zetaTrunc = 0.01, useKmeansIni = TRUE, meansIni = 0L) {
+    .Call('jkmeans_jkmeansEM', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni)
 }
 
-jkmeansEMBatch <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, useKmeansIni = TRUE, meansIni = 0L) {
-    .Call('jkmeans_jkmeansEMBatch', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, useKmeansIni, meansIni)
+jkmeansEMBatch <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, flexJ = FALSE, zetaTrunc = 0.01, useKmeansIni = TRUE, meansIni = 0L) {
+    .Call('jkmeans_jkmeansEMBatch', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni)
 }
 

### 2016-10-09 Rmd/convergenceRatesImbalanced.R

@@ -20,10 +20,12 @@ mu[(n/2+1):(2*n),]<- rep(c(0.7),each=(n/2*3))
 sigma<-  sqrt(1.7661/30)
 
 for(b in 1:batchN){
-yBatch[,,b]<- matrix( rnorm(n*p*2,mu,sigma), n*K, p)
+  yBatch[,,b]<- matrix( rnorm(n*p*2,mu,sigma), n*K, p)
 }
 
 mu_ini<- matrix(mu0,K,p)
+
+
 
 
 dataHist<- data.frame("Mu"=as.factor(mu),    "y"= yBatch[,,1])
@@ -31,32 +33,35 @@ dataHist<- data.frame("Mu"=as.factor(mu),    "y"= yBatch[,,1])
 ggplot(dataHist, aes(y, fill = as.factor(Mu))) + geom_histogram(alpha = 0.2,bins = 30,  position="identity")
 
 
+
 computeMError<- function(jk, n,batchN){
-error<- numeric(batchN)
-for(b in 1:batchN){
-tM0<- rank(jk$mu[,1,b])-1
-trueM<- c(rep(tM0[1],(n/2)),rep(tM0[2],(n/2*3)))
-error[b]<- sum ( jk$M[,b] != trueM) /K/n
-if( min(jk$w[,b])<0.1){
-error[b]<- NA
-}
-}
-error
+  error<- numeric(batchN)
+  for(b in 1:batchN){
+    tM0<- rank(jk$mu[,1,b])-1
+    trueM<- c(rep(tM0[1],(n/2)),rep(tM0[2],(n/2*3)))
+    error[b]<- sum ( jk$M[,b] != trueM) /K/n
+    if( min(jk$w[,b])<0.1){
+        error[b]<- NA
+    }
+  }
+  error
 }
 
 computeRMSE<- function(jk, mu0, batchN){
-rmse<- numeric(batchN)
-for(b in 1:batchN){
-newOrder<- order(jk$mu[,1,b],decreasing = F)
-
-muEst<-as.matrix(jk$mu[,,b])[newOrder,]
-
-rmse[b]<- sqrt( mean((muEst - mu0)^2))
-if( min(jk$w[,b])<0.1)
-rmse[b]<- NA
+  rmse<- numeric(batchN)
+  for(b in 1:batchN){
+    newOrder<- order(jk$mu[,1,b],decreasing = F)
+    
+    muEst<-as.matrix(jk$mu[,,b])[newOrder,]
+   
[…]

### 2016-10-09 src/RcppExports.cpp

@@ -7,8 +7,8 @@
 using namespace Rcpp;
 
 // jkmeansEM
-Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps, double tol, bool fixW, bool useKmeansIni, const arma::mat& meansIni);
-RcppExport SEXP jkmeans_jkmeansEM(SEXP ySEXP, SEXP kSEXP, SEXP jSEXP, SEXP stepsSEXP, SEXP tolSEXP, SEXP fixWSEXP, SEXP useKmeansIniSEXP, SEXP meansIniSEXP) {
+Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps, double tol, bool fixW, bool flexJ, double zetaTrunc, bool useKmeansIni, const arma::mat& meansIni);
+RcppExport SEXP jkmeans_jkmeansEM(SEXP ySEXP, SEXP kSEXP, SEXP jSEXP, SEXP stepsSEXP, SEXP tolSEXP, SEXP fixWSEXP, SEXP flexJSEXP, SEXP zetaTruncSEXP, SEXP useKmeansIniSEXP, SEXP meansIniSEXP) {
 BEGIN_RCPP
     Rcpp::RObject __result;
     Rcpp::RNGScope __rngScope;
@@ -18,15 +18,17 @@ BEGIN_RCPP
     Rcpp::traits::input_parameter< int >::type steps(stepsSEXP);
     Rcpp::traits::input_parameter< double >::type tol(tolSEXP);
     Rcpp::traits::input_parameter< bool >::type fixW(fixWSEXP);
+    Rcpp::traits::input_parameter< bool >::type flexJ(flexJSEXP);
+    Rcpp::traits::input_parameter< double >::type zetaTrunc(zetaTruncSEXP);
     Rcpp::traits::input_parameter< bool >::type useKmeansIni(useKmeansIniSEXP);
     Rcpp::traits::input_parameter< const arma::mat& >::type meansIni(meansIniSEXP);
-    __result = Rcpp::wrap(jkmeansEM(y, k, j, steps, tol, fixW, useKmeansIni, meansIni));
+    __result = Rcpp::wrap(jkmeansEM(y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni));
     return __result;
 END_RCPP
 }
 // jkmeansEMBatch
[…]

### 2016-10-09 src/jkmeans.cpp

@@ -13,15 +13,16 @@ using namespace arma;
 
 // [[Rcpp::export]]
 Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps = 1000,
-                     double tol = 1E-8, bool fixW = true,
-                     bool useKmeansIni = true, const arma::mat& meansIni = 0) {
+                     double tol = 1E-8, bool fixW = true, bool flexJ = false,
+                     double zetaTrunc = 0.01, bool useKmeansIni = true,
+                     const arma::mat& meansIni = 0) {
   Mixture mix(y, k, j);
 
   if (j > k) {
     throw std::range_error("j needs be no bigger than k");
   }
 
-  mix.initialize(meansIni, useKmeansIni, fixW);
+  mix.initialize(meansIni, useKmeansIni, fixW, flexJ, zetaTrunc);
 
   mix.runEM(steps, tol);
 
@@ -34,6 +35,7 @@ Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps = 1000,
 // [[Rcpp::export]]
 Rcpp::List jkmeansEMBatch(const arma::cube& y, int k, int j, int steps = 1000,
                           double tol = 1E-8, bool fixW = true,
+                          bool flexJ = false, double zetaTrunc = 0.01,
                           bool useKmeansIni = true,
                           const arma::mat& meansIni = 0) {
   if (j > k) {
@@ -54,7 +56,7 @@ Rcpp::List jkmeansEMBatch(const arma::cube& y, int k, int j, int steps = 1000,
   for (int i = 0; i < batchN; ++i) {
     mat localY = y.slice(i);
     Mixture mix(localY, k, j);
-    mix.initialize(meansIni, useKmeansIni, fixW);
+    mix.initialize(meansIni, useKmeansIni, fixW, flexJ, zetaTrunc);
     mix.runEM(steps, tol);
 
     mu.slice(i) = mix.mu;

### 2016-10-09 src/mixture.hpp

@@ -17,6 +17,8 @@ class Mixture {
   mat hessian;
 
   bool fixW;
+  bool flexJ;
+  double zetaTrunc;
 
   Mixture(mat _y, int _K, int _J) {
     y = _y;
@@ -42,7 +44,8 @@ class Mixture {
     hessian = zeros<mat>(p * K, p * K);
   }
 
-  void initialize(mat meansInput, bool useKmeansIni, bool _fixW) {
+  void initialize(mat meansInput, bool useKmeansIni, bool _fixW, bool _flexJ,
+                  double _zetaTrunc) {
     if (useKmeansIni) {
       mat means;
       bool status = kmeans(means, trans(y), K, random_subset, 10, false);
@@ -52,6 +55,8 @@ class Mixture {
     }
 
     fixW = _fixW;
+    flexJ = _flexJ;
+    zetaTrunc = _zetaTrunc;
   }
 
   double loglik(rowvec x, rowvec mu, double sigma2) {
@@ -93,18 +98,20 @@ class Mixture {
 
       local_zeta -= local_zeta.max();
 
-      // if(local_zeta.has_nan()){
-      //   cout<< local_zeta<<endl;
-      //   cout<< zeta<<endl;
-      // }
+      if (!flexJ) {
+        uvec indices = sort_index(local_zeta, "descend");
 
-      uvec indices = sort_index(local_zeta, "descend");
+        int trunc_idx = indices(J - 1);
 
-      int trunc_idx = indices(J - 1);
+        local_zeta = exp(local_zeta);
 
-      local_zeta = exp(local_zeta);
-
-      local_zeta(find(local_zeta < local_zeta(trunc_idx))).fill(0);
+        local_zeta(find(local_zeta < local_zeta(trunc_idx))).fill(0);
+      }
+      if (flexJ) {
+        local_zeta = exp(local_zeta);
+        local_zeta /= accu(local_zeta);
+        local_zeta(find(local_zeta < zetaTrunc)).fill(0);
+      }
 
       zeta.row(i) = local_zeta / accu(local_zeta);
     }

### 2016-10-10 docs/ver1.tex
sync

--- 2016-09-01 docs/latex/ver1.tex
@@
 \title
-    {{Sparse Mixture Model (with jk-means a special case)}}
+    {?? Mixture Model}
 
@@
 
-\section{Sparse Mixture Model}
+\section{Introduction}
+
+hard clustering:
+several overly strong assumptions in k-means
+
+soft clustering:
+large variance induced by way too many non-zeros in the probability simplex.
+
+variance-bias tradeoff?
+reducing variability on the weight
+
+\section{?? Mixture Model}
 
@@
 
-On a separable metrics space $\{\mathcal{X},\mathcal{F}\}$, we assume each data $y_i$ has the following probabilty density as the likelihood:
+On a separable metrics space $\{\mathcal{X},\mathcal{F}\}$, we assume each data $y_i$ has the following probability density as the likelihood:
 
@@
 \begin{aligned}
-& [ y_i ] = \sum_{k=1}^{\kappa} w_{k,i} \delta_{\theta_{k}}(y_i) \\
+& \pi(y_i) \sim G(\theta_i) \\
+& \pi(\theta_i) = \sum_{k=1}^{\kappa} w_{k,i} \delta_{\theta_{k}}(\theta_i) \\
 & w_{k,i} = \frac{ u_{k,i} v_{k} } { \sum_{k=1}^{\kappa}  u_{k,i} v_{k} }\\
@@
 Due to the $0$'s $u_{k,i}$ introduces, this induces sparsity in the membership probabilty $w_{k,i}$.
-Therefore, we refer this as the sparse mixture model.
 
@@
 
-\subsubsection{JK-Means}
+\section{Theory}
 
-With the above general framework, there is a special case $\{ u_{1,i},u_{2,i},\ldots u_{\kappa,i}\} \sim \text{ J out of K draws}$ and $\nu_k = 1/\kappa$.
-We refer this as JK-Means as when $J=1$, this reduces to standard K-means method.
-The estimation procedure is the same as above except step \ref{flex_w_step} is skipped.
+variance reduction in the weight parameter
+
+\section{Simulation}
+
+large K with small J in 1d
+
+high p setting
+
+\section{Application}
+\section{Discussion}
 

### 2016-10-18 Rmd/convergenceRatesImbalancedSmallSigma.Rmd
update large and small overlap
bug fixed; moved Rmd out of R package

--- 2016-10-04 Rmd/convergenceRatesImbalanced.Rmd
@@
 
-sigma<-  sqrt(1.7661/30)
+sigma<-  sqrt(1.7661/10000)
 
@@
 
-  jk12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)
-  jk22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)
-  jGMM12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F)
-  jGMM22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F)
+  jk12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T,sigma2_ini = sigma^2)
+  jk22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T,sigma2_ini = sigma^2)
+  jGMM12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F,sigma2_ini = sigma^2)
+  jGMM22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F,sigma2_ini = sigma^2)
 
@@
 
-if(FALSE)
+# if(FALSE)
 {
@@
 
-  save(result, file="resultRatesImbalanced.Rda")
+  save(result, file="resultRatesImbalancedSmallSigma.Rda")
 }
@@
 
-load("resultRatesImbalanced.Rda")
+load("resultRatesImbalancedSmallSigma.Rda")
 
@@
 pE + theme_bw()
+
 ```

### 2016-10-18 R/RcppExports.R

@@ -1,11 +1,0 @@
-# This file was generated by Rcpp::compileAttributes
-# Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
-
-jkmeansEM <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, flexJ = FALSE, zetaTrunc = 0.01, useKmeansIni = TRUE, meansIni = 0L) {
-    .Call('jkmeans_jkmeansEM', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni)
-}
-
-jkmeansEMBatch <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, flexJ = FALSE, zetaTrunc = 0.01, useKmeansIni = TRUE, meansIni = 0L) {
-    .Call('jkmeans_jkmeansEMBatch', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni)
-}
-

### 2016-10-18 jkmeans/R/RcppExports.R

@@ -1,0 +1,11 @@
+# This file was generated by Rcpp::compileAttributes
+# Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
+
+jkmeansEM <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, flexJ = FALSE, zetaTrunc = 0.01, useKmeansIni = TRUE, meansIni = 0L, sigma2_ini = 0.1) {
+    .Call('jkmeans_jkmeansEM', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini)
+}
+
+jkmeansEMBatch <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, flexJ = FALSE, zetaTrunc = 0.01, useKmeansIni = TRUE, meansIni = 0L, sigma2_ini = 0.1) {
+    .Call('jkmeans_jkmeansEMBatch', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini)
+}
+

### 2016-10-18 jkmeans/src/RcppExports.cpp

@@ -7,8 +7,8 @@
 using namespace Rcpp;
 
 // jkmeansEM
-Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps, double tol, bool fixW, bool flexJ, double zetaTrunc, bool useKmeansIni, const arma::mat& meansIni);
-RcppExport SEXP jkmeans_jkmeansEM(SEXP ySEXP, SEXP kSEXP, SEXP jSEXP, SEXP stepsSEXP, SEXP tolSEXP, SEXP fixWSEXP, SEXP flexJSEXP, SEXP zetaTruncSEXP, SEXP useKmeansIniSEXP, SEXP meansIniSEXP) {
+Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps, double tol, bool fixW, bool flexJ, double zetaTrunc, bool useKmeansIni, const arma::mat& meansIni, double sigma2_ini);
+RcppExport SEXP jkmeans_jkmeansEM(SEXP ySEXP, SEXP kSEXP, SEXP jSEXP, SEXP stepsSEXP, SEXP tolSEXP, SEXP fixWSEXP, SEXP flexJSEXP, SEXP zetaTruncSEXP, SEXP useKmeansIniSEXP, SEXP meansIniSEXP, SEXP sigma2_iniSEXP) {
 BEGIN_RCPP
     Rcpp::RObject __result;
     Rcpp::RNGScope __rngScope;
@@ -22,13 +22,14 @@ BEGIN_RCPP
     Rcpp::traits::input_parameter< double >::type zetaTrunc(zetaTruncSEXP);
     Rcpp::traits::input_parameter< bool >::type useKmeansIni(useKmeansIniSEXP);
     Rcpp::traits::input_parameter< const arma::mat& >::type meansIni(meansIniSEXP);
-    __result = Rcpp::wrap(jkmeansEM(y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni));
+    Rcpp::traits::input_parameter< double >::type sigma2_ini(sigma2_iniSEXP);
+    __result = Rcpp::wrap(jkmeansEM(y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini));
     return __result;
 END_RCPP
 }
 // jkmeansEMBatch
[…]

### 2016-10-18 jkmeans/src/jkmeans.cpp

@@ -15,14 +15,14 @@ using namespace arma;
 Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps = 1000,
                      double tol = 1E-8, bool fixW = true, bool flexJ = false,
                      double zetaTrunc = 0.01, bool useKmeansIni = true,
-                     const arma::mat& meansIni = 0) {
+                     const arma::mat& meansIni = 0, double sigma2_ini = 0.1) {
   Mixture mix(y, k, j);
 
   if (j > k) {
     throw std::range_error("j needs be no bigger than k");
   }
 
-  mix.initialize(meansIni, useKmeansIni, fixW, flexJ, zetaTrunc);
+  mix.initialize(meansIni, useKmeansIni, fixW, flexJ, zetaTrunc, sigma2_ini);
 
   mix.runEM(steps, tol);
 
@@ -37,7 +37,8 @@ Rcpp::List jkmeansEMBatch(const arma::cube& y, int k, int j, int steps = 1000,
                           double tol = 1E-8, bool fixW = true,
                           bool flexJ = false, double zetaTrunc = 0.01,
                           bool useKmeansIni = true,
-                          const arma::mat& meansIni = 0) {
+                          const arma::mat& meansIni = 0,
+                          double sigma2_ini = 0.1) {
   if (j > k) {
     throw std::range_error("j needs be no bigger than k");
   }
@@ -56,7 +57,7 @@ Rcpp::List jkmeansEMBatch(const arma::cube& y, int k, int j, int steps = 1000,
   for (int i = 0; i < batchN; ++i) {
     mat localY = y.slice(i);
     Mixture mix(localY, k, j);
-    mix.initialize(meansIni, useKmeansIni, fixW, flexJ, zetaTrunc);
+    mix.initialize(meansIni, useKmeansIni, fixW, flexJ, zetaTrunc, sigma2_ini);
     mix.runEM(steps, tol);
 
     mu.slice(i) = mix.mu;
@@ -70,20 +71,3 @@ Rcpp::List jkmeansEMBatch(const arma::cube& y, int k, int j, int steps = 1000,
                             Rcpp::Named("sigma2") = sigma2,
                             Rcpp::Named("zeta") = zeta, Rcpp::Named("M") = M);
 }
-
-// // [[Rcpp::export]]
-// Rcpp::List jkmeansQNEM(const arma::mat& y, int k, int j, int steps = 1000) {
-//   Mixture mix(y, k, j);
-
-//     if (j > k) {
[…]

### 2016-10-18 jkmeans/src/mixture.hpp

@@ -33,11 +33,6 @@ class Mixture {
     mu = ones<mat>(K, p);
     mu.randn();
 
-    mat means;
-    bool status = kmeans(means, trans(y), K, random_subset, 10, false);
-
-    mu = means.t();
-
     zeta = ones<mat>(n, K);
 
     gradient = zeros(p * K);
@@ -45,7 +40,9 @@ class Mixture {
   }
 
   void initialize(mat meansInput, bool useKmeansIni, bool _fixW, bool _flexJ,
-                  double _zetaTrunc) {
+                  double _zetaTrunc, double sigma2_ini) {
+    sigma2 = sigma2_ini;
+
     if (useKmeansIni) {
       mat means;
       bool status = kmeans(means, trans(y), K, random_subset, 10, false);

### 2016-10-18 jkmeans/test/build.r

@@ -1,6 +1,6 @@
 library(devtools)
 
-setwd("~/git/")
+setwd("~/git/jkmeans")
 build('jkmeans')
 install.packages("jkmeans_1.0.tar.gz", repos = NULL, type = "source")
 

### 2016-10-18 Rmd/render.r
sync

@@ -1,4 +1,0 @@
-require("rmarkdown")
-rmarkdown::render("1DimTest.Rmd")
-
-

### 2016-10-18 Rmd/render.sh

@@ -1,0 +1,5 @@
+#/bin/bash
+
+R -e "rmarkdown::render('$1')"
+
+

### 2016-10-19 Rmd/convergenceRatesBalancedBigK.Rmd
rerun tests after fixing initialization

--- 2016-10-06 Rmd/convergenceRatesBigK2.Rmd
@@
 n<- 100
-K<- 10
+K<- 5
 p<- 1
@@
   yBatch<- array(0,dim = c(n*K,p,batchN))
-  K<- 10
 
@@
   mu <- rep(mu0, each = n)
-  sigma<-  0.5
 
@@
 
-  jk12 <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)
-  jk22 <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = K,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)
-  jGMM12 <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F)
-  jGMM22 <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = K,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F)
+  jk1 <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T,sigma2_ini = sigma^2)
+  jk2 <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T,sigma2_ini = sigma^2)
+  jk3 <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = 3,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T,sigma2_ini = sigma^2)
+  jk4 <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = 4,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T,sigma2_ini = sigma^2)
+  jkK <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = K,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T,sigma2_ini = sigma^2)
 
-  error <- cbind( computeMError(jk12,n, batchN), computeMError(jk22,n,batchN), computeMError(jGMM12,n,batchN),computeMError(jGMM22,n,batchN))
-  rmse <- cbind( computeRMSE(jk12,mu0,batchN), computeRMSE(jk22,mu0,batchN), computeRMSE(jGMM12,mu0,batchN),computeRMSE(jGMM22,mu0,batchN))
+  jGMM1 <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F,sigma2_ini = sigma^2)
+  jGMM2 <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F,sigma2_ini = sigma^3)
+  jGMM3 <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = 3,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F,sigma2_ini = sigma^2)
+  jGMM4 <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = 4,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F,sigma2_ini = sigma^2)
+  jGMMK <- jkmeans::jkmeansEMBatch(yBatch, k=K, j = K,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F,sigma2_ini = sigma^2)
+
+  runRes <- list(jk1,jk2,jk3,jk4,jkK, jGMM1, jGMM2, jGMM3, jGMM4, jGMMK)
+
+  error<- sapply(runRes,function(x){computeMError(x,n,batchN)})
+  rmse<- sapply(runRes,function(x){computeRMSE(x,mu0,batchN)})
 
@@
 
-  save(result, file="resultRatesBigK2.Rda")
+  save(result, file="resultRatesBigK.Rda")
 }
@@
 
-```{r}
+```{r echo=FALSE}
 
-load("resultRatesBigK2.Rda")
+load("resultRatesBigK.Rda")
 
-plot1<- data.frame( "n" = rep(nSeries, 4),
-            "MC14Error"= c(result$MCEmean),
-            "MC14ErrorL"= c(result$MCEq25),
-            "MC14ErrorU"= c(result$MCEq975),
-            "Model"= rep(c("K-Means","W-constraint-GMM (22-Means)","1-sparse-GMM","GMM"),each= length(nSeries))
+plot1<- data.frame( "n" = rep(nSeries, 5),
+                    "MC14Error"= c(result$MCEmean[,1:5]),
+                    "MC14ErrorL"= c(result$MCEq25[,1:5]),
+                    "MC14ErrorU"= c(result$MCEq975[,1:5]),
+                    "Model"= rep(c("w fixed, J=1","w fixed, J=2","w fixed, J=3","w fixed, J=4","w fixed, J=K"
+                    ),each= length(nSeries))
 )
@@
 p+ theme_bw()
-pE + theme_bw()
+# pE + theme_bw()
+```
+
+```{r echo=FALSE}
+
+plot1<- data.frame( "n" = rep(nSeries, 5),
+                    "MC14Error"= c(result$MCEmean[,6:10]),
+                    "MC14ErrorL"= c(result$MCEq25[,6:10]),
+                    "MC14ErrorU"= c(result$MCEq975[,6:10]),
+                    "Model"= rep(c("w flex, J=1","w flex, J=2","w flex, J=3","w flex, J=4","w flex, J=K"
+                    ),each= length(nSeries))
+)
+
+p<- ggplot(data=plot1, aes())
+p<- p+  geom_line(aes(x= n, y=MC14Error, colour=Model)) 
+pE<- p+  geom_line(aes(x= n, y=MC14ErrorL, colour=Model),linetype=2) 
+pE<- pE+  geom_line(aes(x= n, y=MC14ErrorU, colour=Model),linetype=2)
+
+# p+ geom_errorbar( aes(x=n,ymax = MC14ErrorU, ymin=MC14ErrorL, colour=Model),width=0.2)
+p+ theme_bw()
+# pE + theme_bw()
 ```
@@
 
-```{r}
+```{r echo=FALSE}
 
-plot1<- data.frame( "n" = rep(nSeries, 4),
-            "RMSE"= c(result$RMSEmean),
-            "RMSEL"= c(result$RMSEq25),
-            "RMSEU"= c(result$RMSEq975),
-            "Model"= rep(c("K-Means","W-constraint-GMM (22-Means)","1-sparse-GMM","GMM"),each= length(nSeries))
-            )
+plot1<- data.frame( "n" = rep(nSeries, 5),
+                    "RMSE"= c(result$RMSEmean[,1:5]),
+                    "RMSEL"= c(result$RMSEq25[,1:5]),
+                    "RMSEU"= c(result$RMSEq975[,1:5]),
+                    "Model"= rep(c("w fixed, J=1","w fixed, J=2","w fixed, J=3","w fixed, J=4","w fixed, J=K"
+
+                    ),each= length(nSeries))            )
 
@@
 p+ theme_bw()
-pE + theme_bw()
+# pE + theme_bw()
 ```
+
+```{r echo=FALSE}
+
+plot1<- data.frame( "n" = rep(nSeries, 5),
+                    "RMSE"= c(result$RMSEmean[,6:10]),
+                    "RMSEL"= c(result$RMSEq25[,6:10]),
+                    "RMSEU"= c(result$RMSEq975[,6:10]),
+                    "Model"= rep(c(
+                                   "w flex, J=1","w flex, J=2","w flex, J=3","w flex, J=4","w flex, J=K"
+                    ),each= length(nSeries))            )
+
+p<- ggplot(data=plot1, aes())
+p<- p+  geom_line(aes(x= n, y=RMSE, colour=Model))
+
+pE<- p+  geom_line(aes(x= n, y=RMSEL, colour=Model),linetype=2) 
+pE<- pE+  geom_line(aes(x= n, y=RMSEU, colour=Model),linetype=2)
+
+p+ theme_bw()
+# pE + theme_bw()
+```

### 2016-10-19 Rmd/convergenceRatesImbalanced.Rmd
rerun tests after fixing initialization

@@
 
-  sigma<-  sqrt(1.7661/30)
-
   for(b in 1:batchN){
@@
 
-  jk12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)
-  jk22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T)
-  jGMM12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F)
-  jGMM22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F)
+  jk12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T,sigma2_ini = sigma^2)
+  jk22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = T,sigma2_ini = sigma^2)
+  jGMM12 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F,sigma2_ini = sigma^2)
+  jGMM22 <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 2,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F,sigma2_ini = sigma^2)
 
-  jGMMFlex <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F,flexJ = T,zetaTrunc = 0.05)
+  jGMMFlex <- jkmeans::jkmeansEMBatch(yBatch, k=2, j = 1,  1000,tol = 1E-10,useKmeansIni = F, meansIni = mu_ini, fixW = F,flexJ = T,zetaTrunc = 0.05,sigma2_ini = sigma^2)
 

### 2016-10-19 Rmd/convergenceRatesImbalancedLargeSigma.Rmd
update large and small overlap
rerun tests after fixing initialization

--- 2016-10-04 Rmd/convergenceRatesImbalanced.Rmd
@@
 
-sigma<-  sqrt(1.7661/30)
+sigma<-  sqrt(1.7661/10)
 
@@
 
-  save(result, file="resultRatesImbalanced.Rda")
+  save(result, file="resultRatesImbalancedLargeSigma.Rda")
 }
@@
 
-load("resultRatesImbalanced.Rda")
+load("resultRatesImbalancedLargeSigma.Rda")
 

### 2016-10-19 Rmd/convergenceRatesBalancedLabels.R

@@ -1,138 +1,0 @@
-
-require('ggplot2')
-require('reshape')
-require('jkmeans')
-
-setwd("~/git/jkmeans/Rmd/")
-
-n<- 100
-K<- 2
-p<- 1
-batchN<- 1000
-
-yBatch<- array(0,dim = c(n*K,p,batchN))
-
-mu0<- c(c(0.3), c(0.7))
-mu<- matrix( 0,n*2,p)
-mu[1:(n),]<- rep(c(0.3),each=(n))
-mu[(n+1):(2*n),]<- rep(c(0.7),each=(n))
-
-sigma<-  sqrt(1.7661/30)
-
-for(b in 1:batchN){
-yBatch[,,b]<- matrix( rnorm(n*p*2,mu,sigma), n*K, p)
-}
-
-mu_ini<- matrix(mu0,K,p)
-
-
-dataHist<- data.frame("Mu"=as.factor(mu),    "y"= yBatch[,,1])
-
-ggplot(dataHist, aes(y, fill = as.factor(Mu))) + geom_histogram(alpha = 0.2,bins = 30,  position="identity")
-
-
-
-
-computeMError<- function(jk, n,batchN){
-error<- numeric(batchN)
-for(b in 1:batchN){
-tM0<- rank(jk$mu[,1,b])-1
-trueM<- rep(tM0,each=n)
-error[b]<- sum ( jk$M[,b] != trueM) /K/n
-if( min(jk$w[,b])<0.1){
-error[b]<- NA
-}
-}
-error
-}
-
-computeRMSE<- function(jk, mu0, batchN){
-rmse<- numeric(batchN)
-for(b in 1:batchN){
-newOrder<- order(jk$mu[,1,b],decreasing = F)
-
-muEst<-as.matrix(jk$mu[,,b])[newOrder,]
-
-rmse[b]<- sqrt( mean((muEst - mu0)^2))
-if( min(jk$w[,b])<0.1)
-rmse[b]<- NA
-}
-rmse
-}
-
-
-
-experiment<- function(n){
-  
-  K<- 2
-  p<- 1
-  batchN<- 1000
-  
-  yBatch<- array(0,dim = c(n*K,p,batchN))
-  
-  mu0<- c(c(0.3), c(0.7))
-  mu<- matrix( 0,n*2,p)
-  mu[1:(n),]<- rep(c(0.3),each=(n))
[…]

### 2016-10-19 Rmd/convergenceRatesBigK.r

@@ -1,131 +1,0 @@
-
-require('ggplot2')
-require('reshape')
-require('jkmeans')
-
-setwd("~/git/jkmeans/Rmd/")
-
-n<- 100
-K<- 10
-p<- 1
-batchN<- 1000
-
-yBatch<- array(0,dim = c(n*K,p,batchN))
-
-mu0<- c(1:K)
-mu<- matrix( 0,n*K,p)
-mu <- rep(mu0, each = n)
-sigma<-  0.5
-
-for(b in 1:batchN){
-  yBatch[,,b]<- matrix( rnorm(n*p*K,mu,sigma), n*K, p)
-}
-
-mu_ini<- matrix(mu0,K,p)
-
-
-
-dataHist<- data.frame("Mu"=as.factor(mu),    "y"= yBatch[,,1])
-
-ggplot(dataHist, aes(y, fill = as.factor(Mu))) + geom_histogram(alpha = 0.2,bins = 30,  position="identity")
-
-
-computeMError<- function(jk, n,batchN){
-  error<- numeric(batchN)
-  for(b in 1:batchN){
-    tM0<- rank(jk$mu[,1,b])-1
-    trueM<- rep(tM0,each=n)
-    error[b]<- sum ( jk$M[,b] != trueM) /K/n
-    if( min(jk$w[,b])<0.01){
-        error[b]<- NA
-    }
-  }
-  error
-}
-
-computeRMSE<- function(jk, mu0, batchN){
-  rmse<- numeric(batchN)
-  for(b in 1:batchN){
-    newOrder<- order(jk$mu[,1,b],decreasing = F)
-    
-    muEst<-as.matrix(jk$mu[,,b])[newOrder,]
-   
-    rmse[b]<- sqrt( mean((muEst - mu0)^2))
-    if( min(jk$w[,b])<0.01)
-      rmse[b]<- NA
-  }
-  rmse
-}
-
-experiment<- function(n){
-  
-  p<- 1
-  batchN<- 1000
-  
-  yBatch<- array(0,dim = c(n*K,p,batchN))
-  K<- 10
-  
-  mu0<- c(1:K)
-  mu<- matrix( 0,n*K,p)
-  mu <- rep(mu0, each = n)
-  sigma<-  0.5
-
-
-  for(b in 1:batchN){
-    yBatch[,,b]<- matrix( rnorm(n*p*K,mu,sigma), n*K, p)
-  }
-  
-  mu_ini<- matrix(mu0,K,p)
-  
-  
[…]

### 2016-10-19 Rmd/convergenceRatesImbalanced.R

@@ -1,141 +1,0 @@
-
-require('ggplot2')
-require('reshape')
-require('jkmeans')
-
-setwd("~/git/jkmeans/Rmd/")
-
-n<- 100
-K<- 2
-p<- 1
-batchN<- 1000
-
-yBatch<- array(0,dim = c(n*K,p,batchN))
-
-mu0<- c(c(0.3), c(0.7))
-mu<- matrix( 0,n*2,p)
-mu[1:(n/2),]<- rep(c(0.3),each=(n/2))
-mu[(n/2+1):(2*n),]<- rep(c(0.7),each=(n/2*3))
-
-sigma<-  sqrt(1.7661/30)
-
-for(b in 1:batchN){
-  yBatch[,,b]<- matrix( rnorm(n*p*2,mu,sigma), n*K, p)
-}
-
-mu_ini<- matrix(mu0,K,p)
-
-
-
-
-dataHist<- data.frame("Mu"=as.factor(mu),    "y"= yBatch[,,1])
-
-ggplot(dataHist, aes(y, fill = as.factor(Mu))) + geom_histogram(alpha = 0.2,bins = 30,  position="identity")
-
-
-
-computeMError<- function(jk, n,batchN){
-  error<- numeric(batchN)
-  for(b in 1:batchN){
-    tM0<- rank(jk$mu[,1,b])-1
-    trueM<- c(rep(tM0[1],(n/2)),rep(tM0[2],(n/2*3)))
-    error[b]<- sum ( jk$M[,b] != trueM) /K/n
-    if( min(jk$w[,b])<0.1){
-        error[b]<- NA
-    }
-  }
-  error
-}
-
-computeRMSE<- function(jk, mu0, batchN){
-  rmse<- numeric(batchN)
-  for(b in 1:batchN){
-    newOrder<- order(jk$mu[,1,b],decreasing = F)
-    
-    muEst<-as.matrix(jk$mu[,,b])[newOrder,]
-   
-    rmse[b]<- sqrt( mean((muEst - mu0)^2))
-    if( min(jk$w[,b])<0.1)
-      rmse[b]<- NA
-  }
-  rmse
-}
-
-
-
-
-experiment<- function(n){
-  
-  K<- 2
-  p<- 1
-  batchN<- 1000
-  
-  yBatch<- array(0,dim = c(n*K,p,batchN))
-  
-  mu0<- c(c(0.3), c(0.7))
[…]

### 2016-10-19 Rmd/rmd2rscript.r

@@ -1,46 +1,0 @@
-#' ## script for converting .Rmd files to .R scripts.
-
-#' #### Kevin Keenan 2014
-#' 
-#' This function will read a standard R markdown source file and convert it to 
-#' an R script to allow the code to be run using the "source" function.
-#' 
-#' The function is quite simplisting in that it reads a .Rmd file and adds 
-#' comments to non-r code sections, while leaving R code without comments
-#' so that the interpreter can run the commands.
-#' 
-#' 
-rmd2rscript <- function(infile){
-  # read the file
-  flIn <- readLines(infile)
-  # identify the start of code blocks
-  cdStrt <- which(grepl(flIn, pattern = "```{r*", perl = TRUE))
-  # identify the end of code blocks
-  cdEnd <- sapply(cdStrt, function(x){
-    preidx <- which(grepl(flIn[-(1:x)], pattern = "```", perl = TRUE))[1]
-    return(preidx + x)
-  })
-  # define an expansion function
-  # strip code block indacators
-  flIn[c(cdStrt, cdEnd)] <- ""
-  expFun <- function(strt, End){
-    strt <- strt+1
-    End <- End-1
-    return(strt:End)
-  }
-  idx <- unlist(mapply(FUN = expFun, strt = cdStrt, End = cdEnd, 
-                       SIMPLIFY = FALSE))
-  # add comments to all lines except code blocks
-  comIdx <- 1:length(flIn)
-  comIdx <- comIdx[-idx]
-  for(i in comIdx){
-    flIn[i] <- paste("#' ", flIn[i], sep = "")
-  }
-  # create an output file
-  nm <- strsplit(infile, split = "\\.")[[1]][1]
-  flOut <- file(paste(nm, "[rmd2r].R", sep = ""), "w")
-  for(i in 1:length(flIn)){
-    cat(flIn[i], "\n", file = flOut, sep = "\t")
-  }
-  close(flOut)
-}

### 2016-10-19 Rmd/convergenceRatesBalancedBigK.Rmd
sync

@@
 
+```{r include=FALSE}
+
+nSeries<- c(seq(20,500,by = 20))
+
+BEsim<- function(n){
+  p<- 1
+  batchN<- 1000
+
+  yBatch<- array(0,dim = c(n*K,p,batchN))
+
+  mu0<- c(1:K)
+  mu<- matrix( 0,n*K,p)
+  mu <- rep(mu0, each = n)
+
+  for(b in 1:batchN){
+    yBatch[,,b]<- matrix( rnorm(n*p*K,mu,sigma), n*K, p)
+  }
+
+  dist<- abs(outer(c(yBatch), mu0, "-"))
+  bayesclassified <- apply(dist, 1, function(x){c(1:K)[x==min(x)]})
+  #bayes error
+  (sum(bayesclassified != rep(c(1:K),each=n))/n/K/batchN)
+}
+
+bayesErorr<- BEsim(100)
+
+```
+
 #Misclassification error
@@
 # p+ geom_errorbar( aes(x=n,ymax = MC14ErrorU, ymin=MC14ErrorL, colour=Model),width=0.2)
-p+ theme_bw()
+p+ theme_bw() + geom_hline(yintercept = bayesErorr)
 # pE + theme_bw()

### 2016-10-19 Rmd/convergenceRatesImbalanced.Rmd
sync

@@
 
-# if(FALSE)
+if(FALSE)
 {
@@
 
+```{r include=FALSE}
+
+nSeries<- c(seq(20,500,by = 20))
+
+BEsim<- function(n){
+  p<- 1
+  batchN<- 1000
+
+  sigma<-  sqrt(1.7661/30)
+
+  yBatch<- array(0,dim = c(n*K,p,batchN))
+
+  mu0<- c(c(0.3), c(0.7))
+  mu<- matrix( 0,n*2,p)
+  mu[1:(n/2),]<- rep(c(0.3),each=(n/2))
+  mu[(n/2+1):(2*n),]<- rep(c(0.7),each=(n/2*3))
+
+  for(b in 1:batchN){
+    yBatch[,,b]<- matrix( rnorm(n*p*2,mu,sigma), n*K, p)
+  }
+
+  dist<- abs(outer(c(yBatch), mu0, "-"))
+
+  dist<- t(t(exp(-dist^2/sigma^2/2))*c(0.25,0.75))
+
+  bayesclassified <- apply(dist, 1, function(x){c(1:K)[x==max(x)]})
+  #bayes error
+  (sum(bayesclassified != c(rep(1,n/2),rep(2,n/2*3)))/n/K/batchN)
+}
+
+bayesErorr<- BEsim(100)
+
+```
+
 #Misclassification error
@@
 # p+ geom_errorbar( aes(x=n,ymax = MC14ErrorU, ymin=MC14ErrorL, colour=Model),width=0.2)
-p+ theme_bw()
+p+ theme_bw() + geom_hline(yintercept = bayesErorr)
 pE + theme_bw()

### 2016-10-20 Rmd/convergenceRatesImbalancedBigK.Rmd
ordered Mu after clustering is done, before passing back to R

--- 2016-10-19 Rmd/convergenceRatesBalancedBigK.Rmd
@@
 mu0<- c(1:K)
-mu<- matrix( 0,n*K,p)
-mu <- rep(mu0, each = n)
+# mu<- matrix( 0,n*K,p)
+# mu <- rep(mu0, each = n)
+
+mu<- c(rep(1,n*3/2), rep(2, n/2),  rep(3, n/2*3), rep(4,n/2), rep(5,n))
+
 sigma<-  0.5
@@
 
-Here is a view of the two clusters
+Here is a view of the clusters
 ```{r}
@@
 ```{r}
-computeMError<- function(jk, n,batchN){
+computeMError<- function(jk, batchN, trueM){
   error<- numeric(batchN)
   for(b in 1:batchN){
-    tM0<- rank(jk$mu[,1,b])-1
-    trueM<- rep(tM0,each=n)
-    error[b]<- sum ( jk$M[,b] != trueM) /K/n
+    # tM0<- rank(jk$mu[,1,b])-1
+    # trueM<- rep(tM0,each=n)
+    error[b]<- mean ( jk$M[,b] != (trueM -1 ))
     if( min(jk$w[,b])<0.01){
@@
 
-computeRMSE<- function(jk, mu0, batchN){
+computeRMSE<- function(jk, batchN, trueMu){
   rmse<- numeric(batchN)
@@
     newOrder<- order(jk$mu[,1,b],decreasing = F)
-
     muEst<-as.matrix(jk$mu[,,b])[newOrder,]
 
-    rmse[b]<- sqrt( mean((muEst - mu0)^2))
+    rmse[b]<- sqrt( mean((muEst - trueMu)^2))
     if( min(jk$w[,b])<0.01)
@@
   p<- 1
-  batchN<- 1000
+  batchN<- 100
 
@@
   mu0<- c(1:K)
-  mu<- matrix( 0,n*K,p)
-  mu <- rep(mu0, each = n)
+  mu<- c(rep(1,n*3/2), rep(2, n/2),  rep(3, n/2*3), rep(4,n/2), rep(5,n))
+
+  trueM<- mu #same as mu
+  trueMu<- mu0
 
@@
 
-  error<- sapply(runRes,function(x){computeMError(x,n,batchN)})
-  rmse<- sapply(runRes,function(x){computeRMSE(x,mu0,batchN)})
+  error<- sapply(runRes,function(x){computeMError(x,batchN, trueM)})
+  rmse<- sapply(runRes,function(x){computeRMSE(x,batchN, trueMu)})
 
@@
 
-nSeries<- c(seq(20,500,by = 20))
+nSeries<- c(seq(20,200,by = 20),300,400,500)
 
@@
 
-  save(result, file="resultRatesBigK.Rda")
+  save(result, file="resultRatesImbalancedBigK.Rda")
 }
@@
 ```{r include=FALSE}
-
-nSeries<- c(seq(20,500,by = 20))
-
-BEsim<- function(n){
-  p<- 1
-  batchN<- 1000
-
-  yBatch<- array(0,dim = c(n*K,p,batchN))
-
-  mu0<- c(1:K)
-  mu<- matrix( 0,n*K,p)
-  mu <- rep(mu0, each = n)
-
-  for(b in 1:batchN){
-    yBatch[,,b]<- matrix( rnorm(n*p*K,mu,sigma), n*K, p)
-  }
-
-  dist<- abs(outer(c(yBatch), mu0, "-"))
-  bayesclassified <- apply(dist, 1, function(x){c(1:K)[x==min(x)]})
-  #bayes error
-  (sum(bayesclassified != rep(c(1:K),each=n))/n/K/batchN)
-}
-
-bayesErorr<- BEsim(100)
+# 
+# nSeries<- c(seq(20,500,by = 20))
+# 
+# 
+# 
+# BEsim<- function(n){
+#   p<- 1
+#   batchN<- 1000
+#   
+#   yBatch<- array(0,dim = c(n*K,p,batchN))
+#   
+#   mu0<- c(1:K)
+#   mu<- matrix( 0,n*K,p)
+#   mu <- rep(mu0, each = n)
+#   
+#   for(b in 1:batchN){
+#     yBatch[,,b]<- matrix( rnorm(n*p*K,mu,sigma), n*K, p)
+#   }
+#   
+#   dist<- abs(outer(c(yBatch), mu0, "-"))
+#   bayesclassified <- apply(dist, 1, function(x){c(1:K)[x==min(x)]})
+#   #bayes error
+#   (sum(bayesclassified != rep(c(1:K),each=n))/n/K/batchN)
+# }
+# 
+# 
+# bayesErorr<- BEsim(100)
 
@@
 
-load("resultRatesBigK.Rda")
+load("resultRatesImbalancedBigK.Rda")
 
@@
 # p+ geom_errorbar( aes(x=n,ymax = MC14ErrorU, ymin=MC14ErrorL, colour=Model),width=0.2)
-p+ theme_bw() + geom_hline(yintercept = bayesErorr)
+p+ theme_bw() #+ geom_hline(yintercept = bayesErorr)
 # pE + theme_bw()

### 2016-10-20 jkmeans/src/mixture.hpp

@@ -45,7 +45,7 @@ class Mixture {
 
     if (useKmeansIni) {
       mat means;
-      bool status = kmeans(means, trans(y), K, random_subset, 10, false);
+      bool status = kmeans(means, trans(y), K, static_spread, 10, false);
       mu = means.t();
     } else {
       mu = meansInput;
@@ -284,7 +284,14 @@ class Mixture {
     }
   }
 
-  uvec clusteringMAP() { return arma::index_max(zeta, 1); }
+  uvec clusteringMAP() {
+    uvec indices = sort_index(mu.col(0));
+    mat reordered_zeta = zeta.cols(indices);
+
+    // cout << indices << endl;
+
+    return arma::index_max(reordered_zeta, 1);
+  }
 
   // void runQNEM(int steps) {
   // Expectation();

### 2016-10-30 docs/ver1.tex
update write-up & BV in mean

@@
 \title
-    {?? Mixture Model}
+    {One-Hot Membership Model}
 
@@
 
-variance-bias tradeoff?
-reducing variability on the weight
-
-\section{?? Mixture Model}
+\section{One-Hot Membership Model}
 
@@
 
-On a separable metrics space $\{\mathcal{X},\mathcal{F}\}$, we assume each data $y_i$ has the following probability density as the likelihood:
+Let $y_i$ be the observed data following the distribution $F$ with parameter $\theta_i$. Consider a  membership model with $\kappa$ possible membership:
 
@@
 \begin{aligned}
-& \pi(y_i) \sim G(\theta_i) \\
-& \pi(\theta_i) = \sum_{k=1}^{\kappa} w_{k,i} \delta_{\theta_{k}}(\theta_i) \\
-& w_{k,i} = \frac{ u_{k,i} v_{k} } { \sum_{k=1}^{\kappa}  u_{k,i} v_{k} }\\
-& u_{k,i} \in \{0,1\}
+& \pi(y_i) \stackrel{indep}{\sim} F(y_i|\theta_i) \\
+& \pi(\theta_i) = \sum_{k=1}^{\kappa} z_{k,i} \delta_{\theta^*_{k}}(\theta_i) \\
 \end{aligned}
-\label{marginal_lik}
+\label{membership_model}
 \end{equation}
-where $\delta_{\theta_{k}}(y_i)$ is a probabilty density corresponds to a component distribution with parameter $\theta_{k}$, and $w_{k,i}$ is the component weight that varies by index $i$ and $\sum_{k=1}^{\kappa} w_{k,i} =1$.
-The weight varying is due to the randomness in $u_{k,i}$.
-Due to the $0$'s $u_{k,i}$ introduces, this induces sparsity in the membership probabilty $w_{k,i}$.
+where $\{ z_{.,i}\}$ denotes the membership with only one $1$ and $(k-1)$ many $0$'s.
 
-\subsection{Estimation}
-
-We now use the latent variable $\{ z_{1,i},z_{2,i},\ldots z_{\kappa,i} \} \sim MultiNomial (\{ w_{1,i},w_{2,i},\ldots w_{\kappa,i}\} )$ that takes value of $\{0,\ldots,1,\ldots 0\}$ that assign randomly one $1$ to one of $\kappa$ vertices in the simplex.
-The likelihood can be rewritten as:
+To allow borrowing of strength among $\{ z_{.,i}\}$, we assume $z_{.,i}$ follows:
 
@@
 \begin{aligned}
-& [ y_i ] =   \sum_{k=1}^{\kappa} \int z_{k,i} \delta_{\theta_{k}}(y_i) P(d z_{k,i}) = \int \sum_{k=1}^{\kappa} z_{k,i} \delta_{\theta_{k}}(y_i) P(d z_{k,i})
+& \pi (z_{.,i}) \propto  \prod_{k=1}^{\kappa} (u_{k,i}v_{k} )^{z_{k,i}}  \\
+ & v_{k} \sim Dir_{\kappa}(1,1,\ldots 1)\\
+  & u_{.,i} \stackrel{iid}{\sim} Dir_{\kappa}(\epsilon,\epsilon,\ldots \epsilon)\\
 \end{aligned}
+\label{one_hot}
 \end{equation}
-where $P(d z_{k,i})$ is the measure of the multinomial distribution aforementioned, the last equation is due to Fubini theorem.
-Note due to the $z_{k,i}$ takes only one $1$ out of $K$, this allows a simple log-density augmented with $z_{k,i}$:
+where $\epsilon$ is a small number close to $0$, which causes $u_{k,i}$ to have only one value close to $1$ and the rest close to $0$.
+Due to the $z_{.,i}$ will take one-of-$\kappa$ memberships with probability almost $1$, we refer this  as one-hot membership model.
+
+To compare with other methods, note when $\{z_{.,i}\}$ is assumed to be independent for each $i$, (\ref{membership_model}) is exactly the same as the k-means model; when $u_{k,i}$ is fixed to $1/\kappa$ for all $k,i$, the membership $z_{.,i}$ can be integrated out to obtain general finite mixture model.
+
+The key difference is in one-hot membership model, the multinomial distribution of $z_{.,i}$ has its probability weight concentrated to one vertex, so that the estimate of $z_{.,i}$ is obtained via maximization, instead of expectation or sampling as in the general mixture model.
+
+\subsection{Estimation}
 
@@
 \begin{algorithm}[H]
-\caption{EM algorithm}\label{alg:euclid}
+\caption{Estimation algorithm}\label{alg:euclid}
 \begin{algorithmic}[1]
-% \Procedure{Euclid}{$a,b$}\Comment{The g.c.d. of a and b}
-% \State $r\gets a\bmod b$
 \While{$||\theta^{(s)}-\theta^{(s+1)}||>\epsilon$}%\Comment{We have the answer if r is 0}
-\State Maximize over $\{ u_{i,k}\}_k$ by $ \underset{\{u_{i,k}\}_k} {\text{argmax }}\{u_{i,k}  v_{k}\delta_{\theta_{k}}(y_i)\}$ for all $i$; \Comment{\small pick top J out K based on (\ref{marginal_lik})}
-\State Compute $ \mathbb{E}(z_{k,i})= \frac{w_{k,i} \delta_{\theta_{k}}(y_i)}{\sum_k w_{k,i} \delta_{\theta_{k}}(y_i)}$ with $w_{k,i} = u_{i,k} v_{k}$; \Comment{E step  based on (\ref{conditional_lik})}
-\State Set $\theta^{(s+1)}_{k}= \text{argmax} \sum_{k=1}^{\kappa} \mathbb{E}z_{k,i} \log \delta_{\theta_{k}}(y_i) $ for all $k$ \Comment{M step: obtain MLE based on (\ref{conditional_lik})}
-\State Set $v^{(s+1)}_{k}=  \sum_{i}\mathbb{E}z_{k,i} /  \sum_{i}\sum_{k}\mathbb{E}z_{k,i}$ \Comment{M step: obtain MLE based on (\ref{conditional_lik})}\label{flex_w_step}
+\State Maximize over $\{ z_{i,k}\}_k$ by $ \underset{\{z_{i,k}\}_k} {\text{argmax }}\{z_{i,k}  v_{k}F (y_i|\theta^*_k)\}$ for all $i$;
+\State Update $\theta^{*}_{k}= \text{argmax} \sum_{i=1}^{n} z_{k,i} \log F(y_i|\theta^*_k) $ for all $k$ 
+\State Update $v_{k}=  \sum_{i} z_{k,i} /  \sum_{i}\sum_{k}z_{k,i}$ 
 \EndWhile
-% \State \textbf{return} $b$\Comment{The gcd is b}
-% \EndProcedure
 \end{algorithmic}
@@
 
-variance reduction in the weight parameter
+\subsection{Empirical Result: Bias-Variance Trade-off}
+
+We generate data $y_i$ from $N(\mu_i, 0.5^2)$, using $75\%$ with $\mu_i =1$ and $25\%$ with $\mu_i =2$.
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}[t]{.8\columnwidth}
+\includegraphics[width=\textwidth]{BV2.pdf}
+\caption{The bias-variance trade-off.}
+\end{subfigure}
+\begin{subfigure}[t]{.8\columnwidth}
+\includegraphics[width=\textwidth]{BVMeanDist.pdf}
+\caption{The distribution of MLE for $\mu_1$ and $\mu_2$ over 1,000 simulation data sets, with total $40$ data points.}
+\end{subfigure}
+\begin{subfigure}[t]{.8\columnwidth}
+\includegraphics[width=\textwidth]{BVMeanDist3.pdf}
+\caption{The distribution of MLE for $\mu_1$ and $\mu_2$ over 1,000 simulation data sets, with total $150$ data points.}
+\end{subfigure}
+\end{figure}
+
+\subsection{Variance Reduction in Small Sample Size}
+
+Let superscript denote the variable to be integrated over.
+Using variance decomposition, the variance of the parameter estimator $\hat\theta$ has:
+
+$$\mbox{Var}^{y} (\hat\theta) = \mbox{E}^{z}\mbox{Var}^{y} (\hat\theta| z) + \mbox{Var}^{z}\mbox{E}^{y} (\hat \theta| z)$$
+
+Using subscript $FMM$ and $OH$ to denote the different models.
+As each $z_i$ is maximized to a fixed value in each dataset in one-hot model, It can be immediately seen that $\mbox{Var}_{OH}^{z}\mbox{E}^{y} (\hat \theta| z)=0$ and $\mbox{E}^{z}_{OH}\mbox{Var}^{y} (\hat\theta| z) = \mbox{Var}^{y}_{OH} (\hat\theta| z)$; whereas each $z_i$ is random in finite mixture model, $\mbox{Var}_{FMM}^{z}\mbox{E}^{y} (\hat \theta| z)\ge 0$.
+Therefore, it suffices to have $\mbox{Var}_{OH}^{y} (\hat\theta| z) - \mbox{E}^{z}\mbox{Var}^{y}_{FMM} (\hat\theta| z) < \mbox{Var}_{FMM}^{z}\mbox{E}^{y} (\hat \theta| z)$ for a reduction in variance.
 
 \section{Simulation}
-
-large K with small J in 1d
-
-high p setting
 

### 2016-10-30 Rmd/computeBV.r

@@ -1,0 +1,213 @@
+require("ggplot2")
+
+setwd("~/git/jkmeans/Rmd/")
+
+#good
+load("sims/raw_result_2_0_0.5_1_0.Rda")
+
+# load("sims/raw_result_2_0_0.5_0_1.Rda")
+
+nseries<- length(raw_test_results)
+K<- 2
+
+#likelihood
+
+loglik_GMM<- function(y, mu,sigma2,w){
+  diff<- outer(y,mu,"-")
+  sum(log(exp(-diff^2/sigma2/2)/sqrt(sigma2)) %*% (as.matrix(w)))
+}
+
+
+#get Bias Variance for J=j in GMM
+getBVMean <- function(j,n){
+  #histogram of the mean
+  
+  raw_n<- raw_test_results[[n]]
+  
+  muHat<- raw_n$GMMList[[j]]$mu
+  wHat<- raw_n$GMMList[[j]]$w
+  
+  sigma2Hat<- raw_n$GMMList[[j]]$sigma2
+  
+  cutoff<- 0.05
+  exclude<- (wHat[1,]>(1-cutoff) | wHat[1,]<cutoff)
+  
+  muHat<- muHat[,,!exclude]
+  wHat<- wHat[,!exclude]
+  sigma2Hat<- sigma2Hat[!exclude]
+  yBatch<- raw_n$yBatch[,,!exclude]
+  
+  batchN<- batchN- sum(exclude)
+  
+  # loglik<- sapply(c(1:batchN), function(i){
+  #   y<- yBatch[,i]
+  #   mu<- muHat[,i]
+  #   w<- wHat[,i]
+  #   sigma2<- sigma2Hat[i]
+  #   loglik_GMM(y,mu,sigma2,w)
+  #   })
+  # 
+  # meanLoglik<- mean(loglik)
+  # 
+  # dfPlot <- data.frame(
+  #   "Mu" = c(mu), 
+  #   "Component" = as.factor(rep(c(1:K), batchN))
+  # )
+  # 
+  # meanHist<- ggplot(dfPlot, aes(x=Mu, fill=Component)) + geom_histogram(alpha=0.5, position="identity", bins = 100) + theme_bw()
+  
+  #bias variance of the mean
+  
+  Emuhat<- rowMeans(muHat)
+  bias <- Emuhat - raw_n$mu0
+  variance <-  rowMeans((muHat - Emuhat)^2)
+  
[…]

### 2016-10-30 Rmd/postSim.r

@@ -1,0 +1,104 @@
+require("ggplot2")
+
+setwd("~/git/jkmeans/Rmd/")
+
+#good
+load("sims/raw_result_2_0_0.5_1_0.Rda")
+
+nseries<- length(raw_test_results)
+K<- 2
+
+#likelihood
+
+loglik_GMM<- function(y, mu,sigma2,w){
+  diff<- outer(y,mu,"-")
+  sum(log(exp(-diff^2/sigma2/2)/sqrt(sigma2)) %*% (as.matrix(w)))
+}
+
+
+#get Bias Variance for J=j in GMM
+getBVMean <- function(j){
+  #histogram of the mean
+  
+  muHat<- raw_n$GMMList[[j]]$mu
+  wHat<- raw_n$GMMList[[j]]$w
+  
+  sigma2Hat<- raw_n$GMMList[[j]]$sigma2
+  
+  cutoff<- 0.05
+  exclude<- (wHat[1,]>(1-cutoff) | wHat[1,]<cutoff)
+  
+  muHat<- muHat[,,!exclude]
+  wHat<- wHat[,!exclude]
+  sigma2Hat<- sigma2Hat[!exclude]
+  yBatch<- raw_n$yBatch[,,!exclude]
+  
+  batchN<- batchN- sum(exclude)
+  
+  loglik<- sapply(c(1:batchN), function(i){
+    y<- yBatch[,i]
+    mu<- muHat[,i]
+    w<- wHat[,i]
+    sigma2<- sigma2Hat[i]
+    loglik_GMM(y,mu,sigma2,w)
+    })
+  
+  meanLoglik<- mean(loglik)
+  
+  # dfPlot <- data.frame(
+  #   "Mu" = c(mu), 
+  #   "Component" = as.factor(rep(c(1:K), batchN))
+  # )
+  # 
+  # meanHist<- ggplot(dfPlot, aes(x=Mu, fill=Component)) + geom_histogram(alpha=0.5, position="identity", bins = 100) + theme_bw()
+  
+  #bias variance of the mean
+  
+  Emuhat<- rowMeans(muHat)
+  bias <- Emuhat - raw_n$mu0
+  variance <-  rowMeans((muHat - Emuhat)^2)
+  
+  MSE <- rowMeans((muHat - raw_n$mu0)^2)
+  
+  
+  # bias variance of the zeta
[…]

### 2016-10-30 Rmd/sim2.r

@@ -1,13 +1,14 @@
-K=2
-balanced = T
-sigma = 1
-initial_truth = T
-n= 100
-
 p<- 1
 batchN<- 1000
+n<- 100
+K<- 2
+sigma<- 0.5
+fix_sigma2 <- FALSE
 
 yBatch<- array(0,dim = c(n*K,p,batchN))
+balanced <- FALSE
+initial_truth<- TRUE
+
 
 M <- numeric() #label
 
@@ -35,55 +36,12 @@ for(b in 1:batchN){
 useKmeansIni =  !initial_truth
 
 kMeansList <- lapply(c(1:K), function(J){
-    jkmeans::jkmeansEMBatch(yBatch, k=K, j = J,  1000,tol = 1E-10,useKmeansIni = useKmeansIni, meansIni = as.matrix(mu0), fixW = T,sigma2_ini = sigma^2)
+    jkmeans::jkmeansEMBatch(y= yBatch, k=K, j = J, steps =  1000,tol = 1E-10,useKmeansIni = useKmeansIni, meansIni = as.matrix(mu0), fixW = T,sigma2_ini = sigma^2, fixSigma2 = fix_sigma2,flexJ = F,zetaTrunc = 0.1)
 })
-
-
-kMeans
 
 GMMList <- lapply(c(1:K), function(J){
-    jkmeans::jkmeansEMBatch(yBatch, k=K, j = J,  1000,tol = 1E-10,useKmeansIni = useKmeansIni, meansIni = as.matrix(mu0), fixW = F,sigma2_ini = sigma^2)
+    jkmeans::jkmeansEMBatch(y = yBatch, k=K, j = J, steps =  1000,tol = 1E-10,useKmeansIni = useKmeansIni, meansIni = as.matrix(mu0), fixW = F,sigma2_ini = sigma^2, fixSigma2 = fix_sigma2,flexJ = F,zetaTrunc = 0.1)
 })
-
-    #functions to compute the error
-    computeMError<- function(jk, batchN, trueM){
-      error<- numeric(batchN)
-      for(b in 1:batchN){
-        error[b]<- mean ( jk$M[,b] != (trueM -1 ))
-        if( min(jk$w[,b])<0.01){
-          error[b]<- NA
-        }
-      }
-      error
-    }
-
-    computeRMSE<- function(jk, batchN, trueMu){
-      rmse<- numeric(batchN)
-      for(b in 1:batchN){
-        newOrder<- order(jk$mu[,1,b],decreasing = F)
[…]

### 2016-10-30 Rmd/simulation2.R

@@ -1,0 +1,55 @@
+simulation_raw<- function(K=2, balanced = T, sigma = 1, initial_truth = T, fix_sigma2 = F){
+  
+  experiment<- function(n){
+    
+    p<- 1
+    batchN<- 1000
+    
+    yBatch<- array(0,dim = c(n*K,p,batchN))
+    
+    M <- numeric() #label
+    
+    for(k in 1:K){
+      if(balanced)
+        M<- c(M, rep(k,n))
+      else{
+        if(k==K & K%%2==1){
+          M<- c(M, rep(k, n))
+        }else{
+          M<- c(M, rep(k, n/2 * ((k %% 2)*2+1)  ))
+        }
+      }
+    }
+    
+    w<- table(M)/sum(table(M))
+    
+    mu<- 1*M
+    mu0<- 1*c(1:K)
+    
+    for(b in 1:batchN){
+      yBatch[,,b]<- matrix( rnorm(n*p*K,mu,sigma), n*K, p)
+    }
+    
+    useKmeansIni =  !initial_truth
+    
+    kMeansList <- lapply(c(1:K), function(J){
+      jkmeans::jkmeansEMBatch(y= yBatch, k=K, j = J, steps =  1000,tol = 1E-10,useKmeansIni = useKmeansIni, meansIni = as.matrix(mu0), fixW = T,sigma2_ini = sigma^2, fixSigma2 = fix_sigma2,flexJ = F,zetaTrunc = 0.1)
+    })
+    
+    GMMList <- lapply(c(1:K), function(J){
+      jkmeans::jkmeansEMBatch(y = yBatch, k=K, j = J, steps =  1000,tol = 1E-10,useKmeansIni = useKmeansIni, meansIni = as.matrix(mu0), fixW = F,sigma2_ini = sigma^2, fixSigma2 = fix_sigma2,flexJ = F,zetaTrunc = 0.1)
+    })
+    
+    list("M"= M,"mu0"=mu0, "yBatch"= yBatch, "kMeansList"=kMeansList, "GMMList"= GMMList)
+    
+  }
+  
+  nSeries<- c(seq(20,200,by = 20),300,400,500)
+  
+  raw_test_results <- lapply(nSeries, experiment)
+  
+  filename = paste("sims/raw_result_",paste(K,balanced ,sigma , initial_truth, as.numeric(fix_sigma2),sep =  "_"),".Rda",sep="")
+  save(raw_test_results, file=filename)
+  
+  # result
+}

### 2016-10-30 jkmeans/R/RcppExports.R

@@ -1,11 +1,11 @@
 # This file was generated by Rcpp::compileAttributes
 # Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
 
-jkmeansEM <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, flexJ = FALSE, zetaTrunc = 0.01, useKmeansIni = TRUE, meansIni = 0L, sigma2_ini = 0.1) {
-    .Call('jkmeans_jkmeansEM', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini)
+jkmeansEM <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, flexJ = FALSE, zetaTrunc = 0.01, useKmeansIni = TRUE, meansIni = 0L, sigma2_ini = 0.1, fixSigma2 = FALSE) {
+    .Call('jkmeans_jkmeansEM', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini, fixSigma2)
 }
 
-jkmeansEMBatch <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, flexJ = FALSE, zetaTrunc = 0.01, useKmeansIni = TRUE, meansIni = 0L, sigma2_ini = 0.1) {
-    .Call('jkmeans_jkmeansEMBatch', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini)
+jkmeansEMBatch <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, flexJ = FALSE, zetaTrunc = 0.01, useKmeansIni = TRUE, meansIni = 0L, sigma2_ini = 0.1, fixSigma2 = FALSE) {
+    .Call('jkmeans_jkmeansEMBatch', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini, fixSigma2)
 }
 

### 2016-10-30 jkmeans/src/RcppExports.cpp

@@ -7,8 +7,8 @@
 using namespace Rcpp;
 
 // jkmeansEM
-Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps, double tol, bool fixW, bool flexJ, double zetaTrunc, bool useKmeansIni, const arma::mat& meansIni, double sigma2_ini);
-RcppExport SEXP jkmeans_jkmeansEM(SEXP ySEXP, SEXP kSEXP, SEXP jSEXP, SEXP stepsSEXP, SEXP tolSEXP, SEXP fixWSEXP, SEXP flexJSEXP, SEXP zetaTruncSEXP, SEXP useKmeansIniSEXP, SEXP meansIniSEXP, SEXP sigma2_iniSEXP) {
+Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps, double tol, bool fixW, bool flexJ, double zetaTrunc, bool useKmeansIni, const arma::mat& meansIni, double sigma2_ini, bool fixSigma2);
+RcppExport SEXP jkmeans_jkmeansEM(SEXP ySEXP, SEXP kSEXP, SEXP jSEXP, SEXP stepsSEXP, SEXP tolSEXP, SEXP fixWSEXP, SEXP flexJSEXP, SEXP zetaTruncSEXP, SEXP useKmeansIniSEXP, SEXP meansIniSEXP, SEXP sigma2_iniSEXP, SEXP fixSigma2SEXP) {
 BEGIN_RCPP
     Rcpp::RObject __result;
     Rcpp::RNGScope __rngScope;
@@ -23,13 +23,14 @@ BEGIN_RCPP
     Rcpp::traits::input_parameter< bool >::type useKmeansIni(useKmeansIniSEXP);
     Rcpp::traits::input_parameter< const arma::mat& >::type meansIni(meansIniSEXP);
     Rcpp::traits::input_parameter< double >::type sigma2_ini(sigma2_iniSEXP);
-    __result = Rcpp::wrap(jkmeansEM(y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini));
+    Rcpp::traits::input_parameter< bool >::type fixSigma2(fixSigma2SEXP);
+    __result = Rcpp::wrap(jkmeansEM(y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini, fixSigma2));
[…]

### 2016-10-30 jkmeans/src/jkmeans.cpp

@@ -15,16 +15,20 @@ using namespace arma;
 Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps = 1000,
                      double tol = 1E-8, bool fixW = true, bool flexJ = false,
                      double zetaTrunc = 0.01, bool useKmeansIni = true,
-                     const arma::mat& meansIni = 0, double sigma2_ini = 0.1) {
+                     const arma::mat& meansIni = 0, double sigma2_ini = 0.1,
+                     bool fixSigma2 = false) {
   Mixture mix(y, k, j);
 
   if (j > k) {
     throw std::range_error("j needs be no bigger than k");
   }
 
-  mix.initialize(meansIni, useKmeansIni, fixW, flexJ, zetaTrunc, sigma2_ini);
+  mix.initialize(meansIni, useKmeansIni, fixW, flexJ, zetaTrunc, sigma2_ini,
+                 fixSigma2);
 
   mix.runEM(steps, tol);
+
+  mix.sortBy1stDinMu();
 
   return Rcpp::List::create(
       Rcpp::Named("mu") = mix.mu, Rcpp::Named("w") = mix.w,
@@ -38,7 +42,7 @@ Rcpp::List jkmeansEMBatch(const arma::cube& y, int k, int j, int steps = 1000,
                           bool flexJ = false, double zetaTrunc = 0.01,
                           bool useKmeansIni = true,
                           const arma::mat& meansIni = 0,
-                          double sigma2_ini = 0.1) {
+                          double sigma2_ini = 0.1, bool fixSigma2 = false) {
   if (j > k) {
     throw std::range_error("j needs be no bigger than k");
   }
@@ -57,8 +61,11 @@ Rcpp::List jkmeansEMBatch(const arma::cube& y, int k, int j, int steps = 1000,
   for (int i = 0; i < batchN; ++i) {
     mat localY = y.slice(i);
     Mixture mix(localY, k, j);
-    mix.initialize(meansIni, useKmeansIni, fixW, flexJ, zetaTrunc, sigma2_ini);
+    mix.initialize(meansIni, useKmeansIni, fixW, flexJ, zetaTrunc, sigma2_ini,
+                   fixSigma2);
     mix.runEM(steps, tol);
+
+    mix.sortBy1stDinMu();
 
     mu.slice(i) = mix.mu;
     w.col(i) = mix.w;

### 2016-10-30 jkmeans/src/mixture.hpp

@@ -13,10 +13,8 @@ class Mixture {
   vec w;
   mat zeta;
 
-  vec gradient;
-  mat hessian;
-
   bool fixW;
+  bool fixSigma2;
   bool flexJ;
   double zetaTrunc;
 
@@ -34,13 +32,12 @@ class Mixture {
     mu.randn();
 
     zeta = ones<mat>(n, K);
-
-    gradient = zeros(p * K);
-    hessian = zeros<mat>(p * K, p * K);
   }
 
   void initialize(mat meansInput, bool useKmeansIni, bool _fixW, bool _flexJ,
-                  double _zetaTrunc, double sigma2_ini) {
+                  double _zetaTrunc, double sigma2_ini, bool _fixSigma2) {
+    sigma2 = sigma2_ini;
+
     if (useKmeansIni) {
       mat means;
       bool status = kmeans(means, trans(y), K, static_spread, 10, false);
@@ -50,15 +47,18 @@ class Mixture {
       int _J = J;  // back up the user defined J
       J = 1;
       Expectation();
-      updateSigma2();
+
+      if (!fixSigma2) {
+        updateSigma2();
+      }
 
       J = _J;
 
     } else {
       mu = meansInput;
-      sigma2 = sigma2_ini;
     }
 
+    fixSigma2 = _fixSigma2;
     fixW = _fixW;
     flexJ = _flexJ;
     zetaTrunc = _zetaTrunc;
@@ -142,55 +142,13 @@ class Mixture {
       diff2 += accu(sum(diff % diff, 1) % weights);
       sum_total_weights += total_weights * p;
 
-      // sigma2 = (accu(sum(diff % diff, 1) % weights)) /
-      // (total_weights * p + 0.1);
-
       // compute w
       if (!fixW) w(k) = accu(weights) / (double)n;
     }
 
-    sigma2 = (diff2 / sum_total_weights);
-  }
-
-  vec computeGradient(mat mu) {
-    vec gradient(p * K);
-
-    for (int k = 0; k < K; ++k) {
-      mat diff = y - repmat(mu.row(k), n, 1);
-
-      vec zeta_local = zeta.col(k);
-      vec gradient_local = diff.t() * zeta_local / sigma2;
-
-      double hessian_diag_local =
-          -accu(zeta_local) / sigma2;  // same for all p elements
-
-      int idx = k * p;
-      int end_idx = idx + p - 1;
[…]

### 2016-10-30 jkmeans/test/test2.r

@@ -1,0 +1,22 @@
+require("jkmeans")
+
+mu0<- c(1,2)
+n<- 40
+mu<- c(rep(1,n),rep(2,n*2))
+y<- rnorm(n*3, mu)
+
+fit<- jkmeansEM(as.matrix(y),k = 2, j=1, meansIni = as.matrix(mu0),steps = 1000,tol = 1E-15, fixW = F, fixSigma2 = F, sigma2_ini = 1,useKmeansIni = T)
+
+fit$mu
+fit$w
+fit$sigma2
+
+
+fit2<- jkmeansEM(as.matrix(y),k = 2, j=2, meansIni = as.matrix(mu0),steps = 1000,tol = 1E-15, fixW = F, fixSigma2 = F, sigma2_ini = 0.01,useKmeansIni = T)
+
+fit$mu
+fit$w
+fit$sigma2
+
+require("MCMCpack")
+rdirichlet(1000, rep(1,5))

### 2016-11-07 Rmd/HighP.Rmd
updated write-up, test with large p

# .rs.restartR()
```{r include=FALSE}

require("jkmeans")

n<- 100
K<- 2
p<- 100

sigma<- 1
mu1<- rep(K,n*K)

randomMu<- unlist(sapply(1:K, function(x){rep(x, runif(1,n/10,n))}))
mu1[1:length(randomMu)]<- randomMu

mu<- matrix(0,n*K,p)
mu[,1:5]<- mu1
# mu[,2]<- mu1

y<- matrix( rnorm(n*K*p, mu,sd = (sigma)), n*K)
# hist(y,breaks = 100)

svdY<- svd(y)

plot(svdY$d)

y1<- as.matrix( t(t(svdY$u)*svdY$d) [,c(1)])

# hist(c(sapply(c(1:K),function(x){ rnorm(1000,x, (sigma2/sqrt(n)))})), breaks = 1000)

kmpp <- function(X, k) {
  n <- nrow(X)
  C <- numeric(k)
  C[1] <- sample(1:n, 1)

  for (i in 2:k) {
    dm <- abs(outer(X, X[C, ], "-"))
    pr <- apply(dm, 1, min)
    pr[C] <- 0
    C[i] <- sample(1:n, 1, prob = pr)
  }

   X[C, ]
}

rmse<- function(x){
  v<- x$mu[,1:5]
  v<- v[order(v[,1]),]
  sqrt(mean((v-c(1:2))^2))
}

plot(y,col=mu1)

k<- K

ini<- as.matrix(kmpp(y,k))

kmeans <- jkmeansEM(y,k = k,j= 1,1000,tol = 1E-15,fixW = T, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
# table(kmeans$M)
# lines(kmeans$mu, type="p",pch=2,lwd=3,col="green")

jk<- jkmeansEM(y,k = k,j= 1,1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
# table(jk$M)
# lines(jk$mu, type="p",pch=2,lwd=3,col="blue")
jk$mu[,1]

full<- jkmeansEM(y,k = k,j= k,1000,tol = 1E-10,fixW = F, meansIni = ini,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
# table(full$M)
# lines(full$mu, type="p",pch=3,lwd=3,col="red")
full$mu[,1]

y<- y1
ini<- as.matrix(kmpp(y,k))

kmeans_pca <- jkmeansEM(y,k = k,j= 1,1000,tol = 1E-15,fixW = T, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
# table(kmeans$M)
# lines(kmeans$mu, type="p",pch=2,lwd=3,col="green")
# kmeans$mu[,1]

jk_pca<- jkmeansEM(y,k = k,j= 1,1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
# table(jk$M)
# lines(jk$mu, type="p",pch=2,lwd=3,col="blue")
# jk$mu[,1]

full_pca<- jkmeansEM(y,k = k,j= k,1000,tol = 1E-10,fixW = F, meansIni = ini,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
# table(full$M)
# lines(full$mu, type="p",pch=3,lwd=3,col="red")
# full$mu[,1]

((kmeans$mu[1,1]*svdY$v)[,1]-(kmeans$mu[2,1]*svdY$v)[,1])[1:5]
((jk$mu[1,1]*svdY$v)[,1]-(jk$mu[2,1]*svdY$v)[,1])[1:5]
((full$mu[1,1]*svdY$v)[,1] - (full$mu[2,1]*svdY$v)[,1])[1:5]

rmse_pca<- function(x){
  v<-rbind( (x$mu[1,1]*svdY$v)[1:5,1],  (x$mu[2,1]*svdY$v)[1:5,1])
  v<- v[order(v[,1]),]

  sqrt(mean((v-c(1:2))^2))
}

kmeans_pca$mu 
dim(svdY$v)

```

```{r}
rmse(kmeans)
rmse(jk)
rmse(full)
rmse_pca(kmeans_pca)
rmse_pca(jk_pca)
rmse_pca(full_pca)
```

### 2016-11-07 docs/ver1.tex
updated write-up, test with large p

--- 2016-10-10 docs/ver1.tex
@@
 
+\usepackage{mhequ}
+\newcommand{\be}{\begin{equs}}
+\newcommand{\ee}{\end{equs}}
+\newcommand{\bb}[1]{\mathbb{#1}}
+\newcommand{\mc}[1]{\mathcal{#1}}
+
 \thispagestyle{empty}
@@
 \title
-    {?? Mixture Model}
+    {Complete Membership Model for High Dimensional Clustering}
 
@@
 
-hard clustering:
-several overly strong assumptions in k-means
+Literature review:
 
-soft clustering:
-large variance induced by way too many non-zeros in the probability simplex.
+Finite mixture model clustering.
 
-variance-bias tradeoff?
-reducing variability on the weight
+Difficulty in high dimensional clustering.
 
-\section{?? Mixture Model}
+The difficulty in Determinantal process clustering, repulsive clustering.
+
+Classification EM.
+
+\section{Complete Membership Model}
 
@@
 \begin{aligned}
-& \pi(y_i) \sim G(\theta_i) \\
-& \pi(\theta_i) = \sum_{k=1}^{\kappa} w_{k,i} \delta_{\theta_{k}}(\theta_i) \\
-& w_{k,i} = \frac{ u_{k,i} v_{k} } { \sum_{k=1}^{\kappa}  u_{k,i} v_{k} }\\
-& u_{k,i} \in \{0,1\}
+& \pi(y_i) \stackrel{indep}{\sim} F(y_i|\theta_i) \\
+& \pi(\theta_i) = \sum_{k=1}^{\kappa} z_{k,i} \delta_{\theta^*_{k}}(\theta_i) \\
 \end{aligned}
-\label{marginal_lik}
+\label{membership_model}
 \end{equation}
-where $\delta_{\theta_{k}}(y_i)$ is a probabilty density corresponds to a component distribution with parameter $\theta_{k}$, and $w_{k,i}$ is the component weight that varies by index $i$ and $\sum_{k=1}^{\kappa} w_{k,i} =1$.
-The weight varying is due to the randomness in $u_{k,i}$.
-Due to the $0$'s $u_{k,i}$ introduces, this induces sparsity in the membership probabilty $w_{k,i}$.
+for each $i$, $\{ z_{.,i}\}$ denotes the membership with only one $1$ and $(k-1)$ many $0$'s.
+Note $\{ z_{.,i}\}$ is not random for each $i$, although collectively $\{ \sum_i  z_{1,i}, \sum_i  z_{2,i},\ldots \sum_i  z_{\kappa,i} \}$ follows a multinomial distribution $Multinomial(n, \{ w_1, w_2, \ldots, w_\kappa \})$.
 
@@
 
-We now use the latent variable $\{ z_{1,i},z_{2,i},\ldots z_{\kappa,i} \} \sim MultiNomial (\{ w_{1,i},w_{2,i},\ldots w_{\kappa,i}\} )$ that takes value of $\{0,\ldots,1,\ldots 0\}$ that assign randomly one $1$ to one of $\kappa$ vertices in the simplex.
-The likelihood can be rewritten as:
-
 \begin{equation}
 \begin{aligned}
-& [ y_i ] =   \sum_{k=1}^{\kappa} \int z_{k,i} \delta_{\theta_{k}}(y_i) P(d z_{k,i}) = \int \sum_{k=1}^{\kappa} z_{k,i} \delta_{\theta_{k}}(y_i) P(d z_{k,i})
-\end{aligned}
-\end{equation}
-where $P(d z_{k,i})$ is the measure of the multinomial distribution aforementioned, the last equation is due to Fubini theorem.
-Note due to the $z_{k,i}$ takes only one $1$ out of $K$, this allows a simple log-density augmented with $z_{k,i}$:
-
-\begin{equation}
-\begin{aligned}
-& \log[ \{y_i\}_i, \{z_{k,i}\}_{k,i} ] =  \sum_{i}\sum_{k=1}^{\kappa} z_{k,i} \log \{\delta_{\theta_{k}}(y_i) v_k \}
+& \log[ \{y_i\}_i, \{z_{k,i}\}_{k,i} ] =  \sum_{i}\sum_{k=1}^{\kappa} z_{k,i} \log \{w_k F(y_i\mid \theta_k)  \}
 \end{aligned}
@@
 \begin{algorithm}[H]
-\caption{EM algorithm}\label{alg:euclid}
+\caption{Estimation algorithm}\label{alg:euclid}
 \begin{algorithmic}[1]
-% \Procedure{Euclid}{$a,b$}\Comment{The g.c.d. of a and b}
-% \State $r\gets a\bmod b$
 \While{$||\theta^{(s)}-\theta^{(s+1)}||>\epsilon$}%\Comment{We have the answer if r is 0}
-\State Maximize over $\{ u_{i,k}\}_k$ by $ \underset{\{u_{i,k}\}_k} {\text{argmax }}\{u_{i,k}  v_{k}\delta_{\theta_{k}}(y_i)\}$ for all $i$; \Comment{\small pick top J out K based on (\ref{marginal_lik})}
-\State Compute $ \mathbb{E}(z_{k,i})= \frac{w_{k,i} \delta_{\theta_{k}}(y_i)}{\sum_k w_{k,i} \delta_{\theta_{k}}(y_i)}$ with $w_{k,i} = u_{i,k} v_{k}$; \Comment{E step  based on (\ref{conditional_lik})}
-\State Set $\theta^{(s+1)}_{k}= \text{argmax} \sum_{k=1}^{\kappa} \mathbb{E}z_{k,i} \log \delta_{\theta_{k}}(y_i) $ for all $k$ \Comment{M step: obtain MLE based on (\ref{conditional_lik})}
-\State Set $v^{(s+1)}_{k}=  \sum_{i}\mathbb{E}z_{k,i} /  \sum_{i}\sum_{k}\mathbb{E}z_{k,i}$ \Comment{M step: obtain MLE based on (\ref{conditional_lik})}\label{flex_w_step}
+\State Maximize over $\{ z_{i,k}\}_k$ by $ \underset{\{z_{i,k}\}_k} {\text{argmax }}\{z_{i,k}  w_{k}F (y_i|\theta^*_k)\}$ for all $i$;
+\State Update $\theta^{*}_{k}= \text{argmax} \sum_{i=1}^{n} z_{k,i} \log F(y_i|\theta^*_k) $ for all $k$ 
+\State Update $w_{k}=  \sum_{i} z_{k,i} /  \sum_{i}\sum_{k}z_{k,i}$ 
 \EndWhile
-% \State \textbf{return} $b$\Comment{The gcd is b}
-% \EndProcedure
 \end{algorithmic}
@@
 
+\subsection{High Dimensional Estimation}
+
+Alternative Least Square with regularization
+
 \section{Theory}
 
-variance reduction in the weight parameter
+\subsection{Increased Separation among the Centers}
+
+Consider the finite mixture model with the center estimates as $\mu_k = \frac{\sum_i \bb E z_{k,i} y_i}{\sum_i \bb E z_{k,i}}$, and complete membership model with the center estimates as $\mu^*_k = \frac{\sum_i  z_{k,i} y_i}{\sum_i  z_{k,i}}$.
+We are interested in comparing the pairwise distance among the centers from the two models.
+
+Let the pairwise distance be $||\mu_1 - \mu_2||$ between two centers in the finite mixture.
+As the $||\mu_1 - \mu_2||= \sqrt{ \sum_{l=1}^{p} ||\mu_{1,l} - \mu_{2,l}||^2}$, we focus on one sub-dimension $\mu_{1,l} - \mu_{2,l}$, without loss of generality, we assume $\mu_{1,l} > \mu_{2,l}$.
+
+For any $y_{j,l} \ge  \frac{\sum_{i\ne j} \bb E z_{1,i} y_{i,l} }{  \sum_{i\ne j} E z_{1,i} } \ge \frac{\sum_{i\ne j} \bb E z_{2,i} y_{i,l} }{  \sum_{i\ne j} E z_{2,i} } $ and $ \bb E z_{1,j} \ge \bb E z_{2,j}$,
+\be
+\mu_{1,l} - \mu_{2,l} = & \frac{\sum_{i\ne j} \bb E z_{1,i} y_{i,l}  + \bb E z_{1,j} y_{j,l}}{  \sum_{i\ne j} \bb E z_{1,i}  + \bb E z_{1,j} }  -  \frac{\sum_{i\ne j} \bb E z_{2,i} y_{i,l}  + \bb E z_{2,j} y_{j,l}}{  \sum_{i\ne j} \bb E z_{2,i}  + \bb E z_{2,j} } \\
+ \le  & \frac{\sum_{i\ne j} \bb E z_{1,i} y_{i,l}  +  y_{j,l}}{  \sum_{i\ne j} \bb E z_{1,i}  + 1 }  -  \frac{\sum_{i\ne j} \bb E z_{2,i} y_{i,l} }{  \sum_{i\ne j} \bb E z_{2,i}  }
+\ee
+
+For any $y_{j,l} \le   \frac{\sum_{i\ne j} \bb E z_{2,i} y_{i,l} }{  \sum_{i\ne j} E z_{2,i} } \le  \frac{\sum_{i\ne j} \bb E z_{1,i} y_{i,l} }{  \sum_{i\ne j} E z_{1,i} }$ and $ \bb E z_{1,j} \le \bb E z_{2,j}$,
+
+\be
+\mu_{1,l} - \mu_{2,l}  \le  & \frac{\sum_{i\ne j} \bb E z_{1,i} y_{i,l} }{  \sum_{i\ne j} \bb E z_{1,i}  }  -  \frac{\sum_{i\ne j} \bb E z_{2,i} y_{i,l} +  y_{j,l} }{  \sum_{i\ne j} \bb E z_{2,i}+1  }
+\ee
+
+ By induction, this converts all the $E z_{k,i}$ to $z_{k,i}$ hence $\mu_{1,l} - \mu_{2,l}\le \mu^{*}_{1,l} - \mu^{*}_{2,l}$.
+
+\subsection{Convex Relaxation}
+
+Something similar to:
+
+Agarwal, Alekh, Sahand Negahban, and Martin J. Wainwright. "Noisy matrix decomposition via convex relaxation: Optimal rates in high dimensions." The Annals of Statistics (2012): 1171-1197.
 
@@
 
-large K with small J in 1d
+Note: preliminary
 
-high p setting
+RMSE:
+
+\begin{table}[H]
+\centering
+	\begin{tabular}{| l | r | r |}
+	\hline
+			Model & $n=100, p=100, p^*=5$ &  $n=100, p=100, p^*=100$\\
+	\hline
+		k-means &  0.40 & 0.069\\
+		CM		& 0.43 & 0.069\\
+		GMM &  0.46 & 0.069 \\
+		PCA+ k-means & 0.17 & 0.030\\
+		PCA + CM		& 0.07 & 0.030\\
+		PCA + GMM & 0.10 & 0.030\\
+		new model  & \\
+		new model &\\
+		new model &\\
+			\hline
+	\end{tabular}
+\end{table}
 

### 2016-11-07 Rmd/ExcessK.r

@@ -1,0 +1,37 @@
+require("jkmeans")
+
+n<- 100
+K<- 5
+
+sigma<- 1
+mu<- rep(c(1:K),each=n)
+y<- matrix( rnorm(n*K, mu,sd = (sigma)))
+hist(y,breaks = 100)
+
+hist(c(sapply(c(1:K),function(x){ rnorm(1000,x, (sigma2/sqrt(n)))})), breaks = 1000)
+
+
+k<- K
+
+jk<- jkmeansEM(y,k = k,j= 1,1000,tol = 1E-10,fixW = F, meansIni =  matrix(c(c(1:k))),useKmeansIni = F,sigma2_ini = 0.1)
+
+full<- jkmeansEM(y,k = k,j= k,1000,tol = 1E-10,fixW = F, meansIni =  matrix(c(c(1:k))),useKmeansIni = F,sigma2_ini = 0.1)
+
+oracle <- jkmeansEM(y,k = K,j= K,1000,tol = 1E-10,fixW = F, meansIni =  matrix(c(c(1:K))),useKmeansIni = F,sigma2_ini = 0.1)
+
+flex <- jkmeansEM(y,k = k ,j= k,1000,tol = 1E-10,fixW = F, meansIni =  matrix(c(c(1:k))),useKmeansIni = F,sigma2_ini = 0.1,flexJ = T,zetaTrunc = 0.1)
+
+jk<- jkmeansEM(y,k = k,j= 3,1000,tol = 1E-10,fixW = T, meansIni =  matrix(c(c(1:k))),useKmeansIni = F,sigma2_ini = 1)
+
+
+table(jk$M)
+table(full$M)
+table(oracle$M)
+table(flex$M)
+
+
+jk$mu[jk$w>0.01]
+full$mu[full$w>0.05]
+oracle$mu[oracle$w>0.05]
+flex$mu[flex$w>0.01]
+

### 2016-11-07 Rmd/HighP.R

@@ -1,0 +1,122 @@
+# .rs.restartR()
+require("jkmeans")
+
+n<- 100
+K<- 2
+p<- 100
+
+sigma<- 1
+mu1<- rep(K,n*K)
+
+randomMu<- unlist(sapply(1:K, function(x){rep(x, runif(1,n/10,n))}))
+mu1[1:length(randomMu)]<- randomMu
+
+mu<- matrix(0,n*K,p)
+mu[,1:5]<- mu1
+# mu[,2]<- mu1
+
+
+y<- matrix( rnorm(n*K*p, mu,sd = (sigma)), n*K)
+# hist(y,breaks = 100)
+
+svdY<- svd(y)
+
+plot(svdY$d)
+
+y1<- as.matrix( t(t(svdY$u)*svdY$d) [,c(1)])
+
+# hist(c(sapply(c(1:K),function(x){ rnorm(1000,x, (sigma2/sqrt(n)))})), breaks = 1000)
+
+
+kmpp <- function(X, k) {
+  n <- nrow(X)
+  C <- numeric(k)
+  C[1] <- sample(1:n, 1)
+  
+  for (i in 2:k) {
+    dm <- abs(outer(X, X[C, ], "-"))
+    pr <- apply(dm, 1, min)
+    pr[C] <- 0
+    C[i] <- sample(1:n, 1, prob = pr)
+  }
+  
+   X[C, ]
+}
+
+
+rmse<- function(x){
+  sqrt(mean((x$mu[,1:5]-c(1:2))^2))
+}
+
+
+
+plot(y,col=mu1)
+
+k<- K
+
+ini<- as.matrix(kmpp(y,k))
+
+
+
+kmeans <- jkmeansEM(y,k = k,j= 1,1000,tol = 1E-15,fixW = T, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
+# table(kmeans$M)
+# lines(kmeans$mu, type="p",pch=2,lwd=3,col="green")
+
+
+jk<- jkmeansEM(y,k = k,j= 1,1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
+# table(jk$M)
+# lines(jk$mu, type="p",pch=2,lwd=3,col="blue")
+jk$mu[,1]
+
+
[…]

### 2016-11-07 Rmd/computeBV.r

@@ -136,7 +136,7 @@ df$J<- as.factor(df$J)
 
 pdf("BVMeanDist.pdf",6,3)
 p<- ggplot(data=df, aes())
-p+  geom_histogram(aes(x=value,fill=mu),bins=100, alpha = 0.8)+facet_wrap(~J)
+p+  geom_histogram(aes(x=value,fill=mu),bins=100, alpha = 0.8,position = 'identity')+facet_wrap(~J)
 dev.off()
 
 

### 2016-11-07 Rmd/postSim.r

@@ -3,10 +3,13 @@ require("ggplot2")
 setwd("~/git/jkmeans/Rmd/")
 
 #good
-load("sims/raw_result_2_0_0.5_1_0.Rda")
+# load("sims/raw_result_2_0_0.5_1_0.Rda")
+
+load("sims/raw_result_5_0_1_1_0.Rda")
+
 
 nseries<- length(raw_test_results)
-K<- 2
+K<- 5
 
 #likelihood
 
@@ -17,7 +20,7 @@ loglik_GMM<- function(y, mu,sigma2,w){
 
 
 #get Bias Variance for J=j in GMM
-getBVMean <- function(j){
+getBVMean <- function(j,cutoff=0.05){
   #histogram of the mean
   
   muHat<- raw_n$GMMList[[j]]$mu
@@ -25,7 +28,6 @@ getBVMean <- function(j){
   
   sigma2Hat<- raw_n$GMMList[[j]]$sigma2
   
-  cutoff<- 0.05
   exclude<- (wHat[1,]>(1-cutoff) | wHat[1,]<cutoff)
   
   muHat<- muHat[,,!exclude]
@@ -80,10 +82,11 @@ getBVMean <- function(j){
 }
 
 
-n<-1
+n<-5
 raw_n<- raw_test_results[[n]]
 
-hist(raw_n$M)
+table(raw_n$M)
+
 
 batchN<- dim(raw_n$yBatch)[3]
 yBatch <- raw_n$yBatch
@@ -93,9 +96,15 @@ sliceMeans <- function(x){
 }
 
 
-getBVMean(1)
+raw_n$GMMList[[4]]$zeta
+
+
+getBVMean(1, 0.1)
 getBVMean(2)
 
+sum(getBVMean(1, 0.2)$MSE)
+sum(getBVMean(2, 0.2)$MSE)
+sum(getBVMean(3, 0.2)$MSE)
 
 
 

### 2016-11-07 Rmd/simKmeans.R

@@ -1,0 +1,73 @@
+
+b<- 3
+
+plot(rbind(c(0,0),c(1,b)),type="l",xlim = c(0,5),ylim=c(0,5))
+
+lines(rbind(c(0, b/2+1/2/b),c(b^2/2+1/2,0)),type="l")
+
+
+b<- 1
+rho<- 5
+
+
+kemans_test <-  function(b,rho){
+  
+  mu1<- c(0,0)
+  mu2<- c(1,b)
+  var<- c(1/40,1/40*rho)* 2
+  y1<- t( sqrt(var) *matrix(rnorm(1000*2),nrow=2))
+  y2<-  t( sqrt(var) *matrix(rnorm(1000*2),nrow=2)+ mu2)
+  
+  # plot(y1,xlim = c(-b*2,b*2),ylim=c(-b*2,b*2))
+  # lines(y2, type="p",col="red")
+  
+  y<- rbind(y1,y2)
+  
+  plot(y,col=rep(c(1,2),each=1000), xlim= range(y[,2]), ylim=range(y[,2]))
+  
+  abline(a= b/2+1/2/b, b= -1/b)
+  
+  
+  
+  
+  kmpp <- function(X, k) {
+    n <- nrow(X)
+    C <- numeric(k)
+    C[1] <- sample(1:n, 1)
+    
+    for (i in 2:k) {
+      dm <- abs(outer(X, X[C, ], "-"))
+      pr <- apply(dm, 1, min)
+      pr[C] <- 0
+      C[i] <- sample(1:n, 1, prob = pr)
+    }
+    
+    X[C, ]
+  }
+  
+  ini<- kmpp(y,2)
+  ini<- ini[order(ini[,1]),]
+  
+  # km<- kmeans(y, centers= ini)
+  km<- kmeans(y, centers= rbind(mu1,mu2),iter.max=10)
+  
+  lines(km$centers,type = "p",pch=2,lwd=5,col="blue")
+  
+  
+  rate <-sum(km$cluster != rep(c(1,2),each=1000))/2000
+  print(min(rate,1-rate))
+  
+}
+
+kemans_test(1,5)
+kemans_test(1,10)
+kemans_test(1,20)
+kemans_test(1,30)
+kemans_test(1,40)
+kemans_test(1,100)
+kemans_test(1,1000)
+
+
+kemans_test(2,5)
+kemans_test(2,10)
+kemans_test(5,1000)

### 2016-11-07 jkmeans/R/RcppExports.R

@@ -1,11 +1,11 @@
 # This file was generated by Rcpp::compileAttributes
 # Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
 
-jkmeansEM <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, flexJ = FALSE, zetaTrunc = 0.01, useKmeansIni = TRUE, meansIni = 0L, sigma2_ini = 0.1, fixSigma2 = FALSE) {
-    .Call('jkmeans_jkmeansEM', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini, fixSigma2)
+jkmeansEM <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, flexJ = FALSE, zetaTrunc = 0.01, useKmeansIni = TRUE, meansIni = 0L, sigma2_ini = 0.1, normalizeZeta = FALSE) {
+    .Call('jkmeans_jkmeansEM', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini, normalizeZeta)
 }
 
-jkmeansEMBatch <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, flexJ = FALSE, zetaTrunc = 0.01, useKmeansIni = TRUE, meansIni = 0L, sigma2_ini = 0.1, fixSigma2 = FALSE) {
-    .Call('jkmeans_jkmeansEMBatch', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini, fixSigma2)
+jkmeansEMBatch <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, flexJ = FALSE, zetaTrunc = 0.01, useKmeansIni = TRUE, meansIni = 0L, sigma2_ini = 0.1, normalizeZeta = FALSE) {
+    .Call('jkmeans_jkmeansEMBatch', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini, normalizeZeta)
 }
 

### 2016-11-07 jkmeans/src/RcppExports.cpp

@@ -7,8 +7,8 @@
 using namespace Rcpp;
 
 // jkmeansEM
-Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps, double tol, bool fixW, bool flexJ, double zetaTrunc, bool useKmeansIni, const arma::mat& meansIni, double sigma2_ini, bool fixSigma2);
-RcppExport SEXP jkmeans_jkmeansEM(SEXP ySEXP, SEXP kSEXP, SEXP jSEXP, SEXP stepsSEXP, SEXP tolSEXP, SEXP fixWSEXP, SEXP flexJSEXP, SEXP zetaTruncSEXP, SEXP useKmeansIniSEXP, SEXP meansIniSEXP, SEXP sigma2_iniSEXP, SEXP fixSigma2SEXP) {
+Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps, double tol, bool fixW, bool flexJ, double zetaTrunc, bool useKmeansIni, const arma::mat& meansIni, double sigma2_ini, bool normalizeZeta);
+RcppExport SEXP jkmeans_jkmeansEM(SEXP ySEXP, SEXP kSEXP, SEXP jSEXP, SEXP stepsSEXP, SEXP tolSEXP, SEXP fixWSEXP, SEXP flexJSEXP, SEXP zetaTruncSEXP, SEXP useKmeansIniSEXP, SEXP meansIniSEXP, SEXP sigma2_iniSEXP, SEXP normalizeZetaSEXP) {
 BEGIN_RCPP
     Rcpp::RObject __result;
     Rcpp::RNGScope __rngScope;
@@ -23,14 +23,14 @@ BEGIN_RCPP
     Rcpp::traits::input_parameter< bool >::type useKmeansIni(useKmeansIniSEXP);
     Rcpp::traits::input_parameter< const arma::mat& >::type meansIni(meansIniSEXP);
     Rcpp::traits::input_parameter< double >::type sigma2_ini(sigma2_iniSEXP);
-    Rcpp::traits::input_parameter< bool >::type fixSigma2(fixSigma2SEXP);
-    __result = Rcpp::wrap(jkmeansEM(y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini, fixSigma2));
+    Rcpp::traits::input_parameter< bool >::type normalizeZeta(normalizeZetaSEXP);
[…]

### 2016-11-07 jkmeans/src/jkmeans.cpp

@@ -16,7 +16,7 @@ Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps = 1000,
                      double tol = 1E-8, bool fixW = true, bool flexJ = false,
                      double zetaTrunc = 0.01, bool useKmeansIni = true,
                      const arma::mat& meansIni = 0, double sigma2_ini = 0.1,
-                     bool fixSigma2 = false) {
+                     bool normalizeZeta = false) {
   Mixture mix(y, k, j);
 
   if (j > k) {
@@ -24,7 +24,7 @@ Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps = 1000,
   }
 
   mix.initialize(meansIni, useKmeansIni, fixW, flexJ, zetaTrunc, sigma2_ini,
-                 fixSigma2);
+                 normalizeZeta);
 
   mix.runEM(steps, tol);
 
@@ -42,7 +42,7 @@ Rcpp::List jkmeansEMBatch(const arma::cube& y, int k, int j, int steps = 1000,
                           bool flexJ = false, double zetaTrunc = 0.01,
                           bool useKmeansIni = true,
                           const arma::mat& meansIni = 0,
-                          double sigma2_ini = 0.1, bool fixSigma2 = false) {
+                          double sigma2_ini = 0.1, bool normalizeZeta = false) {
   if (j > k) {
     throw std::range_error("j needs be no bigger than k");
   }
@@ -62,7 +62,7 @@ Rcpp::List jkmeansEMBatch(const arma::cube& y, int k, int j, int steps = 1000,
     mat localY = y.slice(i);
     Mixture mix(localY, k, j);
     mix.initialize(meansIni, useKmeansIni, fixW, flexJ, zetaTrunc, sigma2_ini,
-                   fixSigma2);
+                   normalizeZeta);
     mix.runEM(steps, tol);
 
     mix.sortBy1stDinMu();

### 2016-11-07 jkmeans/src/mixture.hpp

@@ -14,7 +14,7 @@ class Mixture {
   mat zeta;
 
   bool fixW;
-  bool fixSigma2;
+  bool normalizeZeta;
   bool flexJ;
   double zetaTrunc;
 
@@ -35,7 +35,7 @@ class Mixture {
   }
 
   void initialize(mat meansInput, bool useKmeansIni, bool _fixW, bool _flexJ,
-                  double _zetaTrunc, double sigma2_ini, bool _fixSigma2) {
+                  double _zetaTrunc, double sigma2_ini, bool _normalizeZeta) {
     sigma2 = sigma2_ini;
 
     if (useKmeansIni) {
@@ -48,9 +48,7 @@ class Mixture {
       J = 1;
       Expectation();
 
-      if (!fixSigma2) {
-        updateSigma2();
-      }
+      updateSigma2();
 
       J = _J;
 
@@ -58,7 +56,7 @@ class Mixture {
       mu = meansInput;
     }
 
-    fixSigma2 = _fixSigma2;
+    normalizeZeta = _normalizeZeta;
     fixW = _fixW;
     flexJ = _flexJ;
     zetaTrunc = _zetaTrunc;
@@ -110,6 +108,8 @@ class Mixture {
 
         local_zeta = exp(local_zeta);
 
+        local_zeta = local_zeta / accu(local_zeta);
+
         local_zeta(find(local_zeta < local_zeta(trunc_idx))).fill(0);
       }
       if (flexJ) {
@@ -117,8 +117,10 @@ class Mixture {
         local_zeta /= accu(local_zeta);
         local_zeta(find(local_zeta < zetaTrunc)).fill(0);
       }
-
-      zeta.row(i) = local_zeta / accu(local_zeta);
+      if (normalizeZeta)
+        zeta.row(i) = local_zeta / accu(local_zeta);
+      else
+        zeta.row(i) = local_zeta;
     }
   }
 
@@ -146,9 +148,7 @@ class Mixture {
       if (!fixW) w(k) = accu(weights) / (double)n;
     }
 
-    if (!fixSigma2) {
-      sigma2 = (diff2 / sum_total_weights);
-    }
+    sigma2 = (diff2 / sum_total_weights);
   }
 
   void updateSigma2() {
@@ -171,6 +171,8 @@ class Mixture {
   void runEM(int steps, double tol) {
     mat mu0 = mu;
 
+    double cur_loglik = -INFINITY;
+
     for (int i = 0; i < steps; ++i) {
[…]

### 2016-11-21 docs/draft.tex
implemented EM algorithm for Dimension-Reduction-Clustering

\documentclass[12pt]{article}
\pdfoutput=1

\addtolength{\oddsidemargin}{-.875in}
\addtolength{\evensidemargin}{-.875in}
\addtolength{\textwidth}{1.75in}

\addtolength{\topmargin}{-.875in}
\addtolength{\textheight}{1.75in}

\openup 1em

%macro for commenting
\usepackage{color}
\newcommand{\leo}[1]{{\color{blue}{\it leo: #1}}}
\newcommand{\xbeta}{\boldsymbol X_i^T \boldsymbol\beta}

\usepackage[round]{natbib}

\usepackage{rotating}
\usepackage{graphicx}
\usepackage{subcaption}

\usepackage{float}

\usepackage{amsthm,amsmath} 
\usepackage{amssymb}
\usepackage{subcaption}
\captionsetup{compatibility=false}

\usepackage[utf8]{inputenc}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}

\usepackage{mhequ}
\newcommand{\be}{\begin{equs}}
\newcommand{\ee}{\end{equs}}
\newcommand{\bb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\bl}{\boldsymbol}
\newcommand{\X}{\boldsymbol X}
\newcommand{\Y}{\boldsymbol Y}
\newcommand{\Z}{\boldsymbol Z}
\newcommand{\W}{\boldsymbol W}
\newcommand{\I}{\boldsymbol I}
\newcommand{\V}{\boldsymbol V}
\newcommand{\bmu}{\boldsymbol \mu}
\newcommand{\bSigma}{\boldsymbol \Sigma}
\newcommand{\E}{\bb E}

\thispagestyle{empty}
\baselineskip=28pt

\title
    {Automatic Repulsive Clustering \\for Finding Separable Linear Subspace}

\author{
     Leo L. Duan and
     Joshua T. Vogelstein
    % \textsuperscript{*}\footnotemark[2]\and
}

\date{}

\begin{document}

\maketitle

\begin{abstract}
Mixture model based clustering methods are routinely used to divide the heterogeneous data into groups, within each the data are similar and characterized by a center. At the same time, it is often desired to maintain significant difference across groups so that the data can be clearly separated. Repulsive regularization among cluster centers serve this purpose, however, they suffer from the curse of dimensionality and specifying the repulsion parameter inevitably leads to sensitivity issue. In this article, we propose a different but simple regularization by assigning data completely into clusters without randomness, this creates automatic repulsion without need for tuning. This becomes especially useful in clustering with high dimensional data, where a separable linear subspace can be obtained. Simulations illustrate the strengths of the method and substantial gains are demonstrated in an application of clustering synaptomes.\\
{\noindent  KEY WORDS:  Complete Membership, High Dimensional Clustering, Repulsive Regularization.}
\end{abstract}

\section{Introduction}

Model based clustering \citep{fraley2002model} is used extensively in unsupervised learning. The common strategy is to treat the likelihood of each data $y_i$ for $i=1\ldots n$ as a weighted mixture of independent components $L(y_i)=\sum_{k=1}^{K} \pi_k f(y_i|\theta_k)$, where $\pi_k$ is the weight and $\theta_k$ is the parameter for the $k$th component. The standard optimization procedure introduces an important latent variable $z_i$, that assigns each data to a component as a one-out-of-$K$ random draw, coupling with an expectation-maximization (EM) (cite Dempster) algorithm to estimate $\pi_k$ and $\theta_k$. After the algorithm converges, one assigns the data to the most probable choice for $z_i$, dividing the data to $K$ partitions. For multivariate Gaussian data $y_i \in \bb R^p$, the covariance is quite useful to accommodate the different importances in each sub-dimension. For example, large variance on the diagonal could result in significant overlap of components, suggesting the sub-dimension is less importance than the others.

The above method fails in high dimensional data with $p\gg n$. Due to the rank, the $p\times p$ covariance matrix cannot be estimated; even with a $p$-element diagonal matrix, there is still large uncertainty due to the small $n$, leading to poor performance. To solve this problem, it is useful to consider dimension reduction by decomposing the matrix $\bl Y= \bl X \bl V + \bl U$ with $\bl X\in \bb R^{n\times d}$ and $d\ll p$ and then use model-based clustering on $\bl X$. Various matrix factorization methods have been used to obtain the lower dimensional factor $\bl X$, such as principle component analysis (PCA) \citep{liu2003pca} and non-negative matrix factorization \citep{tamayo2007metagene}. A significant drawback, however, is that the top $d$ learned subspaces in $\bl X$ do not guarantee a good separation. For example, PCA generates subspace that maximizes the total variance, but good separation in clustering is related to large between-group variance.

As a remedy, it is possible to re-adjust the orientation of $\bl V$ after clustering is done on $\bl X$. For example, when $\bl X$ is clustered, its mean can be expressed as a product of the latent variable probability $\bl W\in \bb R^{n\times k}$ over $k$ components, and their corresponding $d$-dimensional centers $\bl \mu \in \bb R^{k \times d}$. Then using $\bl{W\mu}$ to replace $\bl X$, one can update the estimate of $\bl V$. Alternating between matrix factorization and clustering aligns the subspace to a certain direction that optimizes the clustering model. However, this re-adjustment alone does not solve the low-separation issue. Indeed, model-based clustering only characterizes the degree of overlap through a mixture framework, but does not enforce good separation among the cluster centers.

Therefore, it is useful to consider regularization to obtain good separation in the reduced dimensional subspace. In this regard, repulsive regularization is useful. Examples include the determinantal point process \citep{Kulesza:2012:DPP:2481023} and repulsive mixture \citep{petralia2012repulsive}. These models show good performance in the original data space. When it comes to the latent low dimensional space, there are several critical issues: the amount of the repulsion is controlled by the hyper-parameter, which is difficult to tune when the outcome is not directly observable, creating sensitivity issue; the computation is costly due to the evaluation of the determinant or pairwise repulsion.

Motivated by these studies, we propose a new repulsive regularization on the component centers in the low dimensional subspace. Instead of directly applying penalty on a short distance, we modify the latent probability matrix $\W$ to a complete membership binary matrix $\hat\Z$, which is learned when maximizing the conditional likelihood. This indirectly creates repulsion among the centers. Then alternating maximization can be utilized to find the subspace where the data are separable. This model is efficient to estimate and requires no tuning, hence we refer it as automatic repulsive clustering.

It is worth mentioning a different class of high dimensional clustering method, namely sparse clustering (see \cite{witten2012framework} and the references therein). The main idea is to select the subset of dimensions directly on the data space via sparsity constraint. Our focus is different, since there are many types of data that do not exhibit significant group pattern on a small dimension subset, but do so on a low dimensional latent space. For example, shape and image data commonly show difference almost everywhere, but the difference can be summarized in a projected low dimensional space. In this scenario, our approach is more suitable.

The article proceeds as follows: in section 2, the modeling framework and the estimation procedure are described; in section 3, theory is provided on the automatic repulsion; in section 4, simulation illustrates the advantages; in section 5, a real data application is demonstrated via synapse clustering.

\section{Automatic Repulsive Clustering in Reduced Dimension}

\subsection{Clustering Under Reduced Dimension}

We first summarize the framework that combines dimension reduction and clustering. We refer this as reduced dimension clustering.

Let $\bl Y\in \bb R^{n\times p}$ be the observed data, then we assume the clustering signal reside in a low dimensional matrix $\bl X \in \bb R^{n\times d}$ with $d\ll p$. The clustering of $\bl X$ can be represented as the assignment probability matrix $\bl W\in \bb R^{n\times k}$ with respect to $k$ components $w_{i,k}= \frac {\pi_k f(x_i|\theta_k)}{\sum_k \pi_k  f(x_i|\theta_k)}$,and the cluster mean $\bl \mu \in \bb R^{k\times d}$ adding a random residual $\bl E \in \bb R^{n\times d}$. Using matrix form:

\be
\bl Y & = \bl {XV + U}
\\     &  = \bl{(W\mu +E)V + U}
\label{low_dim_clustering}
\ee
where $\bl U \in \bb R^{n\times p}$ is a matrix containing noise $U_{ij}\sim N(0, \sigma^2)$ and $\bl E \in \bb R^{n\times d}$ is the Gaussian noise with each row $\bl E'_{i,.}\sim N(\bl 0, \bl \Sigma)$. Common choice for $\bl \Sigma$ includes dense $d\times d$ matrix or simple diagonal matrix.

As a comparison, consider direct clustering on the original space $\bb R^{n\times p}$:
\be
\bl Y & = \bl {W\mu + U}
\ee
where  $\bl \mu \in \bb R^{k\times p}$ and  $\bl U  \in \bb R^{n\times p}$.

The key difference lies in the error structure, in direct clustering, each row $\bl U'_{i,.} \sim   N(\bl 0, \bl\Sigma^*)$ with $\bl \Sigma^*$ as a $p\times p$ matrix, due the large dimension $p$, it is difficult to impose structure in or estimate $\bl \Sigma^*$. In reduced dimension, since the subspace projection $\bl V$ is learned, each row of the error term $\bl E'_{i,.} \bl V+ \bl U'_{i,.} \sim   N(\bl 0,\bl {V' \Sigma V} +\sigma^2 \bl I )$, which is a projection of low rank $d$ matrix to the large $p$ matrix. This low rank structure allows borrowing of strength across $p$ different dimensions and is especially useful when the sample size $n$ is small.

\subsection{Automatic Repulsive Clustering (ARC)}

We now introduce the automatic repulsive regularization. Typically the regularization is applied on $\bmu$ directly, causing sensitivity issue with tuning parameter and computing inconvenience. Instead, we regularize by replacing $\W$ with $\hat \Z$, which is the most probable choice of component for each data under $\hat z_{i,k}= 1 \left( k = {\underset{k}{\mbox{argmax}}{ \pi_k f(x_i|\theta_k)}} \right)$. As each $\mu_k$ is the average of $x_i$ weighted by $w_{i,k}$, this leads to automatic repulsion among them, stated by the following theorem:

\begin{theorem}
For any $k\ne k^*$, 
$|| \frac{\sum_i w_{i,k} x_i}{\sum_i w_{i,k}} - \frac{\sum_i w_{i,k^*} x_i}{\sum_i w_{i,k^*}}||\le ||\frac{\sum_i z_{i,k} x_i}{\sum_i z_{i,k}} - \frac{\sum_i z_{i,k^*} x_i}{\sum_i z_{i,k^*}}||$
\end{theorem}
The proof of this theorem is provided in the appendix.

The interpretation of this regularization is to force the estimate centers $\bmu_k$'s to be far apart, so that each row $\X_i$ has one of assignment probability $w_{i,k}\approx 1$. Compared to the other regularization \citep{Kulesza:2012:DPP:2481023}, this is much simple and tuning free. As we show in the next section, the estimation can be carried out by replacing the expectation step in model-based clustering with a maximization step.

\section{Estimation}

We divide the estimation into two parts: estimate the subspace by updating matrix $\bl X$, give the clustering; use ARC to cluster $\bl X$. The estimation proceeds by alternating between these two steps:

\subsection{Updating the Low Dimensional Subspace}

Given the clustering matrix $\bl Z$ and $\bl \mu$, the log-likelihood function is:

\be
\log L = - \frac{1}{2} \left \{ {\sum_i^n ||\Y_{i,.} - \X_{i,.} \V||^2}/{\sigma^2} + {\sum_i^n (\X_{i,.} - \Z_{i,.} \bmu)' \Sigma^{-1}  (\X_{i,.} - \Z_{i,.} \bmu)} + n\log \det  \bSigma + np \log \sigma^2 \right\}  + C
\ee

To ensure identifiability, we regularize $V_{i,j}\sim N(0,1)$. It is possible to  maximize the log-likelihood alternatively over $\V$ and $\X$, however, this would underestimate the variability of the $\V$, leading to suboptimal result. Instead, we treat $\V$ as latent variable and use EM algorithm for optimization.

Note the conditional distribution for $\V$ is:
\be
\V_{.,j} \stackrel{indep}{\sim} N\left( (\X'\X+\I)^{-1}\X'\Y_{.,j}, \sigma^2 (\X'\X+\I)^{-1}    \right)
\ee
for $j=1\ldots p$.

This leads to computing the expectation:
\be
\E\V =  (\X'\X+\I)^{-1}\X'\Y,  \qquad\E\V \V' = p\sigma^2 (\X'\X+\I)^{-1} + (\X'\X+\I)^{-1}\X'\Y \Y'\X (\X'\X+\I)^{-1}
\ee

Then maximize over $\X$:

\be
\hat \X =   \{ \Y \E\V'/\sigma^2 + \Z\bmu \bSigma^{-1}  \} (\E\V\V'/\sigma^2 + \bSigma^{-1})^{-1}
\ee

and over $\sigma^2$:

\be
\hat \sigma^2 = \{\mbox{vec} (\Y)'\mbox{vec} (\Y)  -2 \mbox{vec}  ( \hat\X')' \mbox{vec}( \E\V \Y') + \mbox{vec}  ( \hat\X')' \mbox{vec}( \E\V\V' {\hat \X}' )\}/ np
\ee
where $\mbox{vec}(.)$ denotes the column-wise vectorization.

\subsection{Clustering}

The clustering can be carried out by alternative maximization over $\hat Z$, $\bmu_k$ and $\bSigma$.

\be
\hat z_{i,k} & = 1 \left( k = {\underset{k}{\mbox{argmax}}{ \pi_k f(x_i|\theta_k)}} \right)\\
\bmu_k & = \frac{\sum_i z_{i,k} \X_i}{\sum_i z_{i,k}}\\
\bSigma & = \frac{\sum_k\sum_i z_{i,k} \X_i\X'_i}{\sum_k\sum_i z_{i,k}}
\ee

Then the whole estimation can be carried out by alternating in the two steps.

\section{Simulation}

Data generation:

n=300
p=2000
d=2
mu=1,2,3
sigma2=0.5

\begin{table}[H]
\centering
	\begin{tabular}{| l | r |}
	\hline
			Model & $ARI$ \\
	\hline
			GMM on Principle Components &  0.01  \\
			ARC on Principle Components		& 0.31 \\
			GMM on Flexible Reduced Dimension &  0.35  \\
			ARC on Flexible Reduced Dimension		& {\bf 0.44} \\
			\hline
	\end{tabular}
\end{table}

\begin{figure}[H]
 % \centering
  \begin{subfigure}[b]{0.49\textwidth}
 \includegraphics[width=0.8\textwidth]{pics/originalY2d.pdf}
  \caption{The 3 component view in the first 2 dimensions.}
\end{subfigure}
  \hfill
   \begin{subfigure}[b]{0.49\textwidth}
 \includegraphics[width=0.8\textwidth]{pics/ld_arc_truth.pdf}
  \caption{The true latent 2d separable subspace view.}
\end{subfigure}

 \begin{subfigure}[b]{0.49\textwidth}
 \includegraphics[width=0.8\textwidth]{pics/pca_gmm.pdf}
  \caption{GMM on principle components}
\end{subfigure}
  \hfill
   \begin{subfigure}[b]{0.49\textwidth}
 \includegraphics[width=0.8\textwidth]{pics/pca_arc.pdf}
  \caption{ARC on principle components}
\end{subfigure}

 \begin{subfigure}[b]{0.49\textwidth}
 \includegraphics[width=0.8\textwidth]{pics/ld_gmm.pdf}
  \caption{GMM on flexible reduced Dimension}
\end{subfigure}
  \hfill
   \begin{subfigure}[b]{0.49\textwidth}
 \includegraphics[width=0.8\textwidth]{pics/ld_arc.pdf}
  \caption{ARC on flexible reduced Dimension}
\end{subfigure}

 \end{figure}

\section{Application}
\section{Discussion}

\bibliography{reference}
\bibliographystyle{plainnat}

\section{Appendix}

\subsection{Proof of Theorem}

Consider the finite mixture model with the center estimates as $\mu_k = \frac{\sum_i \bb E z_{k,i} y_i}{\sum_i \bb E z_{k,i}}$, and complete membership model with the center estimates as $\mu^*_k = \frac{\sum_i  z_{k,i} y_i}{\sum_i  z_{k,i}}$. We are interested in comparing the pairwise distance among the centers from the two models.

Let the pairwise distance be $||\mu_1 - \mu_2||$ between two centers in the finite mixture. As the $||\mu_1 - \mu_2||= \sqrt{ \sum_{l=1}^{p} ||\mu_{1,l} - \mu_{2,l}||^2}$, we focus on one sub-dimension $\mu_{1,l} - \mu_{2,l}$, without loss of generality, we assume $\mu_{1,l} > \mu_{2,l}$.

For any $y_{j,l} \ge  \frac{\sum_{i\ne j} \bb E z_{1,i} y_{i,l} }{  \sum_{i\ne j} E z_{1,i} } \ge \frac{\sum_{i\ne j} \bb E z_{2,i} y_{i,l} }{  \sum_{i\ne j} E z_{2,i} } $ and $ \bb E z_{1,j} \ge \bb E z_{2,j}$,
\be
\mu_{1,l} - \mu_{2,l} = & \frac{\sum_{i\ne j} \bb E z_{1,i} y_{i,l}  + \bb E z_{1,j} y_{j,l}}{  \sum_{i\ne j} \bb E z_{1,i}  + \bb E z_{1,j} }  -  \frac{\sum_{i\ne j} \bb E z_{2,i} y_{i,l}  + \bb E z_{2,j} y_{j,l}}{  \sum_{i\ne j} \bb E z_{2,i}  + \bb E z_{2,j} } \\
 \le  & \frac{\sum_{i\ne j} \bb E z_{1,i} y_{i,l}  +  y_{j,l}}{  \sum_{i\ne j} \bb E z_{1,i}  + 1 }  -  \frac{\sum_{i\ne j} \bb E z_{2,i} y_{i,l} }{  \sum_{i\ne j} \bb E z_{2,i}  }
\ee

For any $y_{j,l} \le   \frac{\sum_{i\ne j} \bb E z_{2,i} y_{i,l} }{  \sum_{i\ne j} E z_{2,i} } \le  \frac{\sum_{i\ne j} \bb E z_{1,i} y_{i,l} }{  \sum_{i\ne j} E z_{1,i} }$ and $ \bb E z_{1,j} \le \bb E z_{2,j}$,

\be
\mu_{1,l} - \mu_{2,l}  \le  & \frac{\sum_{i\ne j} \bb E z_{1,i} y_{i,l} }{  \sum_{i\ne j} \bb E z_{1,i}  }  -  \frac{\sum_{i\ne j} \bb E z_{2,i} y_{i,l} +  y_{j,l} }{  \sum_{i\ne j} \bb E z_{2,i}+1  }
\ee

 By induction, this converts all the $E z_{k,i}$ to $z_{k,i}$ hence $\mu_{1,l} - \mu_{2,l}\le \mu^{*}_{1,l} - \mu^{*}_{2,l}$.

\end{document}

### 2016-11-21 jkmeans/R/RcppExports.R

@@ -5,7 +5,3 @@ jkmeansEM <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, flexJ = F
     .Call('jkmeans_jkmeansEM', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini, normalizeZeta)
 }
 
-jkmeansEMBatch <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, flexJ = FALSE, zetaTrunc = 0.01, useKmeansIni = TRUE, meansIni = 0L, sigma2_ini = 0.1, normalizeZeta = FALSE) {
-    .Call('jkmeans_jkmeansEMBatch', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini, normalizeZeta)
-}
-

### 2016-11-21 jkmeans/src/RcppExports.cpp

@@ -28,25 +28,3 @@ BEGIN_RCPP
     return __result;
 END_RCPP
 }
-// jkmeansEMBatch
-Rcpp::List jkmeansEMBatch(const arma::cube& y, int k, int j, int steps, double tol, bool fixW, bool flexJ, double zetaTrunc, bool useKmeansIni, const arma::mat& meansIni, double sigma2_ini, bool normalizeZeta);
-RcppExport SEXP jkmeans_jkmeansEMBatch(SEXP ySEXP, SEXP kSEXP, SEXP jSEXP, SEXP stepsSEXP, SEXP tolSEXP, SEXP fixWSEXP, SEXP flexJSEXP, SEXP zetaTruncSEXP, SEXP useKmeansIniSEXP, SEXP meansIniSEXP, SEXP sigma2_iniSEXP, SEXP normalizeZetaSEXP) {
-BEGIN_RCPP
-    Rcpp::RObject __result;
-    Rcpp::RNGScope __rngScope;
-    Rcpp::traits::input_parameter< const arma::cube& >::type y(ySEXP);
-    Rcpp::traits::input_parameter< int >::type k(kSEXP);
-    Rcpp::traits::input_parameter< int >::type j(jSEXP);
-    Rcpp::traits::input_parameter< int >::type steps(stepsSEXP);
-    Rcpp::traits::input_parameter< double >::type tol(tolSEXP);
-    Rcpp::traits::input_parameter< bool >::type fixW(fixWSEXP);
-    Rcpp::traits::input_parameter< bool >::type flexJ(flexJSEXP);
-    Rcpp::traits::input_parameter< double >::type zetaTrunc(zetaTruncSEXP);
-    Rcpp::traits::input_parameter< bool >::type useKmeansIni(useKmeansIniSEXP);
-    Rcpp::traits::input_parameter< const arma::mat& >::type meansIni(meansIniSEXP);
-    Rcpp::traits::input_parameter< double >::type sigma2_ini(sigma2_iniSEXP);
-    Rcpp::traits::input_parameter< bool >::type normalizeZeta(normalizeZetaSEXP);
-    __result = Rcpp::wrap(jkmeansEMBatch(y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini, normalizeZeta));
-    return __result;
-END_RCPP
-}

### 2016-11-21 jkmeans/src/jkmeans.cpp

@@ -32,49 +32,53 @@ Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps = 1000,
 
   return Rcpp::List::create(
       Rcpp::Named("mu") = mix.mu, Rcpp::Named("w") = mix.w,
-      Rcpp::Named("sigma2") = mix.sigma2, Rcpp::Named("zeta") = mix.zeta,
+      Rcpp::Named("Sigma") = mix.Sigma, Rcpp::Named("zeta") = mix.zeta,
       Rcpp::Named("M") = mix.clusteringMAP());
 }
 
-// [[Rcpp::export]]
-Rcpp::List jkmeansEMBatch(const arma::cube& y, int k, int j, int steps = 1000,
-                          double tol = 1E-8, bool fixW = true,
-                          bool flexJ = false, double zetaTrunc = 0.01,
-                          bool useKmeansIni = true,
-                          const arma::mat& meansIni = 0,
-                          double sigma2_ini = 0.1, bool normalizeZeta = false) {
-  if (j > k) {
-    throw std::range_error("j needs be no bigger than k");
-  }
+// // [[Rcpp::export]]
+// Rcpp::List jkmeansEMBatch(const arma::cube& y, int k, int j, int steps =
+// 1000,
+//                           double tol = 1E-8, bool fixW = true,
+//                           bool flexJ = false, double zetaTrunc = 0.01,
+//                           bool useKmeansIni = true,
+//                           const arma::mat& meansIni = 0,
+//                           double sigma2_ini = 0.1, bool normalizeZeta =
+//                           false) {
+//   if (j > k) {
+//     throw std::range_error("j needs be no bigger than k");
+//   }
 
-  int n = y.n_rows;
-  int p = y.n_cols;
-  int batchN = y.n_slices;
+//   int n = y.n_rows;
+//   int p = y.n_cols;
+//   int batchN = y.n_slices;
 
-  cube mu(k, p, batchN);
-  mat w(k, batchN);
-  vec sigma2(batchN);
-  cube zeta(n, k, batchN);
-  umat M(n, batchN);
+//   cube mu(k, p, batchN);
+//   mat w(k, batchN);
+//   vec sigma2(batchN);
+//   cube zeta(n, k, batchN);
+//   umat M(n, batchN);
 
-#pragma omp parallel for
[…]

### 2016-11-21 jkmeans/src/mixture.hpp

@@ -9,7 +9,7 @@ class Mixture {
   int J;  // truncation
 
   mat mu;
-  double sigma2;
+  vec Sigma;
   vec w;
   mat zeta;
 
@@ -27,7 +27,7 @@ class Mixture {
 
     // initialize weights
     w = ones(K) / (double)K;
-    sigma2 = 1;
+    Sigma = ones(p);
     mu = ones<mat>(K, p);
     mu.randn();
 
@@ -36,7 +36,7 @@ class Mixture {
 
   void initialize(mat meansInput, bool useKmeansIni, bool _fixW, bool _flexJ,
                   double _zetaTrunc, double sigma2_ini, bool _normalizeZeta) {
-    sigma2 = sigma2_ini;
+    Sigma.fill(sigma2_ini);
 
     if (useKmeansIni) {
       mat means;
@@ -48,7 +48,7 @@ class Mixture {
       J = 1;
       Expectation();
 
-      updateSigma2();
+      updateSigma();
 
       J = _J;
 
@@ -62,22 +62,24 @@ class Mixture {
     zetaTrunc = _zetaTrunc;
   }
 
-  double loglik(rowvec x, rowvec mu, double sigma2) {
+  double loglik(rowvec x, rowvec mu, vec Sigma) {
     rowvec diff = x - mu;
 
-    return -dot(diff, diff) / 2 / sigma2 - p / 2 * log(sigma2);
+    return -accu(diff % diff / Sigma.t()) / 2.0 - 1.0 / 2.0 * accu(log(Sigma));
   }
 
   double compTotalLoglik() {
-    vec marginal_loglik = zeros<vec>(n);
+    double marginal_loglik = 0;
 
     for (int k = 0; k < K; ++k) {
       mat diff = (y - repmat(mu.row(k), n, 1));
 
-      vec diff2 =
-          exp(-sum(diff % diff, 1) / sigma2 / 2.0 - p / 2.0 * log(sigma2));
+      mat diff2 = diff % diff;
 
-      marginal_loglik += w(k) * diff2;
+      double l = exp(-accu(diff2.each_row() / Sigma.t()) / 2.0 -
+                     1 / 2.0 * accu(log(Sigma)));
+
+      marginal_loglik += w(k) * l;
     }
 
     return accu(log(marginal_loglik));
@@ -88,13 +90,13 @@ class Mixture {
       for (int k = 0; k < K; ++k) {
         rowvec y_local = y.row(i);
[…]

### 2016-11-21 test/ARC.R

@@ -1,0 +1,74 @@
+ARC<-function(Y,K,d=2, randomStart= FALSE, repul=TRUE){
+  
+
+  
+  svdY<-svd(Y,nu = d,nv = d)
+  
+  if(d>2)
+    X<- svdY$u %*%diag(svdY$d[1:d] )
+  else
+    X<-   svdY$u * svdY$d[1]
+  if(randomStart){
+  X<- X + matrix(rnorm(length(c(X)),sd = 1),nrow(X))
+  }
+  
+  
+
+  ini<- as.matrix(kmpp(X,K))
+  
+  # plot(X,X0)
+  
+  
+  jk<- jkmeansEM(X,k = K,j= 1,1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
+  
+  Sigma<-  diag(c(jk$Sigma),d)
+  
+  Zmu<- jk$zeta %*% jk$mu
+  ini_M<- jk$M
+  ini_X<- X
+  
+  
+  # plot(jk$M)
+  
+  # plot(X0,col=mu1)
+  # plot(X,col=jk$M+1)
+  # plot(X0,col=mu1)
+  
+  
+  # e<- sum(jk$M+1!=mu1)/n
+  # min(e,1-e)
+  
+  #################
+  for(i in 1:500){
+    
+    # for(j in 1:10){
+    XX<- t(X)%*%X
+    XY<- t(X)%*%Y
+    XXinv<- solve(XX+diag(100,d))
+    
+    #compute expectation
+    EV<- solve(XX,XY)
+    temp<- XXinv%*%XY
+    EVV<- p*sigma2*XXinv +  temp %*% t(temp)
+    
+    #compute sigma2
+    sigma2<- (sum(c(Y)*c(Y))-2* sum(c(t(X)) * c(EV%*%t(Y))) + sum(c(t(X)) * c(EVV%*%t(X)))) /n/p
+    
+    #get max of X
+    
+    m<- Y%*%t(EV)/sigma2 + Zmu%*%solve(Sigma)
+    v<- solve(EVV/sigma2+solve(Sigma))
+    X<- m%*%v
+    # }
+    
+    #clustering
+    ini<- as.matrix(jk$mu)
+    if(repul){j=1}
+    else{j=K}
[…]

### 2016-11-21 test/HighP.R

@@ -1,0 +1,126 @@
+# .rs.restartR()
+require("jkmeans")
+
+n<- 100
+K<- 2
+p<- 100
+
+sigma<- 1
+mu1<- rep(K,n*K)
+
+randomMu<- unlist(sapply(1:K, function(x){rep(x, runif(1,n/10,n))}))
+mu1[1:length(randomMu)]<- randomMu
+
+mu<- matrix(0,n*K,p)
+mu[,1:5]<- mu1
+# mu[,2]<- mu1
+
+
+y<- matrix( rnorm(n*K*p, mu,sd = (sigma)), n*K)
+# hist(y,breaks = 100)
+
+svdY<- svd(y)
+
+plot(svdY$d)
+
+y1<- as.matrix( t(t(svdY$u)*svdY$d) [,c(1)])
+
+# hist(c(sapply(c(1:K),function(x){ rnorm(1000,x, (sigma2/sqrt(n)))})), breaks = 1000)
+
+
+kmpp <- function(X, k) {
+  n <- nrow(X)
+  C <- numeric(k)
+  C[1] <- sample(1:n, 1)
+  
+  for (i in 2:k) {
+    dm <- abs(outer(X, X[C, ], "-"))
+    pr <- apply(dm, 1, min)
+    pr[C] <- 0
+    C[i] <- sample(1:n, 1, prob = pr)
+  }
+  
+   X[C, ]
+}
+
+
+rmse<- function(x){
+  sqrt(mean((x$mu[,1:5]-c(1:2))^2))
+}
+
+
+
+plot(y,col=mu1)
+
+k<- K
+
+ini<- as.matrix(kmpp(y,k))
+
+
+
+kmeans <- jkmeansEM(y,k = k,j= 1,1000,tol = 1E-15,fixW = T, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
+# table(kmeans$M)
+# lines(kmeans$mu, type="p",pch=2,lwd=3,col="green")
+
+
+jk<- jkmeansEM(y,k = k,j= 1,1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
+# table(jk$M)
+# lines(jk$mu, type="p",pch=2,lwd=3,col="blue")
+jk$mu[,1]
+
+
[…]

### 2016-11-21 test/PCAC.R

@@ -1,0 +1,17 @@
+PCAC<-function(Y,K,d=2, repul=TRUE){
+  
+  svdY<-svd(Y,nu = d,nv = d)
+  
+  if(d>2)
+    X<- svdY$u %*%diag(svdY$d[1:d] )
+  else
+    X<-   svdY$u * svdY$d[1]
+  
+  ini<- as.matrix(kmpp(X,K))
+  
+  if(repul){j=1}
+  else{j=K}
+  jk<- jkmeansEM(X,k = K,j= j,1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
+  
+  return(list("M"=jk$M, "X"=X,"V"=svdY$v,"mu"=jk$mu))
+}

### 2016-11-21 test/em.R

@@ -1,0 +1,46 @@
+n<- 50
+p<- 2000
+d<- 2
+sigma2<- 3
+tau<- 0.05
+Sigma<- diag(tau,d)
+
+Zmu<- matrix( rnorm(n*d),n,d)
+X<- matrix( rnorm(n*d, c(Zmu), sqrt(tau)),n,d)
+V<- matrix( rnorm(p*d),d,p)
+
+Y<- matrix(rnorm(n*p, c(X%*%V), sd= sqrt(sigma2)), n, p)
+
+X0<- X
+
+####
+
+svdY<-svd(Y,nu = d,nv = d)
+
+X<- svdY$u %*%diag(svdY$d[1:d] )
+plot(X,X0)
+plot(Zmu,X0,xlim=c(-3,3),ylim = c(-3,3))
+
+#################
+for(i in 1:1000){
+  XX<- t(X)%*%X
+  XY<- t(X)%*%Y
+  XXinv<- solve(XX+diag(1,d))
+  
+  #compute expectation
+  EV<- solve(XX,XY)
+  temp<- XXinv%*%XY
+  EVV<- p*sigma2*XXinv +  temp %*% t(temp)
+  
+  #compute sigma2
+  sigma2<- (sum(c(Y)*c(Y))-2* sum(c(t(X)) * c(EV%*%t(Y))) + sum(c(t(X)) * c(EVV%*%t(X)))) /n/p
+  
+  #get max of X
+  
+  m<- Y%*%t(EV)/sigma2 + Zmu%*%solve(Sigma)
+  v<- solve(EVV/sigma2+solve(Sigma))
+  X<- m%*%v
+}
+###
+
+plot(X,X0,xlim=c(-3,3),ylim = c(-3,3))

### 2016-11-21 test/em_c.R

@@ -1,0 +1,141 @@
+#
+
+require("sparcl")
+
+n<- 300
+p<- 2000
+d<- 3
+sigma2<- 2
+tau<- 0.5
+Sigma<- diag(tau,d)
+
+###################################
+
+K<- 3
+mu1<- rep(K,n)
+# randomMu<- unlist(sapply(1:K, function(x){rep(x, runif(1,n/10,n))}))
+# mu1[1:length(randomMu)]<- randomMu
+# mu1<- mu1[1:n]
+mu1[1:(n/K)]<-1
+mu1[(n/K+1):(n/K*2)]<-2
+mu1[(n/K*2+1):(n/K*3)]<-3
+
+###############################
+
+
+Zmu<- matrix(mu1,n,d)
+X<- matrix( rnorm(n*d, c(Zmu), sqrt(tau)),n,d)
+V<- matrix( rnorm(p*d),d,p)
+
+Y<- matrix(rnorm(n*p, c(X%*%V), sd= sqrt(sigma2)), n, p)
+
+# R<- svd(matrix(rnorm(n^2),n))$u
+
+# Y<- R%*%Y
+X0<- X
+
+###############
+#Simple PCA+ Cluster
+
+
+
+#######
+
+####initialization####
+setwd("~/git/jkmeans/test/")
+source("kmpp.r")
+require("jkmeans")
+require("mclust")
+source("ARC.R")
+source("PCAC.R")
+
+pca_M<-PCAC(Y,K,d)
+arc_M<-ARC(Y,K,d,randomStart = T)
+pcaGMM_M<-PCAC(Y,K,d,repul = F)
+
+ldgmm_M<-ARC(Y,K,d,randomStart = T,repul = F)
+
+
+pdf("originalY2d.pdf",6,6)
+plot(Y[,1:2],col=mu1, xlab="",ylab="")
+dev.off()
+
+pdf("originalX.pdf",6,6)
+plot(X,col=mu1, xlab="",ylab="")
+dev.off()
+# plot(Y[,1:2],col=mu1)
+
+pdf("pca_gmm.pdf",6,6)
+plot(pcaGMM_M$X, col=pcaGMM_M$M +1, xlab="",ylab="")
+dev.off()
+
+pdf("pca_arc.pdf",6,6)
+plot(pca_M$X, col=pca_M$M +1, xlab="",ylab="")
+dev.off()
+
+pdf("ld_gmm.pdf",6,6)
[…]

### 2016-11-21 test/hd_c.R

@@ -1,0 +1,64 @@
+ARC<-function(Y,K,d=2){
+  
+  source("kmpp.r")
+  require("jkmeans")
+  require("mclust")
+  
+  svdY<-svd(Y,nu = d,nv = d)
+  
+  X<- svdY$u %*%diag(svdY$d[1:d] )
+  ini<- as.matrix(kmpp(X,K))
+  
+  # plot(X,X0)
+  
+  
+  jk<- jkmeansEM(X,k = K,j= 1,1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
+  
+  
+  Zmu<- jk$zeta %*% jk$mu
+  ini_M<- jk$M
+  ini_X<- X
+  
+  
+  plot(jk$M)
+  
+  # plot(X0,col=mu1)
+  plot(X,col=jk$M+1)
+  # plot(X0,col=mu1)
+  
+  
+  # e<- sum(jk$M+1!=mu1)/n
+  # min(e,1-e)
+  
+  #################
+  for(i in 1:200){
+    
+    # for(j in 1:10){
+    XX<- t(X)%*%X
+    XY<- t(X)%*%Y
+    XXinv<- solve(XX+diag(100,d))
+    
+    #compute expectation
+    EV<- solve(XX,XY)
+    temp<- XXinv%*%XY
+    EVV<- p*sigma2*XXinv +  temp %*% t(temp)
+    
+    #compute sigma2
+    sigma2<- (sum(c(Y)*c(Y))-2* sum(c(t(X)) * c(EV%*%t(Y))) + sum(c(t(X)) * c(EVV%*%t(X)))) /n/p
+    
+    #get max of X
+    
+    m<- Y%*%t(EV)/sigma2 + Zmu%*%solve(Sigma)
+    v<- solve(EVV/sigma2+solve(Sigma))
+    X<- m%*%v
+    # }
+    
+    #clustering
+    ini<- as.matrix(jk$mu)
+    jk<- jkmeansEM(X,k = K,j= 1,1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
+    Zmu<- jk$zeta %*% jk$mu
+    
+    Sigma<-  diag(c(jk$Sigma),d)
+  }
+  return(jk$M)
+}

### 2016-11-21 test/highPnewModel.r

@@ -1,0 +1,75 @@
+require("jkmeans")
+
+n<- 100
+K<- 2
+p<- 100
+
+sigma<- 1
+mu1<- rep(K,n*K)
+
+
+randomMu<- unlist(sapply(1:K, function(x){rep(x, runif(1,n/10,n))}))
+mu1[1:length(randomMu)]<- randomMu
+
+
+kmpp <- function(X, k) {
+  n <- nrow(X)
+  C <- numeric(k)
+  C[1] <- sample(1:n, 1)
+  
+  for (i in 2:k) {
+    dm <- abs(outer(X, X[C, ], "-"))
+    pr <- apply(dm, 1, min)
+    pr[C] <- 0
+    C[i] <- sample(1:n, 1, prob = pr)
+  }
+  
+  X[C, ]
+}
+
+mu<- matrix(0,n*K,p)
+mu[,1:5]<- mu1
+
+mu[10:20,1:5]<- 1.5
+
+y<- matrix( rnorm(n*K*p, mu,sd = (sigma)), n*K)
+# hist(y,breaks = 100)
+
+
+svdY<- svd(y)
+
+plot(svdY$d)
+
+V<- svdY$v[1,]
+
+X<- as.matrix(svdY$u[,1]*svdY$d[1])
+
+k<- 3
+
+ini<- as.matrix(kmpp(X,k))
+
+trace_mu<- numeric()
+
+for(i in 1:100){
+  jk<- jkmeansEM(X,k = k,j= 1,1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
+  
+  Zmu<- jk$zeta %*% jk$mu
+  V<- solve(t(Zmu)%*%Zmu ,t(Zmu)%*%y)
+  X<- t(solve(V%*%t(V), t(y%*%t(V))))
+  ini <- as.matrix(jk$mu[order(jk$mu[,1]),])
+  
+  trace_mu<- rbind(trace_mu, t(ini))
+  print(min(sum(jk$M+1 != mu1)/n/K, 1-sum(jk$M+1 != mu1)/n/K))
+}
+
+est<- jk$mu%*%(V)
+est[,1]
+
+jk$w
+
+ts.plot(trace_mu)
+
+# require("mclust")
+
+# adjustedRandIndex(jk$M,mu1)
+

### 2016-11-21 test/kmpp.r

@@ -1,0 +1,14 @@
+kmpp <- function(X, k) {
+  n <- nrow(X)
+  C <- numeric(k)
+  C[1] <- sample(1:n, 1)
+  
+  for (i in 2:k) {
+    dm <- abs(outer(X, X[C, ], "-"))
+    pr <- apply(dm, 1, min)
+    pr[C] <- 0
+    C[i] <- sample(1:n, 1, prob = pr)
+  }
+  
+  X[C, ]
+}

### 2016-11-21 test/multi_test.r

@@ -1,0 +1,30 @@
+n<- 2000
+p<- 2000
+d<- 2
+sigma2<- 2
+tau<- 0.5
+Sigma<- diag(tau,d)
+
+###################################
+
+K<- 3
+mu1<- rep(K,n)
+# randomMu<- unlist(sapply(1:K, function(x){rep(x, runif(1,n/10,n))}))
+# mu1[1:length(randomMu)]<- randomMu
+# mu1<- mu1[1:n]
+mu1[1:(n/K)]<-1
+mu1[(n/K+1):(n/K*2)]<-2
+mu1[(n/K*2+1):(n/K*3)]<-3
+
+
+mu<- matrix(mu1,n,2)
+
+X<- matrix(rnorm(n*2,mu),n)
+X[,2]<- rnorm(n, X[,2],sd=1)
+
+ini<- as.matrix(kmpp(X,K))
+
+
+jk<- jkmeansEM(X,k = K,j= 1,1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
+
+jk$Sigma

### 2016-11-28 docs/draft.tex
updated draft

@@
 \usepackage[utf8]{inputenc}
-\usepackage{algorithm}
-\usepackage[noend]{algpseudocode}
+%\usepackage{algorithm}
+%\usepackage[noend]{algpseudocode}
+\usepackage[]{algorithm2e}
 
@@
 
-To ensure identifiability, we regularize $V_{i,j}\sim N(0,1)$.
+To ensure identifiability, we regularize $V_{i,j}\sim N(0,\nu)$, where $\nu$ is a large variance (e.g.
+$10^6$).
 It is possible to  maximize the log-likelihood alternatively over $\V$ and $\X$, however, this would underestimate the variability of the $\V$, leading to suboptimal result.
@@
 \be
-\V_{.,j} \stackrel{indep}{\sim} N\left( (\X'\X+\I)^{-1}\X'\Y_{.,j}, \sigma^2 (\X'\X+\I)^{-1}    \right)
+\V_{.,j} \stackrel{indep}{\sim} N\left( (\X'\X+\I \nu)^{-1}\X'\Y_{.,j}, \sigma^2 (\X'\X+\I\nu)^{-1}    \right)
 \ee
@@
 \be
-\E\V =  (\X'\X+\I)^{-1}\X'\Y,  \qquad\E\V \V' = p\sigma^2 (\X'\X+\I)^{-1} + (\X'\X+\I)^{-1}\X'\Y \Y'\X (\X'\X+\I)^{-1}
+\E\V =  (\X'\X+\I \nu)^{-1}\X'\Y,  \qquad\E\V \V' = p\sigma^2 (\X'\X+\I \nu)^{-1} + (\X'\X+\I \nu)^{-1}\X'\Y \Y'\X (\X'\X+\I \nu)^{-1}
 \ee
@@
 
+As the loss function, the expected log-likelihood is:
+
+\be
+\bb E \log L = &  - \frac{1}{2}  [    \{\mbox{vec} (\Y)'\mbox{vec} (\Y)  -2 \mbox{vec}  ( \hat\X')' \mbox{vec}( \E\V \Y') + \mbox{vec}  ( \hat\X')' \mbox{vec}( \E\V\V' {\hat \X}' )\}/{\sigma^2} +\\
+& {\sum_i^n (\X_{i,.} - \Z_{i,.} \bmu)' \Sigma^{-1}  (\X_{i,.} - \Z_{i,.} \bmu)} + n\log \det  \bSigma + np \log \sigma^2 ]
+\ee
+
 \subsection{Clustering}
@@
 
-Then the whole estimation can be carried out by alternating in the two steps.
+Then the whole estimation can be carried out by alternating in the two steps. To summarize, the estimating algorithm is shown in Algorithm~\ref{arc_algo}.
 
-\section{Simulation}
+\begin{algorithm}[H]
+ initialization\;
+ \While{Expected likelihood $\bb E \log L$ change $>$ threshold}{
+  Using $\hat \Z \hat \bmu$, compute expectation: $\bb E \V$ and $\bb E \V\V'$\;
+  Using expected values, maximize: $\hat \X$ \;
+   \While{change in $\hat \bmu$ $>$ threshold}{
+   		  Compute MLEs: $\hat \bmu$ and $\hat \bSigma$\;
+		  Compute Most Probable Assignment: $\hat \Z$ \;
+	}
+ Compute expected likelihood $\bb E \log L$\;
+ }
+ \caption{Estimation Algorithm for Automatic Repulsive Clustering}
+ \label{arc_algo}
+\end{algorithm}
 
+\section{Experiments}
+
+\subsection{Synthetic Data}
 Data generation:
@@
  % \centering
-  \begin{subfigure}[b]{0.49\textwidth}
- \includegraphics[width=0.8\textwidth]{pics/originalY2d.pdf}
-  \caption{The 3 component view in the first 2 dimensions.}
+  \begin{subfigure}[b]{0.32\textwidth}
+ \includegraphics[width=0.8\textwidth]{pics/simX}
+  \caption{The lower dimensions used in data generation.}
 \end{subfigure}
   \hfill
-   \begin{subfigure}[b]{0.49\textwidth}
- \includegraphics[width=0.8\textwidth]{pics/ld_arc_truth.pdf}
-  \caption{The true latent 2d separable subspace view.}
-\end{subfigure}
-
- \begin{subfigure}[b]{0.49\textwidth}
- \includegraphics[width=0.8\textwidth]{pics/pca_gmm.pdf}
-  \caption{GMM on principle components}
+   \begin{subfigure}[b]{0.32\textwidth}
+ \includegraphics[width=0.8\textwidth]{pics/simPCAX}
+  \caption{The lower dimensions discovered by PCA.}
 \end{subfigure}
   \hfill
-   \begin{subfigure}[b]{0.49\textwidth}
- \includegraphics[width=0.8\textwidth]{pics/pca_arc.pdf}
-  \caption{ARC on principle components}
+ \begin{subfigure}[b]{0.32\textwidth}
+ \includegraphics[width=0.8\textwidth]{pics/simARCX}
+  \caption{The lower dimensions discovered by ARC.}
 \end{subfigure}
-
- \begin{subfigure}[b]{0.49\textwidth}
- \includegraphics[width=0.8\textwidth]{pics/ld_gmm.pdf}
-  \caption{GMM on flexible reduced Dimension}
-\end{subfigure}
-  \hfill
-   \begin{subfigure}[b]{0.49\textwidth}
- \includegraphics[width=0.8\textwidth]{pics/ld_arc.pdf}
-  \caption{ARC on flexible reduced Dimension}
-\end{subfigure}
-
  \end{figure}
 
-\section{Application}
+\subsection{Real Data}
+
+MNIST:
+
+Four digits: $0,3,7,9$. Pick $100$ samples for each digit.
+
+Compute average ARI over $10$ tests.
+
+\begin{table}[H]
+\centering
+	\begin{tabular}{| l | r |}
+	\hline
+			Model & $ARI$ \\
+	\hline
+			GMM & 0.50 \\
+			PCA + GMM &  0.38  \\
+			ARC		& {\bf 0.57} \\
+			\hline
+	\end{tabular}
+\end{table}
+
 \section{Discussion}

### 2016-11-28 jkmeans/R/RcppExports.R

@@ -1,7 +1,15 @@
 # This file was generated by Rcpp::compileAttributes
 # Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
 
-jkmeansEM <- function(y, k, j, steps = 1000L, tol = 1E-8, fixW = TRUE, flexJ = FALSE, zetaTrunc = 0.01, useKmeansIni = TRUE, meansIni = 0L, sigma2_ini = 0.1, normalizeZeta = FALSE) {
-    .Call('jkmeans_jkmeansEM', PACKAGE = 'jkmeans', y, k, j, steps, tol, fixW, flexJ, zetaTrunc, useKmeansIni, meansIni, sigma2_ini, normalizeZeta)
+jkmeansEM <- function(y, k, meansIni, steps = 1000L, tol = 1E-8, fixW = FALSE, useKmeansIni = FALSE, sigma2_ini = 0.1) {
+    .Call('jkmeans_jkmeansEM', PACKAGE = 'jkmeans', y, k, meansIni, steps, tol, fixW, useKmeansIni, sigma2_ini)
+}
+
+rDARC <- function(Y, d, k, meansIni, steps = 1000L, tol = 1E-8, randomStart = TRUE, fixW = FALSE, sigma2_ini = 0.1, ver = 1L) {
+    .Call('jkmeans_rDARC', PACKAGE = 'jkmeans', Y, d, k, meansIni, steps, tol, randomStart, fixW, sigma2_ini, ver)
+}
+
+kmpp <- function(Y, k) {
+    .Call('jkmeans_kmpp', PACKAGE = 'jkmeans', Y, k)
 }
 

### 2016-11-28 jkmeans/src/RcppExports.cpp

@@ -7,24 +7,52 @@
 using namespace Rcpp;
 
 // jkmeansEM
-Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps, double tol, bool fixW, bool flexJ, double zetaTrunc, bool useKmeansIni, const arma::mat& meansIni, double sigma2_ini, bool normalizeZeta);
-RcppExport SEXP jkmeans_jkmeansEM(SEXP ySEXP, SEXP kSEXP, SEXP jSEXP, SEXP stepsSEXP, SEXP tolSEXP, SEXP fixWSEXP, SEXP flexJSEXP, SEXP zetaTruncSEXP, SEXP useKmeansIniSEXP, SEXP meansIniSEXP, SEXP sigma2_iniSEXP, SEXP normalizeZetaSEXP) {
+Rcpp::List jkmeansEM(const arma::mat& y, int k, const arma::mat& meansIni, int steps, double tol, bool fixW, bool useKmeansIni, double sigma2_ini);
+RcppExport SEXP jkmeans_jkmeansEM(SEXP ySEXP, SEXP kSEXP, SEXP meansIniSEXP, SEXP stepsSEXP, SEXP tolSEXP, SEXP fixWSEXP, SEXP useKmeansIniSEXP, SEXP sigma2_iniSEXP) {
 BEGIN_RCPP
     Rcpp::RObject __result;
     Rcpp::RNGScope __rngScope;
     Rcpp::traits::input_parameter< const arma::mat& >::type y(ySEXP);
     Rcpp::traits::input_parameter< int >::type k(kSEXP);
-    Rcpp::traits::input_parameter< int >::type j(jSEXP);
+    Rcpp::traits::input_parameter< const arma::mat& >::type meansIni(meansIniSEXP);
     Rcpp::traits::input_parameter< int >::type steps(stepsSEXP);
     Rcpp::traits::input_parameter< double >::type tol(tolSEXP);
     Rcpp::traits::input_parameter< bool >::type fixW(fixWSEXP);
-    Rcpp::traits::input_parameter< bool >::type flexJ(flexJSEXP);
-    Rcpp::traits::input_parameter< double >::type zetaTrunc(zetaTruncSEXP);
     Rcpp::traits::input_parameter< bool >::type useKmeansIni(useKmeansIniSEXP);
-    Rcpp::traits::input_parameter< const arma::mat& >::type meansIni(meansIniSEXP);
[…]

### 2016-11-28 jkmeans/src/jkmeans.cpp

@@ -1,8 +1,13 @@
 // -*- mode: C++; c-indent-level: 4; c-basic-offset: 4; indent-tabs-mode: nil;
 // -*-
 
+#include <memory>
 #include "RcppArmadillo.h"
+
+#include "kmpp.hpp"
+
 #include "mixture.hpp"
+#include "reducedDimARC.hpp"
 
 // via the depends attribute we tell Rcpp to create hooks for
 // RcppArmadillo so that the build process will know what to do
@@ -12,19 +17,12 @@
 using namespace arma;
 
 // [[Rcpp::export]]
-Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps = 1000,
-                     double tol = 1E-8, bool fixW = true, bool flexJ = false,
-                     double zetaTrunc = 0.01, bool useKmeansIni = true,
-                     const arma::mat& meansIni = 0, double sigma2_ini = 0.1,
-                     bool normalizeZeta = false) {
-  Mixture mix(y, k, j);
+Rcpp::List jkmeansEM(const arma::mat& y, int k, const arma::mat& meansIni,
+                     int steps = 1000, double tol = 1E-8, bool fixW = false,
+                     bool useKmeansIni = false, double sigma2_ini = 0.1) {
+  Mixture mix(y, k);
 
-  if (j > k) {
-    throw std::range_error("j needs be no bigger than k");
-  }
-
-  mix.initialize(meansIni, useKmeansIni, fixW, flexJ, zetaTrunc, sigma2_ini,
-                 normalizeZeta);
+  mix.initialize(meansIni, useKmeansIni, fixW, sigma2_ini);
 
   mix.runEM(steps, tol);
 
@@ -36,6 +34,27 @@ Rcpp::List jkmeansEM(const arma::mat& y, int k, int j, int steps = 1000,
       Rcpp::Named("M") = mix.clusteringMAP());
 }
 
+// [[Rcpp::export]]
+Rcpp::List rDARC(const arma::mat& Y, int d, int k, const arma::mat& meansIni,
+                 int steps = 1000, double tol = 1E-8, bool randomStart = true,
+                 bool fixW = false, double sigma2_ini = 0.1, int ver = 1) {
+  reducedDimARC rdARC(Y, d, k, randomStart, sigma2_ini, ver, fixW);
[…]

### 2016-11-28 jkmeans/src/kmpp.hpp

@@ -1,0 +1,36 @@
+using namespace arma;
+
+mat kmeanspp(mat X, int K) {
+  int n = X.n_rows;
+  uvec C(K);
+
+  std::random_device rd;
+  std::mt19937 gen(rd());
+
+  vec pr = ones(n);
+
+  auto randDiscrete = [&](vec pr) {
+    std::vector<double> p_stdvec = conv_to<std::vector<double> >::from(pr);
+    std::discrete_distribution<int> distribution(p_stdvec.begin(),
+                                                 p_stdvec.end());
+
+    return distribution(gen);
+  };
+
+  C(0) = randDiscrete(pr);
+
+  mat D = X.each_row() - X.row(C(0));
+  vec D2 = sum(D % D, 1);
+
+  for (int i = 1; i < K; i++) {
+    pr = D2;
+    pr(C(span(0, i - 1))).fill(0);
+    C(i) = randDiscrete(pr);
+
+    mat D = X.each_row() - X.row(C(i));
+    vec D2new = sum(D % D, 1);
+
+    D2 = min(D2, D2new);
+  }
+  return X.rows(C);
+}

### 2016-11-28 jkmeans/src/mixture.hpp

@@ -6,7 +6,6 @@ class Mixture {
   int n;
   int p;
   int K;  // max components
-  int J;  // truncation
 
   mat mu;
   vec Sigma;
@@ -14,14 +13,10 @@ class Mixture {
   mat zeta;
 
   bool fixW;
-  bool normalizeZeta;
-  bool flexJ;
-  double zetaTrunc;
 
-  Mixture(mat _y, int _K, int _J) {
+  Mixture(mat _y, int _K) {
     y = _y;
     K = _K;
-    J = _J;
     n = y.n_rows;
     p = y.n_cols;
 
@@ -34,8 +29,8 @@ class Mixture {
     zeta = ones<mat>(n, K);
   }
 
-  void initialize(mat meansInput, bool useKmeansIni, bool _fixW, bool _flexJ,
-                  double _zetaTrunc, double sigma2_ini, bool _normalizeZeta) {
+  void initialize(mat meansInput, bool useKmeansIni, bool _fixW,
+                  double sigma2_ini) {
     Sigma.fill(sigma2_ini);
 
     if (useKmeansIni) {
@@ -44,27 +39,18 @@ class Mixture {
       mu = means.t();
 
       // initialize sigma2 via K-means
-      int _J = J;  // back up the user defined J
-      J = 1;
       Expectation();
-
       updateSigma();
-
-      J = _J;
 
     } else {
       mu = meansInput;
     }
 
-    normalizeZeta = _normalizeZeta;
     fixW = _fixW;
-    flexJ = _flexJ;
-    zetaTrunc = _zetaTrunc;
   }
 
   double loglik(rowvec x, rowvec mu, vec Sigma) {
     rowvec diff = x - mu;
-
     return -accu(diff % diff / Sigma.t()) / 2.0 - 1.0 / 2.0 * accu(log(Sigma));
   }
 
@@ -91,38 +77,22 @@ class Mixture {
         rowvec y_local = y.row(i);
         rowvec mu_local = mu.row(k);
         zeta(i, k) = log(w(k)) + loglik(y_local, mu_local, Sigma);
-        if (std::isnan(zeta(i, k))) {
-          cout << w(k) << endl;
-          cout << loglik(y_local, mu_local, Sigma) << endl;
-          cout << y_local << endl;
-          cout << mu_local << endl;
-          // cout << Sigma << endl;
-        }
+        // if (std::isnan(zeta(i, k))) {
+        //   cout << w(k) << endl;
+        //   cout << loglik(y_local, mu_local, Sigma) << endl;
[…]

### 2016-11-28 jkmeans/src/reducedDimARC.hpp

@@ -1,0 +1,102 @@
+using namespace arma;
+
+class reducedDimARC {
+ public:
+  int n, d, p, K;
+  mat Y;
+
+  mat X;
+  mat EV;
+  mat EVV;
+  mat Sigma;
+  mat Zmu;
+  double sigma2;
+  bool randomStart;
+
+  std::shared_ptr<Mixture> mix;
+  unsigned int ver;
+
+  reducedDimARC(mat _Y, int _d, int _K, bool _randomStart, double sigma2_ini,
+                unsigned int _ver, bool fixW) {
+    ver = _ver;
+
+    Y = _Y;
+    d = _d;
+    K = _K;
+    n = Y.n_rows;
+    p = Y.n_cols;
+    randomStart = _randomStart;
+
+    {
+      mat U;
+      vec s;
+      mat V;
+      svd(U, s, V, Y);
+      U = U.cols(span(0, d - 1));
+      s = s(span(0, d - 1));
+      X = U * diagmat(s);
+    }
+
+    if (randomStart) {
+      // X += randn(n, d);
+      X = randn(n, d);
+    }
+
+    Sigma = eye(d, d);
+    sigma2 = 1;
+
+    mix.reset(new Mixture(X, K));
+
+    // use kmeans++ for meansInput
+
+    mat meansInput = kmeanspp(X, K);
+
+    mix->initialize(meansInput, false, fixW, sigma2_ini);
+
+    mix->Expectation();
+    mix->Maximization();
+
+    Zmu = (mix->zeta) * (mix->mu);
+  };
+
+  void Estep() {
+    mat XX = X.t() * X;
+    mat XY = X.t() * Y;
+    mat XXinv = inv(XX + eye(d, d) * 1E-5);
+
+    // expectation
+    EV = XXinv * XY;
+    EVV = p * sigma2 * XXinv + EV * EV.t();
+  }
+
+  void Mstep() {
+    // sigma2
+    sigma2 = (accu(Y % Y) - 2 * accu(vectorise(X.t()) % vectorise(EV * Y.t())) +
+              accu(vectorise(X.t()) % vectorise(EVV * X.t()))) /
+             n / p;
+    // compute X
+    mat invSigma = inv(Sigma);
+    mat m = Y * EV.t() / sigma2 + Zmu * invSigma;
+    mat v = inv(EVV / sigma2 + invSigma);
+    X = m * v;
+
+    switch (ver) {
+      case 1:
[…]

### 2016-11-28 mnist_data/bin2csv.m

@@ -1,0 +1,11 @@
+m = ones(28*28, 0);
+for i = 0:9
+    infile = strcat('data',num2str(i))
+    fid = fopen(infile);
+    [ t1, N ] = fread(fid, [28 * 28 1000], 'uchar');
+    fclose(fid);
+    m = [m t1];
+end
+
+ofile = strcat('mnist.csv');
+csvwrite(ofile, m);

### 2016-11-28 test/ARC.R

@@ -1,6 +1,6 @@
 ARC<-function(Y,K,d=2, randomStart= FALSE, repul=TRUE){
   
-
+  
   
   svdY<-svd(Y,nu = d,nv = d)
   
@@ -9,17 +9,17 @@ ARC<-function(Y,K,d=2, randomStart= FALSE, repul=TRUE){
   else
     X<-   svdY$u * svdY$d[1]
   if(randomStart){
-  X<- X + matrix(rnorm(length(c(X)),sd = 1),nrow(X))
+    X<- X+matrix(rnorm(length(c(X)),sd = 1),nrow(X))
   }
   
   
-
+  
   ini<- as.matrix(kmpp(X,K))
   
   # plot(X,X0)
   
   
-  jk<- jkmeansEM(X,k = K,j= 1,1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
+  jk<- jkmeansEM(y= X,k = K,steps = 1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1)
   
   Sigma<-  diag(c(jk$Sigma),d)
   
@@ -44,12 +44,12 @@ ARC<-function(Y,K,d=2, randomStart= FALSE, repul=TRUE){
     # for(j in 1:10){
     XX<- t(X)%*%X
     XY<- t(X)%*%Y
-    XXinv<- solve(XX+diag(100,d))
+    XXinv<- solve(XX+diag(1E-3,d))
     
     #compute expectation
-    EV<- solve(XX,XY)
-    temp<- XXinv%*%XY
-    EVV<- p*sigma2*XXinv +  temp %*% t(temp)
+    # EV<- solve(XX,XY)
+    EV<- XXinv%*%XY
+    EVV<- p*sigma2*XXinv +  EV %*% t(EV)
     
     #compute sigma2
     sigma2<- (sum(c(Y)*c(Y))-2* sum(c(t(X)) * c(EV%*%t(Y))) + sum(c(t(X)) * c(EVV%*%t(X)))) /n/p
@@ -57,18 +57,20 @@ ARC<-function(Y,K,d=2, randomStart= FALSE, repul=TRUE){
     #get max of X
     
     m<- Y%*%t(EV)/sigma2 + Zmu%*%solve(Sigma)
-    v<- solve(EVV/sigma2+solve(Sigma))
[…]

### 2016-11-28 test/PCAC.R

@@ -9,9 +9,11 @@ PCAC<-function(Y,K,d=2, repul=TRUE){
   
   ini<- as.matrix(kmpp(X,K))
   
-  if(repul){j=1}
-  else{j=K}
-  jk<- jkmeansEM(X,k = K,j= j,1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
+  # if(repul){j=1}
+  # else{j=K}
+  # jk<- jkmeansEM(X,k = K,j= j,1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
+  jk<- jkmeansEM(y= X,k = K,steps = 1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1)
+  
   
   return(list("M"=jk$M, "X"=X,"V"=svdY$v,"mu"=jk$mu))
 }

### 2016-11-28 test/em_c.R

@@ -50,41 +50,41 @@ source("ARC.R")
 source("PCAC.R")
 
 pca_M<-PCAC(Y,K,d)
+# pcaGMM_M<-PCAC(Y,K,d,repul = F)
+
+# ldgmm_M<-ARC(Y,K,d,randomStart = F,repul = F)
 arc_M<-ARC(Y,K,d,randomStart = T)
-pcaGMM_M<-PCAC(Y,K,d,repul = F)
-
-ldgmm_M<-ARC(Y,K,d,randomStart = T,repul = F)
 
 
-pdf("originalY2d.pdf",6,6)
+# pdf("originalY2d.pdf",6,6)
 plot(Y[,1:2],col=mu1, xlab="",ylab="")
-dev.off()
+# dev.off()
 
-pdf("originalX.pdf",6,6)
+# pdf("originalX.pdf",6,6)
 plot(X,col=mu1, xlab="",ylab="")
-dev.off()
+# dev.off()
 # plot(Y[,1:2],col=mu1)
 
-pdf("pca_gmm.pdf",6,6)
+# pdf("pca_gmm.pdf",6,6)
 plot(pcaGMM_M$X, col=pcaGMM_M$M +1, xlab="",ylab="")
-dev.off()
+# dev.off()
 
-pdf("pca_arc.pdf",6,6)
+# pdf("pca_arc.pdf",6,6)
 plot(pca_M$X, col=pca_M$M +1, xlab="",ylab="")
-dev.off()
+# dev.off()
 
-pdf("ld_gmm.pdf",6,6)
+# pdf("ld_gmm.pdf",6,6)
 plot(ldgmm_M$X, col=ldgmm_M$M +1, xlab="",ylab="")
-dev.off()
+# dev.off()
 
-pdf("ld_arc.pdf",6,6)
+# pdf("ld_arc.pdf",6,6)
 plot(arc_M$X, col=arc_M$M +1, xlab="",ylab="")
-dev.off()
+# dev.off()
 
 
-pdf("ld_arc_truth.pdf",6,6)
+# pdf("ld_arc_truth.pdf",6,6)
 plot(arc_M$X, col=4-mu1, xlab="",ylab="")
-dev.off()
+# dev.off()
 
 
 adjustedRandIndex(pca_M$M,mu1)
@@ -105,13 +105,12 @@ Y<- as.matrix(iris[,1:4])
 # plot(X)
 K<-3
 ini<- as.matrix(kmpp(Y,K))
[…]

### 2016-11-28 test/kmpp.r

@@ -3,11 +3,16 @@ kmpp <- function(X, k) {
   C <- numeric(k)
   C[1] <- sample(1:n, 1)
   
+  D2<-  apply(X,1,function(x){       sum((x-X[C[1],])^2)})
+  
   for (i in 2:k) {
-    dm <- abs(outer(X, X[C, ], "-"))
-    pr <- apply(dm, 1, min)
+    
+    pr<- D2
     pr[C] <- 0
     C[i] <- sample(1:n, 1, prob = pr)
+    
+    D2new <-  apply(X,1,function(x){       sum((x-X[C[i],])^2)})
+    D2<- pmin(D2new,D2)
   }
   
   X[C, ]

### 2016-11-28 test/mnist_test.r

@@ -1,0 +1,61 @@
+setwd("~/git/jkmeans/test/")
+source("kmpp.r")
+require("jkmeans")
+require("mclust")
+source("PCAC.R")
+
+
+mnist<- read.csv("../mnist_data/mnist.csv",header = F)
+
+pick<- c(0,3,7,9)
+# pick<- c(0:9)
+
+subset<- c(sapply(pick*1000, function(x){x+ c(1:100)*2}))
+
+
+mnist_subset<- mnist[,subset]
+Y<-t(mnist_subset)
+
+
+M<- rep(pick,each=100)
+
+
+K<- 4
+
+# mclust<-Mclust(Y,G = K,modelNames = "VEI")
+# raw_M<-apply(mclust$z, 1, function(x){c(1:K)[x==max(x)]})
+
+# svd0<- svd(Y)
+
+######
+
+#######
+
+
+mclust<-Mclust(Y,G = K)
+raw_M<-apply(mclust$z, 1, function(x){c(1:K)[x==max(x)]})
+adjustedRandIndex(raw_M, M)
+
+ari<- numeric()
+for(i in 1:10){
+  pca_M<-PCAC(Y,K,d = 6)
+  ari<- c(ari,adjustedRandIndex(pca_M$M, M))
+}
+
+mean(ari)
+
+arc_ari<- numeric()
+for(i in 1:10){
+  arc_M<- rDARC(Y,d = 6,k = K,meansIni = matrix(1,2),steps = 1E3,sigma2_ini = 0.1,randomStart = F,ver = 1,fixW = T)
+  arc_ari<- c(arc_ari,  adjustedRandIndex(arc_M$M, M))
+}
+
+mean(arc_ari)
+
+
+plot(raw_M)
+plot(pca_M$M)
+plot(arc_M$M)
+
+
+

### 2016-11-28 test/multi_test.r

@@ -1,3 +1,5 @@
+setwd("~/git/jkmeans/test/")
+source("kmpp.r")
 n<- 2000
 p<- 2000
 d<- 2
@@ -19,12 +21,15 @@ mu1[(n/K*2+1):(n/K*3)]<-3
 
 mu<- matrix(mu1,n,2)
 
-X<- matrix(rnorm(n*2,mu),n)
+X<- matrix(rnorm(n*2,mu,0.5),n)
 X[,2]<- rnorm(n, X[,2],sd=1)
 
 ini<- as.matrix(kmpp(X,K))
 
+require("jkmeans")
 
-jk<- jkmeansEM(X,k = K,j= 1,1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 0.1,normalizeZeta = T)
+jk<- jkmeansEM(y= X,k = K,steps = 1000,tol = 1E-15,fixW = F, meansIni = ini ,useKmeansIni = F,sigma2_ini = 1)
 
 jk$Sigma
+
+jk$mu

### 2016-11-28 test/rdARCtest.r

@@ -1,0 +1,91 @@
+require("jkmeans")
+require("mclust")
+setwd("~/git/jkmeans/test/")
+source("kmpp.r")
+source("PCAC.R")
+
+
+Y<- matrix(rnorm(1000),nrow = 100)
+
+mu<- rep(c(1:3),each=100)
+
+Y<- matrix(rnorm(300*3,mu, sd = 0.5),300)
+
+# fit<- rDARC(Y,d = 2,k = 2,meansIni = matrix(1,2))
+
+# kmpp(Y,3)
+
+
+#
+
+
+n<- 300
+p<- 1000
+d<- 2
+sigma2<- 1
+tau<- 0.1
+Sigma<- diag(tau,d)
+
+###################################
+
+K<- 3
+mu1<- matrix(K,n,d)
+# randomMu<- unlist(sapply(1:K, function(x){rep(x, runif(1,n/10,n))}))
+# mu1[1:length(randomMu)]<- randomMu
+# mu1<- mu1[1:n]
+M<- rep(c(1:K),each=n/K)
+
+mu1[1:(n/K),]<-1
+# mu1[1:(n/K),2]<-2
+
+mu1[(n/K+1):(n/K*2),]<-2
+# mu1[(n/K+1):(n/K*2),2]<-1
+
+mu1[(n/K*2+1):(n/K*3),]<-3
+# mu1[(n/K*2+1):(n/K*3),2]<-3
+
+# mu1[1:50,]<- 1.5
+
+###
+
+Zmu<- matrix(mu1,n,d)
+X<- matrix( rnorm(n*d, c(Zmu), sqrt(tau)),n,d)
+V<- matrix( rnorm(p*d),d,p)
+
+Y<- matrix(rnorm(n*p, c(X%*%V), sd= sqrt(sigma2)), n, p)
+
+# plot(mu1)
+plot(X,col=M)
+
+###############################
+
+
+
+
+mclust<-Mclust(Y,G = K,modelNames = "VEI")
+raw_M<-apply(mclust$z, 1, function(x){c(1:K)[x==max(x)]})
+
+
+pca_M<-PCAC(Y,K,d=2)
+arc_M<- rDARC(Y,d = 2,k = K,meansIni = matrix(1,2),steps = 1E3,sigma2_ini = 0.1,randomStart = F,ver = 2)
+# arc_M<- ARC(Y,K,d = 2)
+
+pdf("simX.pdf",6,6)
[…]

### 2016-11-28 docs/draft.tex
added multithreading

@@
  initialization\;
- \While{Expected likelihood $\bb E \log L$ change $>$ threshold}{
+ \While{Change in expected likelihood $\Delta \bb E \log L$ $>$ threshold}{
   Using $\hat \Z \hat \bmu$, compute expectation: $\bb E \V$ and $\bb E \V\V'$\;
   Using expected values, maximize: $\hat \X$ \;
-   \While{change in $\hat \bmu$ $>$ threshold}{
+   \While{change in estimated $\Delta \hat \bmu$ $>$ threshold}{
    		  Compute MLEs: $\hat \bmu$ and $\hat \bSigma$\;

### 2016-11-28 jkmeans/src/mixture.hpp

@@ -55,8 +55,9 @@ class Mixture {
   }
 
   double compTotalLoglik() {
-    double marginal_loglik = 0;
+    vec marginal_loglik = zeros<vec>(K);
 
+#pragma omp parallel for
     for (int k = 0; k < K; ++k) {
       mat diff = (y - repmat(mu.row(k), n, 1));
 
@@ -65,13 +66,14 @@ class Mixture {
       double l = exp(-accu(diff2.each_row() / Sigma.t()) / 2.0 -
                      1 / 2.0 * accu(log(Sigma)));
 
-      marginal_loglik += w(k) * l;
+      marginal_loglik(k) = w(k) * l;
     }
 
-    return accu(log(marginal_loglik));
+    return log(accu(marginal_loglik));
   }
 
   void Expectation() {
+#pragma omp parallel for
     for (int i = 0; i < n; ++i) {
       for (int k = 0; k < K; ++k) {
         rowvec y_local = y.row(i);
@@ -100,6 +102,7 @@ class Mixture {
     double diff2 = 0;
     double sum_total_weights = 0;
 
+#pragma omp parallel for
     for (int k = 0; k < K; ++k) {
       vec weights = zeta.col(k);
       double total_weights = accu(weights);
@@ -118,21 +121,22 @@ class Mixture {
   }
 
   void updateSigma() {
-    vec diff2 = zeros(p);
-    double sum_total_weights = 0;
+    mat diff2K = zeros<mat>(p, K);
+    vec sum_total_weights = zeros<vec>(K);
+
+#pragma omp parallel for
     for (int k = 0; k < K; ++k) {
       mat diff = y.each_row() - mu.row(k);
 
       vec zeta_local = zeta.col(k);
 
-      diff2 += (zeta_local.t() * (diff % diff)).t();
+      diff2K.col(k) = (zeta_local.t() * (diff % diff)).t();
 
-      // diff2 += accu(sum(diff % diff, 1) % zeta_local);
       double total_weights = accu(zeta_local);
-      sum_total_weights += total_weights;
+      sum_total_weights(k) = total_weights;
     }
 
-    Sigma = (diff2 / sum_total_weights);
+    Sigma = sum(diff2K, 1) / accu(sum_total_weights);
   }
 
   void runEM(int steps, double tol) {

### 2016-11-28 test/mnist_test.r

@@ -10,8 +10,9 @@ mnist<- read.csv("../mnist_data/mnist.csv",header = F)
 pick<- c(0,3,7,9)
 # pick<- c(0:9)
 
-subset<- c(sapply(pick*1000, function(x){x+ c(1:100)*2}))
+subset<- c(sapply(pick*1000, function(x){x+ c(1:100)*5}))
 
+# subset<- c(sapply(pick*1000, function(x){x+ c(1:1000)}))
 
 mnist_subset<- mnist[,subset]
 Y<-t(mnist_subset)
@@ -19,7 +20,7 @@ Y<-t(mnist_subset)
 
 M<- rep(pick,each=100)
 
-
+# M<- rep(pick,each=1000)
 K<- 4
 
 # mclust<-Mclust(Y,G = K,modelNames = "VEI")
@@ -46,16 +47,46 @@ mean(ari)
 
 arc_ari<- numeric()
 for(i in 1:10){
-  arc_M<- rDARC(Y,d = 6,k = K,meansIni = matrix(1,2),steps = 1E3,sigma2_ini = 0.1,randomStart = F,ver = 1,fixW = T)
+  arc_M<- rDARC(Y,d = 6,k = K,meansIni = matrix(1,2),steps = 1E3,sigma2_ini = 0.1,randomStart = F,ver = 1,fixW = F)
   arc_ari<- c(arc_ari,  adjustedRandIndex(arc_M$M, M))
 }
+
+image(matrix(arc_M$mu[1,]%*%arc_M$EV,28,28))
+image(matrix(arc_M$mu[2,]%*%arc_M$EV,28,28))
+image(matrix(arc_M$mu[3,]%*%arc_M$EV,28,28))
+image(matrix(arc_M$mu[4,]%*%arc_M$EV,28,28))
+
+plot(arc_M$M)
+
+
+arc_M$X[4,]
+arc_M$mu[4,]
+
+image(matrix(arc_M$X[1,]%*%arc_M$EV,28,28))
+image(matrix(arc_M$mu[1,]%*%arc_M$EV,28,28))
+image(matrix(Y[1,],28,28))
+
+image(matrix(arc_M$X[101,]%*%arc_M$EV,28,28))
+image(matrix(arc_M$mu[2,]%*%arc_M$EV,28,28))
+image(matrix(Y[101,],28,28))
+
[…]

### 2016-11-28 test/rdARCtest.r

@@ -23,7 +23,7 @@ n<- 300
 p<- 1000
 d<- 2
 sigma2<- 1
-tau<- 0.1
+tau<- 0.2
 Sigma<- diag(tau,d)
 
 ###################################
@@ -66,8 +66,8 @@ mclust<-Mclust(Y,G = K,modelNames = "VEI")
 raw_M<-apply(mclust$z, 1, function(x){c(1:K)[x==max(x)]})
 
 
-pca_M<-PCAC(Y,K,d=2)
-arc_M<- rDARC(Y,d = 2,k = K,meansIni = matrix(1,2),steps = 1E3,sigma2_ini = 0.1,randomStart = F,ver = 2)
+pca_M<-PCAC(Y,K,d=3)
+arc_M<- rDARC(Y,d = 3,k = K,meansIni = matrix(1,2),steps = 1E3,sigma2_ini = 0.1,randomStart = F,ver = 2)
 # arc_M<- ARC(Y,K,d = 2)
 
 pdf("simX.pdf",6,6)

### 2016-12-05 docs/draft.tex
update draft

@@
 
-It is worth mentioning a different class of high dimensional clustering method, namely sparse clustering (see \cite{witten2012framework} and the references therein).
+%It is worth mentioning a different class of high dimensional clustering method, namely sparse clustering (see \cite{witten2012framework} and the references therein).
 The main idea is to select the subset of dimensions directly on the data space via sparsity constraint.
@@
 
-\subsection{Synthetic Data}
-Data generation:
+\subsection{Automatic Repulsion}
 
-n=300
-p=2000
-d=2
-mu=1,2,3
-sigma2=0.5
+Before we carry out clustering simulations, we first illustrate the effects of automatic repulsion.
+We generate a mixture of two Gaussian components with mean $1$ and $2$.
+We then change the variance in $0.3^2$, $0.5^2$ and $1^2$, so that the data exhibit different degree of overlap (Figure~\ref{fig:simRepulY}).
+We then use ARC to estimate the two component means.
+We repeat the experiment for $1,000$ times so that the distribution of the mean estimates can be visualized  (Figure~\ref{fig:simRepulM}).
+
+ When the data has little overlap (with variance  $0.3^2$), the mean estimate for the two components show almost no repulsion.
+As the variance increases (with variance  $0.5^2$), the mean estimates start to be pushed apart.
+When the overlapping becomes severe  (with variance  $1^2$), the two mean estimates are further  so that they still remain well separated.
+This behavior ensures the clustering generates clear boundary between clusters.
+The regularization is automatic and requires no tuning.
+
+\begin{figure}[H]
+ \centering
+  \begin{subfigure}[b]{1\textwidth}
+ \includegraphics[width=1\textwidth]{pics/repulData}
+  \caption{The distribution of the data.}
+\label{fig:simRepulY}
+\end{subfigure}
+  \hfill
+   \begin{subfigure}[b]{1\textwidth}
+ \includegraphics[width=1\textwidth]{pics/repulMean}
+  \caption{The distribution of the mean under the automatic repulsion.}
+  \label{fig:simRepulM}
+\end{subfigure}
+\caption{Simulation illustrating the automatic repulsion in the mean estimates, under different degree of overlap (standard deviation) in the generated data.}
+ \end{figure}
+
+\subsection{Clustering Synthetic Data}
+
+To illustrate the effect of clustering in low dimension, we simulate $3$ component data with $n_1=50$, $n_2=150$ and $n_3=100$ in two dimensions, with means $(1,2)$, $(2,1)$ and $(3,3)$.
+We set the variance to be $0.2$ so that the components are well separated.
+
+Then we project the matrix into $p$ dimensions via a random matrix $V$ with $V_{ij}\sim N(0,1)$.
+And we test three models (i) the Gaussian mixture model (GMM) directly on the the high dimension, (ii) GMM on the reduced dimension learned by PCA and (iii) ARC on adapted reduced dimension.
+We obtain the clustering labels estimated from the models and compute the  adjusted random index with respect to the true labels.
+Each experiment is repeated $10$ times with mean and standard deviation reported.
+
+We first simply vary $p$ from $50$ to $1,000$ to test the performance of models.
+Since each dimension contains signals, not surprisingly, the simple GMM on the large dimensions performs the best among all three models (Table~\ref{tabs1}, row 1), followed closely by the ARC model.
+However, such setting is very unrealistic as often most of the large dimensions are irrelevant.
+To simulate this setting, we keep the projected dimensions fixed at $20$ but add $(p-20)$ interfering dimensions from pure noise $y_{i,j}\sim N(0,1)$.
+The performance of the high dimensional GMM starts to deteriorate when the irrelevant dimensions to start to interfere with the signals (Table~\ref{tabs1}, row 2).
 
 \begin{table}[H]
+\small
 \centering
-	\begin{tabular}{| l | r |}
+	\begin{tabular}{| l | r | r| r| r| r| }
+			\multicolumn{2}{l}{Without irrelevant dimensions} \\
 	\hline
-			Model & $ARI$ \\
+			p & 50 & 100 & 200 & 500 &1000  \\
 	\hline
-			GMM on Principle Components &  0.01  \\
-			ARC on Principle Components		& 0.31 \\
-			GMM on Flexible Reduced Dimension &  0.35  \\
-			ARC on Flexible Reduced Dimension		& {\bf 0.44} \\
+			GMM &   {\bf 0.83} $\pm$ 0.04 &  {\bf 0.84} $\pm$ 0.04  &  {\bf 0.84} $\pm$ 0.03 &  {\bf 0.87} $\pm$ 0.03 &  {\bf 0.85} $\pm$ 0.05 \\
+			PCA+GMM		& 0.64 $\pm$ 0.20 & 0.77 $\pm$ 0.19 & 0.66 $\pm$ 0.29 & 0.79 $\pm$ 0.22 & 0.69 $\pm$ 0.30\\
+			ARC 	&  0.72 $\pm$ 0.12 & 0.82 $\pm$ 0.16 & 0.79 $\pm$ 0.14 & 0.73 $\pm$ 0.19 & 0.86 $\pm$ 0.13\\
+			\hline
+			\multicolumn{2}{l}{With ($p-20$) irrelevant dimensions } \\
+	\hline
+			p & 50 & 100 & 200 & 500 &1000  \\
+	\hline
+			GMM & 0.76 $\pm$ 0.03  &  0.69 $\pm$ 0.13     &  0.70 $\pm$ 0.22  &  0.63 $\pm$ 0.11 &  0.53 $\pm$ 0.07\\
+			PCA+GMM	&0.75 $\pm$ 0.23 	& 0.48 $\pm$ 0.25  &  0.63 $\pm$ 0.21   &  0.48 $\pm$ 0.24&   0.57 $\pm$ 0.21\\
+			ARC 	& {\bf 0.80} $\pm$ 0.11  & {\bf 0.75} $\pm$ 0.11 &  {\bf 0.83 } $\pm$ 0.10   &  {\bf 0.79} $\pm$ 0.09&  {\bf 0.76} $\pm$ 0.09\\
 			\hline
 	\end{tabular}
+	\caption{Clustering performance using adjusted random index with respect to the true labels.}
+	\label{tabs1}
 \end{table}
+
+In this regard, the reduced dimensional method is advantageous as it serves to discard the irrelevant dimensions.
+Nevertheless, clustering based on PCA-generated subspace almost never perform well.
+This can be explained by Figure~\ref{fig:sim_pca} that the first two principle components show little separation.
+This problem is solved with the repulsive regularization.
+As shown in  Figure~\ref{fig:sim_arc}, the discovered low dimension space demonstrates clear and wide boundaries among clusters, leading to superior performance in ARI.
 
@@
  \includegraphics[width=0.8\textwidth]{pics/simX}
-  \caption{The lower dimensions used in data generation.}
+  \caption{The true low dimensions in data generation.}
+\label{fig:sim_truth}
 \end{subfigure}
@@
   \caption{The lower dimensions discovered by PCA.}
+  \label{fig:sim_pca}
 \end{subfigure}
@@
   \caption{The lower dimensions discovered by ARC.}
+  \label{fig:sim_arc}
 \end{subfigure}
+	\caption{Scatterplots illustrating the learning of the low dimensional subspace using PCA and ARC.
+The ground truth, clustered labels under PCA+GMM and clustered labels under ARC are shown in colors.}
  \end{figure}
 
-\subsection{Real Data}
+\subsection{Clustering Real Data: Handwritten Digits Clustering}
 
-MNIST:
+We now test the model on real data with MNIST handwritten digits. The digits are centered in a 28x28 image. We normalize the pixel values by dividing 255.
 
-Four digits: $0,3,7,9$. Pick $100$ samples for each digit.
+Following the choice of digits \citep{lee2015bayesian} for clustering benchmark, we use the images of four digits: $0,3,7,9$.
+To mimic the low sample size and high dimension setting, we randomly draw $50$ samples for each digit.
+Each experiment is repeated 10 times for GMM applied on the data, PCA based GMM and ARC.
+The results are shown in Table~\ref{tabs2}.
+The ARC clearly shows clear advantages in clustering the four digits over the other methods.
 
-Compute average ARI over $10$ tests.
+For completeness, we also test the clustering models with all $10$ digits.
+Due to the similarity of the digits and noise in rotation, we observe a significant decrease in ARI.
+Nevertheless, ARC still dominates in the performance over the other methods.
+
+As comparison, we include the test result using sparsity regularized clustering (SparCL,\cite{witten2012framework}).
+Due to the nature of image data, the sparsity assumption is not quite suitable for this data, except for removing common blank background.
+As the result, its performance is almost the same as the simple GMM directly applied on the data.
 
 \begin{table}[H]
+\small
 \centering
-	\begin{tabular}{| l | r |}
+	\begin{tabular}{| l | r | r|}
 	\hline
-			Model & $ARI$ \\
+	Digits & $(0,3,7,9)$ & $0-9$ \\
 		\hline
-			GMM & 0.50 \\
-			PCA + GMM &  0.38  \\
-			ARC		& {\bf 0.57} \\
+			GMM &   0.37 $\pm$ 0.04   &   0.10 $\pm$ 0.11  \\
+			PCA+GMM		& 0.35 $\pm$ 0.14  &   0.24 $\pm$ 0.01  \\
+			SparCL 	&  0.36 $\pm$ 0.12 & 0.11 $\pm$ 0.13 \\
+			ARC 	&  {\bf 0.44} $\pm$ 0.12 &   {\bf 0.30} $\pm$ 0.05  \\
 			\hline
 		\end{tabular}
+	\label{tabs2}
+		\caption{Adjusted rand index as benchmark for the different clustering algorithms applied on the MNIST handwritten digits data.}
 \end{table}
@@
 \section{Discussion}
+
+With the large dimensions contaminated with irrelevant information, clustering with high dimensional data is challenging.
+Applying dimension reduction is an appealing direction and has recently gain more traction.
+Examples include \cite{niu2011dimensionality} combining dimension reduction with spectral clustering to separate data into multiple manifolds and \cite{jing2013dictionary} using dictionary-learning structure to find the low-dimension subspace.
+
+Since one faces the task of choosing top few dimensions, one critical issue remained, it to identify a subspace that it is ``optimal'' in some sense.
+In this article, we restrict the focus on the linear subspace, and seek the one where the clustered components have a clear separating boundary.
+This is a particular useful complement to model-based clustering, where the separation is originally missing in mixture density.
+
+With the tuning-free regularization, it is simple to achieve this task in our model.
+As demonstrated in the examples, it achieves quite satisfactory performance in both simulation and real data applications.
+One extension to the current work would be studying the robustness of the error structure.
+Another one is to combine the regularization with kernel based spectral clustering hence extends the method to non-linear space.
 

### 2016-12-05 jkmeans/src/jkmeans.cpp

@@ -46,9 +46,8 @@ Rcpp::List rDARC(const arma::mat& Y, int d, int k, const arma::mat& meansIni,
       Rcpp::Named("Sigma") = rdARC.mix->Sigma,
       Rcpp::Named("zeta") = rdARC.mix->zeta, Rcpp::Named("X") = rdARC.X,
       Rcpp::Named("EV") = rdARC.EV,
-      Rcpp::Named("M") = rdARC.mix->clusteringMAP()
-
-          );
+      Rcpp::Named("M") = rdARC.mix->clusteringMAP(),
+      Rcpp::Named("loglik") = rdARC.loglik);
 }
 
 // [[Rcpp::export]]

### 2016-12-05 jkmeans/src/reducedDimARC.hpp

@@ -13,6 +13,8 @@ class reducedDimARC {
   double sigma2;
   bool randomStart;
 
+  double loglik;
+
   std::shared_ptr<Mixture> mix;
   unsigned int ver;
 
@@ -38,8 +40,8 @@ class reducedDimARC {
     }
 
     if (randomStart) {
-      // X += randn(n, d);
-      X = randn(n, d);
+      X += randn(n, d);
+      // X = randn(n, d);
     }
 
     Sigma = eye(d, d);
@@ -93,10 +95,28 @@ class reducedDimARC {
     Sigma = diagmat(mix->Sigma);
   }
 
+  double computeLoglik() {
+    mat XZmu = X - Zmu;
+    double logdetSigma;
+    double sign;
+    log_det(logdetSigma, sign, Sigma);
+
+    return accu(XZmu % trans(solve(Sigma, XZmu.t()))) + n * logdetSigma +
+           n * p * log(sigma2) + n * p;
+  }
+
   void runEM(int steps) {
+    double cur_loglik = INFINITY;
     for (int i = 0; i < steps; ++i) {
       Estep();
       Mstep();
+      double new_loglik = computeLoglik();
+      cout << new_loglik << endl;
+      if (new_loglik > cur_loglik & i > 1)
+        break;
+      else
+        cur_loglik = new_loglik;
     }
+    loglik = cur_loglik;
   }
 };

### 2016-12-05 test/mnist_test.r

@@ -7,86 +7,73 @@ source("PCAC.R")
 
 mnist<- read.csv("../mnist_data/mnist.csv",header = F)
 
-pick<- c(0,3,7,9)
-# pick<- c(0:9)
+# pick<- c(0,3,7,9)
+pick<- c(0:9)
+n<- 50
+p<- 28^2
 
-subset<- c(sapply(pick*1000, function(x){x+ c(1:100)*5}))
 
 # subset<- c(sapply(pick*1000, function(x){x+ c(1:1000)}))
 
-mnist_subset<- mnist[,subset]
-Y<-t(mnist_subset)
 
-
-M<- rep(pick,each=100)
-
-# M<- rep(pick,each=1000)
-K<- 4
-
-# mclust<-Mclust(Y,G = K,modelNames = "VEI")
-# raw_M<-apply(mclust$z, 1, function(x){c(1:K)[x==max(x)]})
-
-# svd0<- svd(Y)
 
 ######
 
 #######
 
 
-mclust<-Mclust(Y,G = K)
-raw_M<-apply(mclust$z, 1, function(x){c(1:K)[x==max(x)]})
-adjustedRandIndex(raw_M, M)
+# mclust<-Mclust(Y,G = K)
 
-ari<- numeric()
-for(i in 1:10){
-  pca_M<-PCAC(Y,K,d = 6)
-  ari<- c(ari,adjustedRandIndex(pca_M$M, M))
+
+testMNIST<- function(n=100){
+  
+  s<- sample(1:1000,n,replace = F)
+  subset<- c(sapply(pick*1000, function(x){x+ s}))
+  mnist_subset<- mnist[,subset]
+  Y<-t(mnist_subset)/255
+  M<- rep(pick,each=n)
+  K<- length(pick)
+  # Y<- (Y-mean(Y))/sd(Y)
+  
+  
+  
+  raw<- jkmeansEM(Y,k = K,useKmeansIni = T,meansIni = matrix(rep(1,K*p),nrow = K),sigma2_ini = 0.1,fixW = F)
+  print("raw") 
+  pca<- PCAC(Y,K,d = 6)
+  print("pca") 
+  arc<- rDARC(Y,d = 6,k = K,meansIni = matrix(1,2),steps = 1E3,sigma2_ini = 0.1,randomStart = F,ver = 1)
+  print("arc") 
+  
[…]

### 2016-12-05 test/rdARCtest.r

@@ -20,7 +20,7 @@ Y<- matrix(rnorm(300*3,mu, sd = 0.5),300)
 
 
 n<- 300
-p<- 1000
+p<- 50
 d<- 2
 sigma2<- 1
 tau<- 0.2
@@ -34,58 +34,94 @@ mu1<- matrix(K,n,d)
 # mu1[1:length(randomMu)]<- randomMu
 # mu1<- mu1[1:n]
 M<- rep(c(1:K),each=n/K)
+M<- c( rep(1,n/K/2), rep(2,n/K/2*3), rep(3,n/K))
 
-mu1[1:(n/K),]<-1
-# mu1[1:(n/K),2]<-2
+mu1[1:(n/K/2),]<-1
+mu1[1:(n/K/2),2]<-2
 
-mu1[(n/K+1):(n/K*2),]<-2
-# mu1[(n/K+1):(n/K*2),2]<-1
+mu1[(n/K/2+1):(n/K*2),]<-2
+mu1[(n/K/2+1):(n/K*2),2]<-1
 
 mu1[(n/K*2+1):(n/K*3),]<-3
-# mu1[(n/K*2+1):(n/K*3),2]<-3
+mu1[(n/K*2+1):(n/K*3),2]<-3
 
 # mu1[1:50,]<- 1.5
+# plot(mu1)
 
 ###
-
+# 
 Zmu<- matrix(mu1,n,d)
 X<- matrix( rnorm(n*d, c(Zmu), sqrt(tau)),n,d)
 V<- matrix( rnorm(p*d),d,p)
-
+# 
 Y<- matrix(rnorm(n*p, c(X%*%V), sd= sqrt(sigma2)), n, p)
 
+
+# Zmu1<- matrix(Zmu,n,p)
+
+p_trunc<- p
+# p_trunc<- 50
+# Y[,p_trunc:p]<- rnorm(length(Y[,p_trunc:p]),mean=0,sd= sqrt(sigma2))
+
+# Zmu1[,p_trunc:p]<- 0
+# Y<- matrix( rnorm(n*p,c(Zmu1),sd= sqrt(sigma2)), n, p)
+
+
 # plot(mu1)
-plot(X,col=M)
 
 ###############################
 
 
 
 
-mclust<-Mclust(Y,G = K,modelNames = "VEI")
+mclust<-Mclust(Y[,1:(p_trunc-1)],G = K)
 raw_M<-apply(mclust$z, 1, function(x){c(1:K)[x==max(x)]})
 
[…]

### 2016-12-05 test/rdARCtestVaryingP.R

@@ -1,0 +1,150 @@
+require("jkmeans")
+require("mclust")
+setwd("~/git/jkmeans/test/")
+source("kmpp.r")
+source("PCAC.R")
+
+
+Y<- matrix(rnorm(1000),nrow = 100)
+
+mu<- rep(c(1:3),each=100)
+
+Y<- matrix(rnorm(300*3,mu, sd = 0.5),300)
+
+# fit<- rDARC(Y,d = 2,k = 2,meansIni = matrix(1,2))
+
+# kmpp(Y,3)
+
+
+#
+
+LowDTest<-function(p,   p_trunc= 100){
+  n<- 300
+  d<- 2
+  sigma2<- 1
+  tau<- 0.2
+  Sigma<- diag(tau,d)
+  
+  ###################################
+  
+  K<- 3
+  mu1<- matrix(K,n,d)
+  # randomMu<- unlist(sapply(1:K, function(x){rep(x, runif(1,n/10,n))}))
+  # mu1[1:length(randomMu)]<- randomMu
+  # mu1<- mu1[1:n]
+  M<- rep(c(1:K),each=n/K)
+  M<- c( rep(1,n/K/2), rep(2,n/K/2*3), rep(3,n/K))
+  
+  mu1[1:(n/K/2),]<-1
+  mu1[1:(n/K/2),2]<-2
+  
+  mu1[(n/K/2+1):(n/K*2),]<-2
+  mu1[(n/K/2+1):(n/K*2),2]<-1
+  
+  mu1[(n/K*2+1):(n/K*3),]<-3
+  mu1[(n/K*2+1):(n/K*3),2]<-3
+  
+  Zmu<- matrix(mu1,n,d)
+  X<- matrix( rnorm(n*d, c(Zmu), sqrt(tau)),n,d)
+  V<- matrix( rnorm(p*d),d,p)
+  # 
+  Y<- matrix(rnorm(n*p, c(X%*%V), sd= sqrt(sigma2)), n, p)
+  
+  
+  
+  # p_trunc<- p
+  Y[,p_trunc:p]<- rnorm(length(Y[,p_trunc:p]),mean=0,sd= sqrt(sigma2))
+  
+  # Zmu1[,p_trunc:p]<- 0
+  # Y<- matrix( rnorm(n*p,c(Zmu1),sd= sqrt(sigma2)), n, p)
+  
+  
+  # plot(mu1)
+  
+  
+  
+  
+  # mclust<-Mclust(Y[,1:(p_trunc-1)],G = K)
[…]

### 2016-12-05 test/repulsion_illstration.r

@@ -1,0 +1,104 @@
+require('ggplot2')
+
+sigma<- 0.5
+mu<- c(rep(1,100),rep(2,100))
+Y<- rnorm(200,mu,sigma)
+
+
+# dataHist<- data.frame("Center"=as.factor(mu),    "y"= Y)
+# ggplot(dataHist, aes(y, fill = Center)) + geom_histogram(alpha = 0.5,bins = 30,  position="identity")
+
+K<-2
+
+
+
+sigma<- 0.3
+repMu3<-sapply(c(1:2000), function(x){
+  Y<- rnorm(200,mu,sigma)
+  raw<- jkmeansEM(as.matrix(Y),k = K,useKmeansIni = T,meansIni = matrix(rep(1,2),nrow = K),sigma2_ini = 0.1,fixW = F)
+  raw$mu
+})
+
+gmmMu3<-sapply(c(1:2000), function(x){
+  Y<- rnorm(200,mu,sigma)
+  fi<- Mclust(as.matrix(Y),2,modelNames="E")
+  fi$parameters$mean
+})
+
+
+
+sigma<- 0.5
+repMu5<-sapply(c(1:2000), function(x){
+  Y<- rnorm(200,mu,sigma)
+  raw<- jkmeansEM(as.matrix(Y),k = K,useKmeansIni = T,meansIni = matrix(rep(1,2),nrow = K),sigma2_ini = 0.1,fixW = F)
+  raw$mu
+})
+
+
+gmmMu5<-sapply(c(1:2000), function(x){
+  Y<- rnorm(200,mu,sigma)
+  fi<- Mclust(Y,2,modelNames="E")
+  fi$parameters$mean
+})
+
+sigma<- 1
+repMu10<-sapply(c(1:2000), function(x){
+  Y<- rnorm(150,mu,sigma)
+  raw<- jkmeansEM(as.matrix(Y),k = K,useKmeansIni = T,meansIni = matrix(rep(1,2),nrow = K),sigma2_ini = 0.1,fixW = F)
+  raw$mu
+})
+
+
+gmmMu10<-sapply(c(1:2000), function(x){
+  Y<- rnorm(200,mu,sigma)
+  fi<- Mclust(Y,2,modelNames="E")
+  fi$parameters$mean
+})
+
+
+sigma<- 1.5
+repMu15<-sapply(c(1:2000), function(x){
+  Y<- rnorm(150,mu,sigma)
[…]

### 2016-12-05 test/shape_test.r

@@ -1,0 +1,68 @@
+require("jkmeans")
+require("mclust")
+setwd("~/git/jkmeans/test/")
+source("kmpp.r")
+source("PCAC.R")
+
+
+Y<- matrix(runif(10000,-10,10),ncol=2)
+
+ring<-function(a,b){
+  x1<- Y[,1]
+  x2<- Y[,2]
+  (x1^2+x2^2)>a^2 & (x1^2+x2^2)<b^2
+}
+
+M<- ring(0,1)*1 + ring(2,4)*2 + ring(5,7)*3
+subset<- (ring(0,1)|ring(2,4)|ring(5,7))
+
+Y<- Y[  subset,]
+M<- M[subset]
+plot(Y,col=M)
+
+###################################
+
+K<- 3
+
+###############################
+
+
+
+
+mclust<-Mclust(Y,G = K,modelNames = "VEI")
+raw_M<-apply(mclust$z, 1, function(x){c(1:K)[x==max(x)]})
+
+
+# pca_M<-PCAC(Y,K,d=2)
+
+arc_M<- rDARC(Y,d = 1,k = K,meansIni = matrix(1,2),steps = 1E3,sigma2_ini = 0.1,randomStart = T,ver = 1)
+# arc_M<- ARC(Y,K,d = 2)
+
+testARC<- lapply(c(1:10),function(x) rDARC(Y,d = 2,k = K,meansIni = matrix(1,2),steps = 1E3,sigma2_ini = 0.1,randomStart = F,ver = 1))
+
+l_loglik<- (sapply(testARC, function(x)x$loglik))
+
+arc_M<- testARC[[c(1:10)[l_loglik== min(l_loglik)]]]
+
+arc_M$loglik
+
+pdf("simX.pdf",6,6)
+plot(X,col=M,xlab="X1",ylab = "X2",lwd=2)
+dev.off()
+
+pdf("simPCAX.pdf",6,6)
+plot(pca_M$X,col=pca_M$M+1,xlab="X1",ylab = "X2",lwd=2)
+dev.off()
+
+pdf("simARCX.pdf",6,6)
+plot(arc_M$X,col=arc_M$M+1,xlab="X1",ylab = "X2",lwd=2,ylim=range(pca_M$X[,2]))
+dev.off()
+
+# plot(arc_M$mu,type = "p",lwd=10)
+
+adjustedRandIndex(raw_M,M)
[…]

### 2016-12-05 test/sim_test.R

@@ -1,0 +1,136 @@
+setwd("~/git/jkmeans/test/")
+source("kmpp.r")
+require("jkmeans")
+require("mclust")
+source("PCAC.R")
+
+
+mnist<- read.csv("../mnist_data/mnist.csv",header = F)
+
+# pick<- c(0,3,7,9)
+pick<- c(1,3)
+
+subset<- c(sapply(pick*1000, function(x){x+ c(1:100)}))
+
+# subset<- c(sapply(pick*1000, function(x){x+ c(1:1000)}))
+
+mnist_subset<- mnist[,subset]
+Y<-t(mnist_subset)
+
+
+M<- rep(pick,each=100)
+
+# M<- rep(pick,each=1000)
+K<- length(pick)
+
+# mclust<-Mclust(Y,G = K,modelNames = "VEI")
+# raw_M<-apply(mclust$z, 1, function(x){c(1:K)[x==max(x)]})
+
+# svd0<- svd(Y)
+
+######
+
+#######
+
+
+# mclust<-Mclust(Y,G = K)
+# raw_M<-apply(mclust$z, 1, function(x){c(1:K)[x==max(x)]})
+# adjustedRandIndex(raw_M, M)
+
+#ari<- numeric()
+#for(i in 1:10){
+#  pca_M<-PCAC(Y,K,d = 6)
+#  ari<- c(ari,adjustedRandIndex(pca_M$M, M))
+#}
+#max(ari)
+
+resultPCA<- lapply(c(1:10),function(x) PCAC(Y,K,d = 6))
+
+resultARC<- lapply(c(1:10),function(x) rDARC(Y,d = 6,k = K,meansIni = matrix(1,2),steps = 1E3,sigma2_ini = 0.1,randomStart = F,ver = 1))
+
+
+l_loglik<- (sapply(resultARC, function(x)x$loglik))
+arc_M<- resultARC[[c(1:10)[l_loglik== min(l_loglik)]]]
+
+
+# save(resultPCA,file="resultPCAMNIST.Rda")
+# save(resultARC,file="resultARCMNIST.Rda")
+
+# arc_M$loglik
+
+image(matrix(arc_M$mu[1,]%*%arc_M$EV,28,28))
+image(matrix(arc_M$mu[2,]%*%arc_M$EV,28,28))
+
+
+image(matrix(arc_M$X[1,]%*%arc_M$EV,28,28))
[…]

### 2016-12-19 docs/draft.tex
sync

@@
 
+\subsection{The Necessity of Dimension Reduction for High Dimensional Clustering}
+
+Before we introduce the method, we first examine the conditions that would lead to the failure of the clustering in high dimension.
+We first consider a simple case of weighted Euclidean distance, which is used in Gaussian mixture model with covariance restricted to diagonal matrix.
+
+\begin{theorem} (High Dimension Clustering Failure Rate)\\
+Assume the true parameters for the two cluster centers are $\{\mu_{11},\mu_{12},\ldots \mu_{1m},\mu_{0(m+1)},\mu_{0(m+2)},\ldots \mu_{0(p-m)}\}'$ and $\{\mu_{21},\mu_{22},\ldots \mu_{2m},\mu_{0(m+1)},\mu_{0(m+2)},\ldots \mu_{0(p-m)}\}'$, in which out of $p$ dimensions, $m$ dimensions are different.
+The parameter estiamtes for the two centers are $\{\hat\mu_{11},\hat\mu_{12},\ldots \hat\mu_{1m},\hat\mu_{1(m+1)},\hat\mu_{1(m+2)},\ldots \hat\mu_{1(p-m)}\}'$ and $\{\hat\mu_{21},\hat\mu_{22},\ldots \hat\mu_{2m},\hat\mu_{2(m+1)},\hat\mu_{2(m+2)},\ldots \hat\mu_{2(p-m)}\}'$ based on cluster size of $n_1>n_2$.
+Then for any point $\{ y_1,\ldots y_p\}'$, the ratio of assigning the data to cluster 1 over  cluster 2 is:
+
+$$ \exp \left ( - \frac{1}{2} \left [ \sum_{j=1}^m \{ (y_{j}-\hat\mu_{1j})^2 - (y_{j}-\hat\mu_{2j})^2\}/\sigma^2_j - \sum_{j=m+1}^{p} \{  (y_{j}-\hat\mu_{2j})^2 -(y_{j}-\hat\mu_{1j})^2 \}/\sigma^2_j \right]\right).$$
+When the noise dimension ${(p-m)\rightarrow \infty}$,
+
+$$\bb E \sum_{j=m+1}^{p} \{  (y_{j}-\hat\mu_{2j})^2 -(y_{j}-\hat\mu_{1j})^2\}/\sigma^2_j  = (p-m)(\frac{1}{n_2}-\frac{1}{n_1}) \rightarrow \infty $$
+\end{theorem}
+
+As the result, any data that is generated from cluster 2 would be mis-clustered with label 1.
+We use different cluster sizes as a condition to induce the different variances in the cluster means.
+This condition can be further relaxed for any small difference in the variance.
+The noise accumulates over the dimensions and quickly distorts the clustering.
+
+To illustrate, we use an arbitrary setting with $\mu_1= \{1, 0\ldots 0\}$ and $\mu_2= \{2, 0\ldots 0\}$, with $1000$ $0$'s as the noisy dimensions.
+We add small noise $N(0,0.1)$ to the $0$'s in $\mu_1$ and $N(0,0.101)$ in $\mu_2$, and assign the data $y=\{y_1,y_2,\ldots, y_p\}$ with $y_2\ldots y_p \sim N(0, 0.5)$.
+We set $y_1$ in range from $1.5$ to $5.0$ to see the mis-classification rate.
+
+Figure~\ref{fig:hd_cluster} shows the large error caused by the small difference in the mean estimates in the interfering dimension.
+Even at $y_1=2.0$, there is still a $40\%$ of misclassification error, which extends well above $2.0$.
+Therefore, we conclude that dimension reduction is necessary to exclude out the interfering dimensions.
+
+\begin{figure}[H]
+ \centering
+  \begin{subfigure}[b]{.45\textwidth}
+ \includegraphics[width=1\textwidth]{pics/misrate}
+\label{fig:simRepulY}
+\end{subfigure}
+\caption{Misclassification rate under the interference of the noisy dimensions.}
+\label{fig:hd_cluster}
+ \end{figure}
+
+\leo{extend to any metric?}
+\\
+\leo{input on simulation conditions?}
+
 \subsection{Clustering Under Reduced Dimension}
@@
 \end{theorem}
-The proof of this theorem is provided in the appendix.
 
@@
 As we show in the next section, the estimation can be carried out by replacing the expectation step in model-based clustering with a maximization step.
+
+\leo{Conjecture}
+\begin{theorem} (Effects of Repulsion on Classification)\\
+The effects of repulsion do not interfere the classification when the difference of $b_1$ and $b_2$ are bounded:
+$$ \exp \left ( - \frac{1}{2} \left [ \sum_{j=1}^m \{ (y_{j}-\hat\mu_{1j} + b_1)^2 - (y_{j}-\hat\mu_{2j} + b_2)^2\}/\sigma^2_j \right] \right).$$
+\end{theorem}
 

### 2016-12-19 jkmeans/R/RcppExports.R

@@ -1,12 +1,12 @@
-# This file was generated by Rcpp::compileAttributes
+# Generated by using Rcpp::compileAttributes() -> do not edit by hand
 # Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
 
 jkmeansEM <- function(y, k, meansIni, steps = 1000L, tol = 1E-8, fixW = FALSE, useKmeansIni = FALSE, sigma2_ini = 0.1) {
     .Call('jkmeans_jkmeansEM', PACKAGE = 'jkmeans', y, k, meansIni, steps, tol, fixW, useKmeansIni, sigma2_ini)
 }
 
-rDARC <- function(Y, d, k, meansIni, steps = 1000L, tol = 1E-8, randomStart = TRUE, fixW = FALSE, sigma2_ini = 0.1, ver = 1L) {
-    .Call('jkmeans_rDARC', PACKAGE = 'jkmeans', Y, d, k, meansIni, steps, tol, randomStart, fixW, sigma2_ini, ver)
+rDARC <- function(Y, d, k, meansIni, steps = 1000L, tol = 1E-8, randomStart = TRUE, fixW = FALSE, useEstep = TRUE, sigma2_ini = 0.1, ver = 1L) {
+    .Call('jkmeans_rDARC', PACKAGE = 'jkmeans', Y, d, k, meansIni, steps, tol, randomStart, fixW, useEstep, sigma2_ini, ver)
 }
 
 kmpp <- function(Y, k) {

### 2016-12-19 jkmeans/src/RcppExports.cpp

@@ -1,4 +1,4 @@
-// This file was generated by Rcpp::compileAttributes
+// Generated by using Rcpp::compileAttributes() -> do not edit by hand
 // Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
 
 #include <RcppArmadillo.h>
@@ -10,8 +10,8 @@ using namespace Rcpp;
 Rcpp::List jkmeansEM(const arma::mat& y, int k, const arma::mat& meansIni, int steps, double tol, bool fixW, bool useKmeansIni, double sigma2_ini);
 RcppExport SEXP jkmeans_jkmeansEM(SEXP ySEXP, SEXP kSEXP, SEXP meansIniSEXP, SEXP stepsSEXP, SEXP tolSEXP, SEXP fixWSEXP, SEXP useKmeansIniSEXP, SEXP sigma2_iniSEXP) {
 BEGIN_RCPP
-    Rcpp::RObject __result;
-    Rcpp::RNGScope __rngScope;
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
     Rcpp::traits::input_parameter< const arma::mat& >::type y(ySEXP);
     Rcpp::traits::input_parameter< int >::type k(kSEXP);
     Rcpp::traits::input_parameter< const arma::mat& >::type meansIni(meansIniSEXP);
@@ -20,16 +20,16 @@ BEGIN_RCPP
     Rcpp::traits::input_parameter< bool >::type fixW(fixWSEXP);
     Rcpp::traits::input_parameter< bool >::type useKmeansIni(useKmeansIniSEXP);
     Rcpp::traits::input_parameter< double >::type sigma2_ini(sigma2_iniSEXP);
-    __result = Rcpp::wrap(jkmeansEM(y, k, meansIni, steps, tol, fixW, useKmeansIni, sigma2_ini));
-    return __result;
+    rcpp_result_gen = Rcpp::wrap(jkmeansEM(y, k, meansIni, steps, tol, fixW, useKmeansIni, sigma2_ini));
+    return rcpp_result_gen;
 END_RCPP
 }
 // rDARC
-Rcpp::List rDARC(const arma::mat& Y, int d, int k, const arma::mat& meansIni, int steps, double tol, bool randomStart, bool fixW, double sigma2_ini, int ver);
[…]

### 2016-12-19 jkmeans/src/jkmeans.cpp

@@ -37,8 +37,9 @@ Rcpp::List jkmeansEM(const arma::mat& y, int k, const arma::mat& meansIni,
 // [[Rcpp::export]]
 Rcpp::List rDARC(const arma::mat& Y, int d, int k, const arma::mat& meansIni,
                  int steps = 1000, double tol = 1E-8, bool randomStart = true,
-                 bool fixW = false, double sigma2_ini = 0.1, int ver = 1) {
-  reducedDimARC rdARC(Y, d, k, randomStart, sigma2_ini, ver, fixW);
+                 bool fixW = false, bool useEstep = true,
+                 double sigma2_ini = 0.1, int ver = 1) {
+  reducedDimARC rdARC(Y, d, k, randomStart, sigma2_ini, ver, fixW, useEstep);
   rdARC.runEM(steps);
 
   return Rcpp::List::create(

### 2016-12-19 jkmeans/src/reducedDimARC.hpp

@@ -12,6 +12,7 @@ class reducedDimARC {
   mat Zmu;
   double sigma2;
   bool randomStart;
+  bool useEstep;
 
   double loglik;
 
@@ -19,7 +20,7 @@ class reducedDimARC {
   unsigned int ver;
 
   reducedDimARC(mat _Y, int _d, int _K, bool _randomStart, double sigma2_ini,
-                unsigned int _ver, bool fixW) {
+                unsigned int _ver, bool fixW, bool _useEstep) {
     ver = _ver;
 
     Y = _Y;
@@ -28,6 +29,7 @@ class reducedDimARC {
     n = Y.n_rows;
     p = Y.n_cols;
     randomStart = _randomStart;
+    useEstep = _useEstep;
 
     {
       mat U;
@@ -66,9 +68,14 @@ class reducedDimARC {
     mat XY = X.t() * Y;
     mat XXinv = inv(XX + eye(d, d) * 1E-5);
 
-    // expectation
+    // expectation or maximize
     EV = XXinv * XY;
-    EVV = p * sigma2 * XXinv + EV * EV.t();
+
+    // expectation
+    if (useEstep) EVV = p * sigma2 * XXinv + EV * EV.t();
+    // maximize
+    else
+      EVV = EV * EV.t();
   }
 
   void Mstep() {

### 2016-12-19 test/mis-clustering.R

@@ -1,0 +1,22 @@
+mean(1/rgamma(10000,2,1))
+hist(1/rgamma(10000,2,1))
+
+
+
+mrate<-function(y){
+test<-sapply(c(1:1000),function(x){
+y<- c(y, rnorm(1000,0,0.5))
+
+mu1<- c(1,rnorm(1000,0,0.1))
+mu2<- c(2,rnorm(1000,0,0.11))
+exp( - (sum((y-mu1)^2) - sum((y-mu2)^2)))
+})
+sum(test>1)/1000
+}
+
+result<-sapply(seq(1.5,5.0,by = 0.1), mrate)
+
+pdf("misrate.pdf",4,3)
+plot(seq(1.5,5.0,by = 0.1),result,type = "l")
+abline(v=2,lty=2,col="red")
+dev.off()

### 2016-12-19 test/mnist_test.r

@@ -7,8 +7,8 @@ source("PCAC.R")
 
 mnist<- read.csv("../mnist_data/mnist.csv",header = F)
 
-# pick<- c(0,3,7,9)
-pick<- c(0:9)
+pick<- c(0,3,7,9)
+# pick<- c(0:9)
 n<- 50
 p<- 28^2
 
@@ -17,9 +17,6 @@ p<- 28^2
 
 
 
-######
-
-#######
 
 
 # mclust<-Mclust(Y,G = K)
@@ -30,29 +27,37 @@ testMNIST<- function(n=100){
   s<- sample(1:1000,n,replace = F)
   subset<- c(sapply(pick*1000, function(x){x+ s}))
   mnist_subset<- mnist[,subset]
-  Y<-t(mnist_subset)/255
+  Y<-t(mnist_subset)
+  Y= Y+rnorm(length(Y),mean = 0,sd = 20)
+  Y= Y/255
   M<- rep(pick,each=n)
   K<- length(pick)
+  
   # Y<- (Y-mean(Y))/sd(Y)
   
-  
-  
   raw<- jkmeansEM(Y,k = K,useKmeansIni = T,meansIni = matrix(rep(1,K*p),nrow = K),sigma2_ini = 0.1,fixW = F)
+  adjustedRandIndex(raw$M,M)
   print("raw") 
   pca<- PCAC(Y,K,d = 6)
+  adjustedRandIndex(pca$M, M)
   print("pca") 
-  arc<- rDARC(Y,d = 6,k = K,meansIni = matrix(1,2),steps = 1E3,sigma2_ini = 0.1,randomStart = F,ver = 1)
+  arcM<- rDARC(Y,d = 6,k = K,meansIni = matrix(1,2),steps = 1E3,sigma2_ini = 0.1,randomStart = F,ver = 1,useEstep = F)
+  arc<- rDARC(Y,d = 6,k = K,meansIni = matrix(1,2),steps = 1E3,sigma2_ini = 0.1,randomStart = F,ver = 1,useEstep = T)
+  
+  adjustedRandIndex(arc$M,M)
+  
   print("arc") 
   
   c(adjustedRandIndex(raw$M,M),
     adjustedRandIndex(pca$M, M),
+    adjustedRandIndex(arcM$M, M),
     adjustedRandIndex(arc$M, M))
 }
 
-t50<- sapply(c(1:10), function(x)testMNIST(50))
[…]

### 2017-01-02 test/eigenface.Rmd
eigenface

```{r}
require("pixmap")
require("jkmeans")
require("mclust")

setwd("~/git/jkmeans/lfwcrop_face/lfwcrop_grey/pick/")

faces<- sapply(list.files("."),function(f){
  x=read.pnm(file = f)
  c(x@grey)
})

mFaces<- 0# rowMeans(faces)
faces<- faces-mFaces

M<- sapply(list.files("."), function(x){substr(x,1,5)})

X<- t(faces)

nf<- 64

K<- 2

plotFace<- function(x){
  image(t(matrix(rev(x),64,64)),col = grey(seq(0, 1, length = 256)))
}

raw<- jkmeansEM(X,k = K,useKmeansIni = T,meansIni = matrix(rep(1,2*10),nrow = 2),sigma2_ini = 0.1,fixW = F)

plotFace(raw$mu[1,] + mFaces)
plotFace(raw$mu[2,]+ mFaces)

plot(raw$M,title="classification")

adjustedRandIndex(raw$M,M)

arcM<- rDARC(X,d = 9,k = K,meansIni = matrix(1,2),steps = 2E3,sigma2_ini = 1,randomStart = F,ver = 2,useEstep = F,tol = 1E-9)

adjustedRandIndex(arcM$M,M)

plotFace(arcM$mu[1,]%*%arcM$EV+ mFaces)
plotFace(arcM$mu[2,]%*%arcM$EV+ mFaces)

plot(arcM$M,title="classification")
```

### 2017-01-02 test/eigenface2.Rmd
eigenface

--- 2017-01-02 test/eigenface.Rmd
@@
 M<- sapply(list.files("."), function(x){substr(x,1,5)})
+faces<- faces[,M=="Alvar"]
 
@@
 
-K<- 2
+K<- 4
 
@@
 plotFace(raw$mu[2,]+ mFaces)
+plotFace(raw$mu[3,]+ mFaces)
+plotFace(raw$mu[4,]+ mFaces)
+# plotFace(raw$mu[5,]+ mFaces)
+# plotFace(raw$mu[6,]+ mFaces)
+# plotFace(raw$mu[7,]+ mFaces)
+# plotFace(raw$mu[8,]+ mFaces)
 
@@
 
-adjustedRandIndex(raw$M,M)
-
 arcM<- rDARC(X,d = 9,k = K,meansIni = matrix(1,2),steps = 2E3,sigma2_ini = 1,randomStart = F,ver = 2,useEstep = F,tol = 1E-9)
-
-adjustedRandIndex(arcM$M,M)
 
@@
 plotFace(arcM$mu[2,]%*%arcM$EV+ mFaces)
+plotFace(arcM$mu[3,]%*%arcM$EV+ mFaces)
+plotFace(arcM$mu[4,]%*%arcM$EV+ mFaces)
 

### 2017-01-02 test/eigenface.r

@@ -1,0 +1,44 @@
+require("pixmap")
+require("jkmeans")
+require("mclust")
+
+setwd("~/git/jkmeans/lfwcrop_face/lfwcrop_grey/pick/")
+
+faces<- sapply(list.files("."),function(f){
+  x=read.pnm(file = f)
+  c(x@grey)
+})
+
+
+M<- sapply(list.files("."), function(x){substr(x,1,5)})
+
+X<- t(faces)
+
+nf<- 64
+
+K<- 2
+
+plotFace<- function(x){
+  image(matrix(x,64,64),col = grey(seq(0, 1, length = 256)))
+}
+
+raw<- jkmeansEM(X,k = K,useKmeansIni = T,meansIni = matrix(rep(1,2*10),nrow = 2),sigma2_ini = 0.1,fixW = F)
+
+plotFace(raw$mu[1,])
+plotFace(raw$mu[2,])
+
+
+plot(raw$M)
+
+adjustedRandIndex(raw$M,M)
+
+
+arcM<- rDARC(X,d = 9,k = K,meansIni = matrix(1,2),steps = 5E3,sigma2_ini = 0.1,randomStart = T,ver = 1,useEstep = F)
+
+adjustedRandIndex(arcM$M,M)
+
+
+plotFace(arcM$mu[1,]%*%arcM$EV)
+plotFace(arcM$mu[2,]%*%arcM$EV)
+
+plot(arcM$M)

### 2017-01-02 test/mis-clustering.R

@@ -4,19 +4,19 @@ hist(1/rgamma(10000,2,1))
 
 
 mrate<-function(y){
-test<-sapply(c(1:1000),function(x){
-y<- c(y, rnorm(1000,0,0.5))
-
-mu1<- c(1,rnorm(1000,0,0.1))
-mu2<- c(2,rnorm(1000,0,0.11))
-exp( - (sum((y-mu1)^2) - sum((y-mu2)^2)))
-})
-sum(test>1)/1000
+  test<-sapply(c(1:1000),function(x){
+    y<- c(y, rnorm(1000,0,0.5))
+    
+    mu1<- c(1,rnorm(1000,0,0.1))
+    mu2<- c(2,rnorm(1000,0,0.11))
+    exp( - (sum((y-mu1)^2) - sum((y-mu2)^2)))
+  })
+  sum(test>1)/1000
 }
 
 result<-sapply(seq(1.5,5.0,by = 0.1), mrate)
 
-pdf("misrate.pdf",4,3)
-plot(seq(1.5,5.0,by = 0.1),result,type = "l")
+pdf("misrate.pdf",5,5)
+plot(seq(1.5,5.0,by = 0.1),result,type = "l",xlab="y1",ylab="Misclassification Rate")
 abline(v=2,lty=2,col="red")
 dev.off()

### 2017-01-03 test/eigenface_spec.R
eigenface

@@ -1,0 +1,85 @@
+require("pixmap")
+require("jkmeans")
+require("mclust")
+
+setwd("~/git/jkmeans/lfwcrop_face/lfwcrop_grey/pick/")
+
+faces<- sapply(list.files("."),function(f){
+  x=read.pnm(file = f)
+  c(x@grey)
+})
+
+
+M<- sapply(list.files("."), function(x){substr(x,1,5)})
+
+X<- t(faces)
+
+nf<- 64
+
+K<- 2
+
+plotFace<- function(x){
+  image(matrix(x,64,64),col = grey(seq(0, 1, length = 256)))
+}
+
+raw<- jkmeansEM(X,k = K,useKmeansIni = T,meansIni = matrix(rep(1,2*10),nrow = 2),sigma2_ini = 0.1,fixW = F)
+
+plotFace(raw$mu[1,])
+plotFace(raw$mu[2,])
+
+
+plot(raw$M)
+
+adjustedRandIndex(raw$M,M)
+
+
+arcM<- rDARC(X,d = 9,k = K,meansIni = matrix(1,2),steps = 5E3,sigma2_ini = 0.1,randomStart = F,ver = 1,useEstep = F)
+
+adjustedRandIndex(arcM$M,M)
+
+
+plotFace(arcM$mu[1,]%*%arcM$EV)
+plotFace(arcM$mu[2,]%*%arcM$EV)
+
+plot(arcM$M)
+
+
+
+specClust<-function(X1,K,tau=50){
+  N<- nrow(X1)
+  dist<- matrix(0,N,N)
+  for(i in 1:N){
+    for(j in 1:i){
+      dist[i,j]<- exp(-tau* sum((X1[i,]-X1[j,])^2))
+      dist[j,i]<- dist[i,j]
+    }
+  }
+  
+  D<- (colSums(dist))
+  L<- t(dist/sqrt(D))/sqrt(D)
+
+  svdL<- svd(L,nu = K)
+  X2<- svdL$u
+  
+  nX2<- X2/ sqrt( rowSums(X2^2))
+  
+  clust<-kmeans(nX2,K)
+  clust$cluster
+}
+
+
+spcM<-specClust(X,2, 0.0001)
+
+adjustedRandIndex(spcM,M)
+
+plot(spcM)
+plot(as.numeric(as.factor(M)))
+
+miscError<- function(x){
+  rate=sum(x != as.numeric(as.factor(M)))/N
[…]

### 2017-01-03 test/spectralClust.R

@@ -1,0 +1,41 @@
+N<- 1000
+X<- matrix(rnorm(N*2),N)
+
+circle<-function(a,b){
+ radius<- (X[,1]^2 +X[,2]^2)
+ radius>a^2 & radius<b^2
+}
+
+X1<- rbind( X[circle(1,1.5),], X[circle(0,0.6),])
+
+plot(X1)
+
+N<- nrow(X1)
+dist<- matrix(0,N,N)
+for(i in 1:N){
+  for(j in 1:i){
+    dist[i,j]<- exp(  - 100* sum((X1[i,]-X1[j,])^2))
+    dist[j,i]<- dist[i,j]
+  }
+}
+
+D<- (colSums(dist))
+L<- t(dist/sqrt(D))/sqrt(D)
+# image(L)
+
+K<- 2
+
+svdL<- svd(L,nu = K)
+X2<- svdL$u
+
+nX2<- X2/ sqrt( rowSums(X2^2))
+
+# plot(X2)
+plot(nX2)
+
+clust<-kmeans(nX2,K)
+
+plot(nX2,col=clust$cluster)
+
+
+plot(X1,col=clust$cluster)

### 2017-01-09 test/mnist_test.Rmd
tested GMM/kmeans/1-max on eigenface mnist

---
title: "GMM, ECM and kmeans"
author: "Author"
date: '`r Sys.Date()`'
output:
  md_document: 
    variant: markdown_github
---

```{r}
# .rs.restartR()
setwd("~/git/jkmeans/test/")
source("kmpp.r")
require("jkmeans")
require("mclust")

mnist<- read.csv("../mnist_data/mnist.csv",header = F)

pick<- c(0:5)
p<- 28^2

n<-100
s<- sample(1:1000,n,replace = F)

subset<- c(sapply(pick*1000, function(x){x+ s}))
mnist_subset<- mnist[,subset]
Y<-t(mnist_subset)
Y<- Y/255
K<- length(pick)

gmm<- jkmeansEM(Y,k = K,useKmeansIni = T,meansIni = matrix(rnorm(K*p),nrow = K),sigma2_ini = 0.1,fixW = F,tau = 1,steps = 1000)

jk<- jkmeansEM(Y,k = K,useKmeansIni = T,meansIni = matrix(rep(1,K*p),nrow = K),sigma2_ini = 0.1,fixW = F,tau = 1000)

kmeans <- jkmeansEM(Y,k = K,useKmeansIni = T,meansIni = matrix(rep(1,K*p),nrow = K),sigma2_ini = 0.1,fixW = T,tau = 1000)

M<- rep(pick,each=n)

adjustedRandIndex(jk$M,M)
adjustedRandIndex(gmm$M,M)
adjustedRandIndex(kmeans$M,M)

plot(jk$M)
```

### 2017-01-09 test/mnist_test.r

@@ -1,84 +1,101 @@
+# .rs.restartR()
 setwd("~/git/jkmeans/test/")
 source("kmpp.r")
 require("jkmeans")
 require("mclust")
-source("PCAC.R")
+
 
 
 mnist<- read.csv("../mnist_data/mnist.csv",header = F)
 
-pick<- c(0,3,7,9)
-# pick<- c(0:9)
-n<- 50
+# pick<- c(0,3,7,9)
+pick<- c(0:3)
 p<- 28^2
 
 
 # subset<- c(sapply(pick*1000, function(x){x+ c(1:1000)}))
 
+n<-100
+s<- sample(1:1000,n,replace = F)
+# s<- c(1:1000)
+
+subset<- c(sapply(pick*1000, function(x){x+ s}))
+mnist_subset<- mnist[,subset]
+Y<-t(mnist_subset)
+Y<- Y/255
+K<- length(pick)
+
+jk<- jkmeansEM(Y,k = K,useKmeansIni = T,meansIni = matrix(rep(1,K*p),nrow = K),sigma2_ini = 0.1,fixW = F,tau = 1000)
+
+gmm<- jkmeansEM(Y,k = K,useKmeansIni = T,meansIni = matrix(rnorm(K*p),nrow = K),sigma2_ini = 0.1,fixW = F,tau = 1,steps = 1000)
+
+kmeans <- jkmeansEM(Y,k = K,useKmeansIni = T,meansIni = matrix(rep(1,K*p),nrow = K),sigma2_ini = 0.1,fixW = T,tau = 1000)
 
 
+M<- rep(pick,each=n)
 
+adjustedRandIndex(jk$M,M)
+adjustedRandIndex(gmm$M,M)
+adjustedRandIndex(kmeans$M,M)
 
-# mclust<-Mclust(Y,G = K)
+plot(jk$M)
 
-
-testMNIST<- function(n=100){
-  
-  s<- sample(1:1000,n,replace = F)
-  subset<- c(sapply(pick*1000, function(x){x+ s}))
-  mnist_subset<- mnist[,subset]
-  Y<-t(mnist_subset)
-  Y= Y+rnorm(length(Y),mean = 0,sd = 20)
-  Y= Y/255
-  M<- rep(pick,each=n)
-  K<- length(pick)
-  
-  # Y<- (Y-mean(Y))/sd(Y)
-  
[…]

### 2017-01-09 test/repulsion_illstration.r

@@ -1,5 +1,5 @@
 require('ggplot2')
-
+require("jkmeans")
 sigma<- 0.5
 mu<- c(rep(1,100),rep(2,100))
 Y<- rnorm(200,mu,sigma)
@@ -10,12 +10,10 @@ Y<- rnorm(200,mu,sigma)
 
 K<-2
 
-
-
 sigma<- 0.3
 repMu3<-sapply(c(1:2000), function(x){
   Y<- rnorm(200,mu,sigma)
-  raw<- jkmeansEM(as.matrix(Y),k = K,useKmeansIni = T,meansIni = matrix(rep(1,2),nrow = K),sigma2_ini = 0.1,fixW = F)
+  raw<- jkmeansEM(as.matrix(Y),k = K,useKmeansIni = T,meansIni = matrix(rep(1,2),nrow = K),sigma2_ini = 0.1,fixW = F,tau=10)
   raw$mu
 })
 
@@ -44,7 +42,7 @@ gmmMu5<-sapply(c(1:2000), function(x){
 sigma<- 1
 repMu10<-sapply(c(1:2000), function(x){
   Y<- rnorm(150,mu,sigma)
-  raw<- jkmeansEM(as.matrix(Y),k = K,useKmeansIni = T,meansIni = matrix(rep(1,2),nrow = K),sigma2_ini = 0.1,fixW = F)
+  raw<- jkmeansEM(as.matrix(Y),k = K,useKmeansIni = T,meansIni = matrix(rep(1,2),nrow = K),sigma2_ini = 0.1,fixW = F, tau = 10)
   raw$mu
 })
 
@@ -55,21 +53,21 @@ gmmMu10<-sapply(c(1:2000), function(x){
   fi$parameters$mean
 })
 
-
-sigma<- 1.5
-repMu15<-sapply(c(1:2000), function(x){
-  Y<- rnorm(150,mu,sigma)
-  raw<- jkmeansEM(as.matrix(Y),k = K,useKmeansIni = T,meansIni = matrix(rep(1,2),nrow = K),sigma2_ini = 0.1,fixW = F)
-  raw$mu
-})
-
-
-gmmMu15<-sapply(c(1:2000), function(x){
-  Y<- rnorm(200,mu,sigma)
-  fi<- Mclust(Y,2,modelNames="E")
-  fi$parameters$mean
-})
-
+# 
+# sigma<- 1.5
+# repMu15<-sapply(c(1:2000), function(x){
[…]

### 2017-01-09 test/spectralClust.R

@@ -30,12 +30,12 @@ X2<- svdL$u
 
 nX2<- X2/ sqrt( rowSums(X2^2))
 
-# plot(X2)
-plot(nX2)
+plot(X2)
+# plot(nX2)
 
 clust<-kmeans(nX2,K)
 
-plot(nX2,col=clust$cluster)
+plot(X2,col=clust$cluster)
 
 
 plot(X1,col=clust$cluster)

### 2017-01-09 test/breast_cancer.Rmd
added Wisconsin breast cancer test

---
title: "GMM, ECM and kmeans"
author: "Author"
date: '`r Sys.Date()`'
output:
  md_document: 
    variant: markdown_github
---

```{r}

require("jkmeans")
require("mclust")

setwd("~/git/jkmeans/wdbc/")

wdbc<- read.csv("./wdbc.data.txt",header = F,stringsAsFactors = F)

M<- wdbc$V2
X<- wdbc[,3:ncol(wdbc)]
X<- as.matrix(X)

table(M)
K<- 2

gmm<- jkmeansEM(X,k = K,useKmeansIni = T,meansIni = matrix(rep(1,2*10),nrow = 2),sigma2_ini = 1,fixW = F, tau = 1)

kmeans<- jkmeansEM(X,k = K,useKmeansIni = T,meansIni = matrix(rep(1,2*10),nrow = 2),sigma2_ini = 1,fixW = T, tau = 1000)

jk<- jkmeansEM(X,k = K,useKmeansIni = T,meansIni = matrix(rep(1,2*10),nrow = 2),sigma2_ini = 1,fixW = F,   tau = 1000)

jk_point1<- jkmeansEM(X,k = K,useKmeansIni = T,meansIni = matrix(rep(1,2*10),nrow = 2),sigma2_ini = 1,fixW = F,   tau = 0.1)

gmm$zeta[2,]
kmeans$zeta[2,]
jk$zeta[2,]
jk_point1$zeta[2,]

adjustedRandIndex(gmm$M,M)
adjustedRandIndex(kmeans$M,M)
adjustedRandIndex(jk$M,M)
adjustedRandIndex(jk_point1$M,M)

```

### 2017-01-09 test/breast_cancer.r

@@ -1,0 +1,34 @@
+require("jkmeans")
+require("mclust")
+
+
+
+setwd("~/git/jkmeans/wdbc/")
+
+wdbc<- read.csv("./wdbc.data.txt",header = F,stringsAsFactors = F)
+
+M<- wdbc$V2
+X<- wdbc[,3:ncol(wdbc)]
+X<- as.matrix(X)
+
+table(M)
+K<- 2
+
+gmm<- jkmeansEM(X,k = K,useKmeansIni = T,meansIni = matrix(rep(1,2*10),nrow = 2),sigma2_ini = 1,fixW = F, tau = 1)
+
+kmeans<- jkmeansEM(X,k = K,useKmeansIni = T,meansIni = matrix(rep(1,2*10),nrow = 2),sigma2_ini = 1,fixW = T, tau = 1000)
+
+jk<- jkmeansEM(X,k = K,useKmeansIni = T,meansIni = matrix(rep(1,2*10),nrow = 2),sigma2_ini = 1,fixW = F,   tau = 1000)
+
+jk_point1<- jkmeansEM(X,k = K,useKmeansIni = T,meansIni = matrix(rep(1,2*10),nrow = 2),sigma2_ini = 1,fixW = F,   tau = 0.1)
+
+
+gmm$zeta[2,]
+kmeans$zeta[2,]
+jk$zeta[2,]
+jk_point1$zeta[2,]
+
+adjustedRandIndex(gmm$M,M)
+adjustedRandIndex(kmeans$M,M)
+adjustedRandIndex(jk$M,M)
+adjustedRandIndex(jk_point1$M,M)
